### 2021-03-22 bem.tex
First commit

\section{Boundary element methods with Bempp}
In this section we want to provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretisation and solution.

The most simple boundary integral equation is of the form
\begin{equation}
	\label{eq:bnd_integral}
	\int_{\Gamma} g(\bx, \by)\phi(\by)ds(\by) = f(\bx),~\bx\in\Gamma.
\end{equation}
The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$. To provide a concrete example. To compute the electrostatic capacity of an object $\Omega$ one solves the above equation with $f(\bx)$ the constant function $1$ and $g(\bx,\by)=\frac{1}{4\pi|\bx - \by|}$. The normalized capacity is then obtained as $c = \frac{1}{4\pi}\int_{\Gamma}\phi(\bx)ds(\bx)$. Many practical problems have a significantly more complex structure and can involve block systems of integral equations. Nevertheless, the fundamental structure of what Bempp-cl does is well described by the above problem and will be described in the following.

\begin{figure}
	\center
	\includegraphics[width=5cm]{img/sphere}
	\label{fig:triangulation}
	\caption{Discretisation of a sphere into flat surface triangles.}
\end{figure}

The first step is the \textbf{definition of the surface} $\Gamma$. Surfaces are represented in Bempp-cl as a triangulation into flat triangles (see \ref{fig:triangulation}). The triangulation is internally represented as an array of node coordinates and an associated connectivity array of node indices that define each triangle. In this step also topology data is computed. In particular, for each triangle the neighboring triangles and the type of intersection (i.e. via joint node or edge) is computed.

Once a triangulation is given we need to define the necessary data structures for the discretisation. Bempp-cl uses a Galerkin discretisation. Consider \eqref{eq:bnd_integral}. We represent the solution $\phi$ as $\phi=\sum_{j=}^N x_j\phi_j$, where the $\phi_i$ are basis functions of a \textbf{trial space}, defined in the most simple case as constant $1$ on the triangle $\tau_j$ and $0$ everywhere also. Several other types of discretisation spaces are available in Bempp. Moreover, we require a \textbf{test (or dual) space} for the discretisation. Let $\psi=\sum_{i=1}^Ny_i \psi_i$, where each $\psi_i$ is a basis function of a space of test functions. For simplicity, here we will assume that the $\psi_i$ are also piecewise constant. The discrete representation of the above problem then takes the form

$$
Ax = b
$$
with 
$$
A_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)ds(\by)ds(\bx)
$$
and $f_i = \int_{\Gamma}\psi_i(\bx)f(\bx)ds(\bx)$. In the case of piecewise constant trial and test functions the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)ds(\by)ds(\bx)$.

In Bempp-cl an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{Grid function object}, which
consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)ds(x)$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.

Once the grid, the space objects, and the operator are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $A_{ij}$. For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $A_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights. In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical, corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.

The values $b_i$ of the right-hand side vector $b$ are similarly computed through a numerical quadrature rule.

In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as Gmres. The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.

In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp
\begin{enumerate}
	\item Import of the surface description as triangulation data
	\item Definition of the trial space, test space and range space
	\item Discretization into a matrix problem $Ax=b$
	\item Evaluation of domain potential operators for visualization and post-processing
\end{enumerate}

All of these steps are accelerated through the use of either Numba or OpenCL. In the following section we provide a high-level overview of the library and how these acceleration techniques are deployed before we deep dive into the design of the computational kernels.

### 2021-03-22 intro.tex
First commit

\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations. The original release came with a simple Python wrapper to the C++ interface. Over time more and more functionality was moved into the Python interface. However, computationally intensive routines and the main data structures remained in C++. This proved a burden when efforts started to modernise the library to be able to make better use of Single-Instruction-Multiple-Data (SIMD) optimisation on CPUs and offload computation to GPUs. In 2018 we decided to rewrite the computational core from scratch. The aims were to support explicit SIMD optimisation on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python. These aims naturally lead to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba. The name of the new library was Bempp-cl. At the end of 2019 we released an initial version 0.1 that we used to test the library in a number of real-world scenarios. This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use. Since then we have used Bempp-cl in a number of practical application and many of our users are slowly migrating to it from the old C++ based Bempp. In this article we want to discuss the design choices behind Bempp-cl and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.

[To be continued with some more details about content and overview of the Sections]

### 2021-03-22 main.tex
First commit

\documentclass{IEEEcsmag}

\usepackage[colorlinks,urlcolor=blue,linkcolor=blue,citecolor=blue]{hyperref}

\usepackage{upmath}
\usepackage{amsmath, amssymb}

\def\bx{\mathbf{x}}
\def\by{\mathbf{y}}

\jvol{XX}
\jnum{XX}
\paper{8}
\jmonth{May/June}
\jname{IT Professional}
\pubyear{2019}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\setcounter{secnumdepth}{0}

\begin{document}

\sptitle{Department: Head}
\editor{Editor: Name, xxxx@email}

\title{Designing a high-performance boundary element library with OpenCL and Numba}

\author{T. Betcke}
\affil{Department of Mathematics, University College London}

\author{M. W. Scroggs}
\affil{Department of Engineering, University of Cambridge}

\markboth{Department Head}{Paper title}

\begin{abstract}
The Bempp boundary element library is a well known library for the simulation of a range of electrostatic, acoustic and electromagnetic problems in homogeneous bounded and unbounded domains. It originally started as a traditional C++ library with a Python interface. Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba. The OpenCL backend implements kernels for GPUs and SIMD optimised for CPUs. In this paper we will discuss the design of Bempp-cl, provide performance comparisons on different compute devices and discuss the advantages and disadvantages of OpenCL as compared to Numba.
\end{abstract}

\maketitle

\input{intro}
\input{bem}

\end{document}

\section{COPYRIGHT AND CLEARANCE}

All CS Magazine authors must obtain clearance from IEEE Computer Society before submitting the final manuscript. The ``\href{https://apps.na.collabserv.com/wikis/home?lang=en-us#!/wiki/W18e544042a85_4b63_915a_1d1ed2cf8338/page/Publication\%20clearance}{Publication Clearance}'' wiki provides details about the procedure. Computer Society employees must use the \href{https://mc.manuscriptcentral.com/cs-ieee}{ScholarOne  Manuscripts Clearance} System to obtain publication approval.

\section{SECTIONS}

Sections following the introduction should present your results and findings. The body of the paper should be approximately 6,000 words. The manuscript should evolve so that each sentence, equation, figure, and table flow smoothly and logically from whatever precedes it. Relevant work by others, as well as relevant products from other companies, should be adequately and accurately cited. Sufficient support should be provided (or cited) for the assertions made and conclusions drawn.

Headings may be numbered or unnumbered (``1 Introduction'' and ``1.2 Numbered level 2 head''), with no ending punctuation. As demonstrated in this document, the initial paragraph after a headingis not indented.

\section{JOURNAL STYLE}

Use American English when writing your paper. The serial comma should be used (``a, b, and c'' not ``a, b and c''). In American English, periods and commas are within quotation marks, like ``this period.'' Other punctuation is ``outside''! The use of technical jargon, slang, and vague or informal English should be avoided. Generic technical terms should instead be used.

\subsection{Acronyms and abbreviations}

All acronyms should be defined at first mention in the abstract and in the main text. Define in figures, tables, and footnotes only if not defined in the discussion of the figure/table. Acronyms consist of capital letters (except where salted with lowercase), but the terms they represent need not be given initial caps unless a proper name is involved (``central processing unit'' [CPU] but ``Fourier transform'' [FT]). Use of ``e.g.'' and ``i.e.'' okay, but refrain from using ``etc.'' It is preferable to use these abbreviations only in parentheses (e.g., like this).

Abbreviate units of time (s, min, hr, day, mo, yr) only in virgule constructions (10~$\umu$g/hr) and in artwork; otherwise, spell out, e.g., 10 days, 3 months, 25 minutes. Units of measure (Kb, MB, kWh, etc.) should always be abbreviated when used with a numeral. If used alone, spell out (``16 MB of RAM'' but ``these values are measured in micrometers'').

\subsection{Numbers}

Spell out numerals that have no unit of measure or time (one, two, $\ldots$ ten), but always use numerals with units of time and measure. Some examples are as follows: 11 through 999; 1,000; 10,000; twentieth century; twofold, tenfold, 20-fold; 2 times; 0.2 cm; $p = .001$; 25\%; 10\% to 25\%.

\section{MATH AND EQUATIONS}

Scalar {\it variables} and {\it physical constants} should be italicized, and a bold (non-italics) font should be used for {\bf vectors} and {\bf matrices}. Do not italicize subscripts unless they are variables.

Equations should be either display (with a number in parentheses) or inline.

Display equations should be flush left and numbered consecutively, with equation numbers in parentheses and flush right.

Be sure the symbols in your equation have been defined before the equation appears or immediately following. Please refer to ``Equation (1),'' not ``Eq. (1)'' or ``equation (1).''

Punctuate display equations when they are part of the sentence preceding it, as in
\begin{equation}
A=\pi r^2.
\end{equation}
In addition, if the text following the equation flows logically as a part of the display equation, 
\begin{equation}
A=\pi r^2,
\end{equation}
use ending punctuation (comma) after the display equation.

\section{LISTS}

Avoid using lists. Instead, use full sentences and flowing paragraphs. If you absolutely must use a list, use them rarely and keep them short:
\begin{itemize}
\item {\it Style for bulleted lists}---This is the style that should be used for bulleted lists.

\item {\it Punctuation in lists}---Each item in the list should end with a period, regardless of whether full sentences are used.
\end{itemize}

\section{GRAPHICAL ABSTRACTS}

This journal accepts graphical abstracts, and they must be peer reviewed, which means the graphical abstract must be submitted with the full paper. graphical abstracts and their specifications. Please read the additional information provided by \href{http://www.ieee.org/publications_standards/publications/graphical_abstract.pdf}{\underline{IEEE about graphical abstracts}}.

%\begin{figure}
%\centerline{\includegraphics[width=18.5pc]{fig1.png}}
%\caption{Note that ``Figure'' is spelled out. There is a period after the figure number, followed by one space. It is good practice to briefly explain the significance of the figure in the caption. (Used, with permission, from [4].)}
%\end{figure}
%
%\begin{figure*}
%\centerline{\includegraphics[width=26pc]{fig1.png}}
%\caption{Note that ``Figure'' is spelled out. There is a period after the figure number, followed by one space. It is good practice to briefly explain the significance of the figure in the caption. (Used, with permission, from [4].)}
%\end{figure*}

\begin{table}
\caption{Units for magnetic properties.}
\label{table}
\small
\begin{tabular*}{17.5pc}{@{}|p{29pt}|p{63pt}<{\raggedright}|p{80pt}<{\raggedright}|@{}}
\hline
Symbol& 
Quantity& 
Conversion from Gaussian and  CGS EMU to SI$^{\mathrm{a}}$ \\
\hline
$\Phi $& 
Magnetic flux& 
1 Mx $\to  10^{-8}$ Wb $= 10^{-8}$ V $\cdot$ s \\
$B$& 
Magnetic flux density,   magnetic induction& 
1 G $\to  10^{-4}$ T $= 10^{-4}$ Wb/m$^{2}$ \\
$H$& 
Magnetic field strength& 
1 Oe $\to  10^{-3}/(4\pi )$ A/m \\
$m$& 
Magnetic moment& 
1 erg/G $=$ 1 emu   $\to 10^{-3}$ A $\cdot$ m$^{2} = 10^{-3}$ J/T \\
$M$& 
Magnetization& 
1 erg/(G $\cdot$ cm$^{3}) =$ 1 emu/cm$^{3}$   $\to 10^{-3}$ A/m \\
4$\pi M$& 
Magnetization& 
1 G $\to  10^{-3}/(4\pi )$ A/m \\
$\sigma $& 
Specific magnetization& 
1 erg/(G $\cdot$ g) $=$ 1 emu/g $\to $ 1 A $\cdot$ m$^{2}$/kg \\
$j$& 
Magnetic dipole   moment& 
1 erg/G $=$ 1 emu   $\to 4\pi \times  10^{-10}$ Wb $\cdot$ m \\
$J$& 
Magnetic polarization& 
1 erg/(G $\cdot$ cm$^{3}) =$ 1 emu/cm$^{3}$  $\to 4\pi \times  10^{-4}$ T \\
$\chi , \kappa $& 
Susceptibility& 
1 $\to  4\pi $ \\
$\chi_{\rho }$& 
Mass susceptibility& 
1 cm$^{3}$/g $\to  4\pi \times  10^{-3}$ m$^{3}$/kg \\
$\mu $& 
Permeability& 
1 $\to  4\pi \times  10^{-7}$ H/m   $= 4\pi \times  10^{-7}$ Wb/(A $\cdot$ m) \\
$\mu_{r}$& 
Relative permeability& 
$\mu \to \mu_{r}$ \\
$w, W$& 
Energy density& 
1 erg/cm$^{3} \to  10^{-1}$ J/m$^{3}$ \\
$N, D$& 
Demagnetizing factor& 
1 $\to  1/(4\pi )$ \\
\hline
\multicolumn{3}{@{}p{17.5pc}@{}}{Vertical lines are optional in tables. Statements that serve as captions for 
the entire table do not need footnote letters. }\\
\multicolumn{3}{@{}p{17.5pc}@{}}{$^{\mathrm{a}}$Gaussian units are the same as cg emu for magnetostatics; Mx 
$=$ maxwell, G $=$ gauss, Oe $=$ oersted; Wb $=$ weber, V $=$ volt, s $=$ 
second, T $=$ tesla, m $=$ meter, A $=$ ampere, J $=$ joule, kg $=$ 
kilogram, H $=$ henry.}
\end{tabular*}
\label{tab1}
\end{table}

\section{FIGURES AND TABLES}

\subsection{In-text callouts for figures and tables}

Figures and tables must be cited in the running text in consecutive order. At first mention, the citation should be boldface ({\bf Figure 1}); subsequent mentions should be Roman type (see Figure 1 and {\bf Table 1}). {\bf Figure 2} shows an example of a figure spanning across two columns.

Previously published figures or tables require permission to reprint. Please obtain permission. Then, add the following text to the figure/table caption: ``From [reference no.], with permission,'' or ``Adapted from [reference no.], with permission.'' {\it Carefully} explain each figure in the text. Each manuscript should be limited to four figures.\pagebreak

\section{END SECTIONS}

\subsection{Appendices}

If multiple appendices are required, they should labeled ``Appendix A,'' ``Appendix B,'' etc. They appear before the ``Acknowledgment'' or the ``References'' section.

\subsection{Acknowledgment}

The ``Acknowledgment'' (no's) section appears immediately after the conclusion. If applicable, this is where you indicate funding for the work. The preferred spelling of the word ``acknowledgment'' in American English is without an ``e'' after the ``g.'' Avoid expressions such as ``One of us (S.B.A.) would like to thank $\ldots$ .'' Instead, write ``We thank $\ldots$.'' Sponsor and financial support acknowledgments are included in the acknowledgment section. For example: This work was supported in part by the U.S. Department of Commerce under Grant BS123456 (sponsor and financial support acknowledgment goes here). Researchers that contributed information or assistance to the article should also be acknowledged in this section. Also, if corresponding authorship is noted in your paper it will be placed in the acknowledgment section. Note that the acknowledgment section is placed at the end of the paper before the reference section.

\subsection{References}

References need not be cited in text. When they are, they appear on the 
line, in square brackets, inside the punctuation. Multiple references are 
each numbered with separate brackets. When citing a section in a book, 
please give the relevant page numbers. In text, refer simply to the 
reference number. Do not use ``Ref.'' or ``reference'' except at the 
beginning of a sentence: ``Reference \cite{CC1} shows $\ldots$ .'' Please do not use 
automatic endnotes in \emph{Word}, rather, type the reference list at the end of the 
paper using the ``References'' style.

Reference numbers are set flush left and form a column of their own, hanging 
out beyond the body of the reference. The reference numbers are on the line, 
end with period. In all references, the given name of the author 
or editor is abbreviated to the initial only and precedes the last name. Use 
them all; use \emph{et al.} only if names are not given. Use commas around Jr., 
Sr., and III in names. Abbreviate conference titles. When citing IEEE 
transactions, provide the issue number, page range, volume number, year, 
and/or month if available. When referencing a patent, provide the day and 
the month of issue, or application. References may not include all 
information; please obtain and include relevant information. Do not combine 
references. There must be only one reference with each number. If there is a 
URL included with the print reference, it can be included at the end of the 
reference.

Other than books, capitalize only the first word in a paper title, except 
for proper nouns and element symbols. For papers published in translation 
journals, please give the English citation first, followed by the original 
foreign-language citation See the end of this document for formats and 
examples of common\break references. For a complete discussion of references and 
their formats, see the IEEE style manual at
\url{http://www.ieee.org/authortools}.

\section{CONCLUSION}

The manuscript should include a conclusion. In this section, summarize what was described in your paper. Future directions may also be included in this section. Authors are strongly encouraged not to reference multiple figures or tables in the conclusion; these should be referenced in the body of the paper.

\section{ACKNOWLEDGMENT}

We thank A, B, and C. This work was supported in part by a grant from XYZ.

\begin{thebibliography}{1}

\bibitem{AA1}
G. M. Amdahl, G. A. Blaauw, and F. P. Brooks, ``Architecture of the IBM System/360,'' {\it IBM J. Res. \& Dev}., vol. 8, no. 2, pp. 87--101, 1964. (journal)

\bibitem{BB1}
IBM Corporation, IBM Knowledge Center - IBM Secure Service Container (Secure Service Container). [Online]. Available: {https://www.ibm.com/support/\break knowledgecenter/en/HW11R/com.ibm.hwmca.kc\_se.doc/\break introductiontotheconsole/wn2131zaci.html} (URL)

\bibitem{CC1}
J. Williams, ``Narrow-band analyzer,'' PhD dissertation, Dept. of Electrical Eng., Harvard Univ., Cambridge, MA: 1993. (Thesis or dissertation)

\bibitem{DD1}
J. M. P. Martinez, R. B. Llavori, M. J. A. Cabo, et al., ``Integrating data warehouses with web data: A survey,'' {\it IEEE Trans. Knowledge and Data Eng}., preprint, Dec. 21, 2007, doi:10.1109/TKDE.2007.190746. (PrePrint)

\bibitem{EE1}
W.-K. Chen, {\it Linear Networks and Systems}, Belmont, CA: Wadsworth, pp. 123--135, 1993. (book)

\bibitem{FF1}
S. P. Bingulac, ``On the compatibility of adaptive controllers,'' {\it Proc. Fourth Ann. Allerton Conf. Circuits and Systems Theory}, pp. 8--16, 1994. (conference proceedings)\vadjust{\vfill\pagebreak}

\bibitem{GG1}
K. Elissa, ``An overview of decision theory,'' unpublished. (Unpublished manuscript)

\bibitem{HH1}
R. Nicole, ``The last word on decision theory,'' {\it J. Computer Vision}, submitted for publication. (Pending publication)

\bibitem{II1}
C. J. Smith and J. S. Smith, Rocky Mountain Research Laboratories, Boulder, CO, personal communication, 1992. (Personal communication)
\end{thebibliography}

\begin{IEEEbiography}{First A. Author}{\,}current employment (is
currently with XYZ Corporation, New York, NY, USA). Author's latest degree received or which he/she is currently pursuing (He received the B.S. degree and the M.S. degree...."). Author's research interest. Association with any official journals or conferences; major professional and/or academic achievements, i.e., best paper awards, research grants, etc.; any publication information (number of papers and titles of books published). Author membership information, e.g., is a member of the IEEE and the IEEE Computer Society, if applicable, is noted at the end of the biography. Biography is limited to single paragraph only (He is a member of IEEE, etc.). All biographies should be limited to one paragraph, five to six sentences, consisting of the following: (e.g., ``Dr. Author received a B.S. degree and an M.S. degree $\ldots$.''), including years achieved; association with any official journals or conferences; major professional and/or academic achievements, i.e., best paper awards, research grants, etc.; any publication information (number of papers and titles of books published); current research interests; association with any professional associations. Author membership information, e.g., is a member of the IEEE and the IEEE Computer Society, if applicable, is noted at the end of the biography. Contact him/her at faauthor@xyz.com.
\end{IEEEbiography}

\begin{IEEEbiography}{Second B. Author, Jr.,}{\,}is with ABC Corporation, Bblingen, Germany. Ms. Author's biography appears here.  Contact him/her at sbauthor@abc.com.
\end{IEEEbiography}

\begin{IEEEbiography}{Third C. Author, III,}{\,}is with DEF Corporation, Tokyo, Japan. Dr. Author's biography appears here.  Contact him/her at tcauthor@def.com.
\end{IEEEbiography}

### 2021-03-29 bem.tex
Finished most of the OpenCL section.

@@
 	\includegraphics[width=5cm]{img/sphere}
+	\caption{Discretisation of a sphere into flat surface triangles.}
 		\label{fig:triangulation}
-	\caption{Discretisation of a sphere into flat surface triangles.}
+
 \end{figure}
@@
 \begin{enumerate}
-	\item Import of the surface description as triangulation data
-	\item Definition of the trial space, test space and range space
-	\item Discretization into a matrix problem $Ax=b$
-	\item Evaluation of domain potential operators for visualization and post-processing
+	\item Import of the surface description as triangulation data.
+	\item Definition of the trial space, test space and range space.
+	\item Discretization into a matrix problem $Ax=b$.
+	\item Solution of the matrix problem by either a direct or iterative solver.
+	\item Evaluation of domain potential operators for visualization and post-processing.
 \end{enumerate}

### 2021-03-29 bempp_overview.tex
Finished most of the OpenCL section.

\section{A high level overview of Bempp-cl}

\begin{figure}
	\center
	\includegraphics[width=7cm]{img/bempp_overview}
	\caption{The layout of Bempp-cl with its computational backends.}
	\label{fig:overview}
\end{figure}

The main user visible component of Bempp-cl is the module \textbf{bempp.api}, which defines all user interface functions and other high-level routines. In particular, it contains the definition of the main object types: \textbf{Grids}, \textbf{Spaces}, \textbf{Gridfunctions}, \textbf{Operators}. All computational backend routines are part of \textbf{bempp.core}. We are currently supporting a full Numba backend and an OpenCL backend. An overview of this structure is provided in Figure \ref{fig:overview}.

Outside the operator discretisation Numba is used in the following contexts.
\begin{itemize}
	\item Topological computations for the grid. Finding out all neighbour relationships between triangles requires parses through the grid data structure with linear complexity, which is accelerated with Numba.
	\item Definition of local-to-global maps for function spaces. Again, this requires traversal through the grid and assigning relationships between global and local indices.
	\item Grid Functions. A right-hand side function $f$ can be defined as a Python callable. This is just-in-time compiled via Numba and then the product with the corresponding basis functions integrated in each triangle via numerical quadrature, again via Numba accelerated routines.
	\item Sparse matrices. Sparse surface matrices are assembled through Numba accelerated routines.
\end{itemize}

The main computational complexity in discretising the boundary integral equation \eqref{eq:bnd_integral} is coming from the discretisation of the left-hand side integral operator. Using dense methods this has quadratic complexity in the number of surface triangles.

Once the user defines an operator with its associated discretisation spaces and calls the \textbf{weak\_form} method to discretize it the code calls a regular integrator to assemble all the interactions between non-adjacent elements and a singular integrator to compute the interactions between adjacent triangles (only necessary if trial and test space are defined on the same grid). The result is stored internally as a dense Numpy array. The corresponding discretisation routines are proxies that forward to computational routines in either Numba or OpenCL, depending on the user preferences. For OpenCL assembly the code checks additional parameters, such as the default vector length for SIMD operators (e.g. 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), and whether the discretisation should proceed in single or double precision. The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed. If the computational backend is Numba, the call is forwarded to the corresponding Numba discretisation routines and executed. For simple piecewise constant function spaces or other spaces, where the support of each basis function is localized to a single triangle, only one call to the computational routines is necessary. If the support of basis functions is larger than a single triangle different threads may need to sum into the same global degree of freedom. We discuss this in more detail in Section \ref{sec:coloring}.

### 2021-03-29 opencl.tex
Finished most of the OpenCL section.

\section{Assembling boundary integral operators with OpenCL}

In this section we discuss in more detail the assembly of boundary integral operators with OpenCL
and how we integrated this into our Python workflow. We start with a brief introduction to OpenCL and then
dive into how we use OpenCL as part of Bempp-cl.

\subsection{What is OpenCL?}

OpenCL (\url{https://www.khronos.org/opencl/}) is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers. At its core OpenCL executes compute kernels that can be written in C99, or more recently, in C++. The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses. OpenCL splits the computational tasks into work-items, each of which represents a single execution of a compute kernel. Work items are grouped together into work-groups, which share local memory. Barrier synchronization is only allowed within a work-group. All work-items are uniquely indexed by a one, two, or three dimensional index space, called NDRange. Kernels are launched onto a compute device (e.g. CPU or GPU) from a host device. OpenCL allows kernels to be loaded as strings and compiled on the fly for a given device, making it well suited for launching from high-productivity languages. To launch an OpenCL kernel the following steps need to be performed.
\begin{itemize}
	\item Initialize buffers on the compute device.
	\item Compute relevant data from host to the compute device.
	\item Load a kernel string and just-in-time compile it for the device.
	\item Execute the kernel for the given buffers and chosen index space.
	\item Copy results back to the host.
\end{itemize}
An excellent feature of OpenCL is its very good support for vectorized operations. OpenCL provides vector data types, such as \textbf{double4}, that can hold four double values in a SIMD register and defines a number of standard operations for these vector types. This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.

Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner, which automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.

\subsection{OpenCL Assembly in Bempp-cl}

\begin{figure*}
	\center
	\includegraphics[width=12cm]{img/kernel_header}
	\caption{Definition of the OpenCL compute kernel for scalar integral equations.}
	\label{fig:kernel_definition}
\end{figure*}

Bempp-cl has OpenCL kernels for all its boundary operators. All operators have the same interface and are launched in the same way. In the first step the relevant data will need to be copied to the compute device. This data consists of:

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grid as flat floating point array, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity information that stores the indices of each node for each triangle.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle.
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item Number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}
The actual kernel definition is shown in Figure \ref{fig:kernel_definition}.
Before the kernel can be launched, it needs to be configured and just-in-time compiled. Kernel configuration happens through C-style preprocessor definitions that are passed through the just-in-time compiler. These include the names of the test and trial space, the name of the function that evaluates the integral kernel, choosing between single and double precision types, and for SIMD enhanced kernel the vector length of the SIMD types.
For example, in Figure \ref{fig:kernel_definition} all floating point types have the name REALTYPE. This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.

Each work-item computes via numerical quadrature all interactions via the integral kernel of basis functions on the trial element with basis functions on the test element. Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangle are adjoint or identical. For this it simply checks if at at least one of the node indices of the test triangle is identical to one of the node indices of the trial triangle. If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer. The reason is that for adjoint triangles a separate singular quadrature rule is used. The effect is that a few work-items do work that is discarded. However, in a grid with $N$ elements the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of interactions is $N^2$. Hence, only a tiny fraction of work-items are discarded.

\subsection{SIMD optimized kernels}

The above description was valid for kernels that do not take advantage of CPU SIMD operations are kernels that run directly on a GPU device. For SIMD optimization on CPUs we proceed slightly differently. The corresponding kernel in principle works as described above, but batches together the interaction on one test triangle with X trial triangles, where X is either 4, 8, or 16, depending on the number of available SIMD lanes. This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation. If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorized kernel.

Each kernel definition is stored in two variants, one with the ending \textit{\_novec.cl} and another one with the ending \textit{\_vec.cl}. The vectorized variant is configured via preprocessor directives for the desired number of vector lanes. Having to develop two OpenCL kernel codes for each operator is a certain amount of overhead. If we write a new kernel, we first implement the non-vectorized version. Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.

Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorize kernels by batching together work items on SIMD lanes, similarly to what we do manually. In our experience this works well for very simple kernels but often fails for more complex OpenCL kernels. This is why we decided to implement this strategy manually.

A completely different SIMD strategy is based on batching together quadrature evaluations within a single test/triangle pair. There are two disadvantages to this approach. First, it only works well if the number of quadrature points is a multiple of the available SIMD lanes. Second, other operations such as the geometry calculations for each element then can not be SIMD optimized as these are only performed once per test/triangle pair.

### 2021-04-06 bem.tex
edititing first section

@@
 \section{Boundary element methods with Bempp}
-In this section we want to provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretisation and solution.
+In this section, we provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretisation and solution.
 
 The most simple boundary integral equation is of the form
-\begin{equation}
+\begin{align}
   \label{eq:bnd_integral}
-	\int_{\Gamma} g(\bx, \by)\phi(\by)ds(\by) = f(\bx),~\bx\in\Gamma.
-\end{equation}
-The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$.
-To provide a concrete example.
-To compute the electrostatic capacity of an object $\Omega$ one solves the above equation with $f(\bx)$ the constant function $1$ and $g(\bx,\by)=\frac{1}{4\pi|\bx - \by|}$.
-The normalized capacity is then obtained as $c = \frac{1}{4\pi}\int_{\Gamma}\phi(\bx)ds(\bx)$.
+  \int_{\Gamma} g(\bx, \by)\phi(\by)\,\mathrm{d}s(\by) &= f(\bx),&&\bx\in\Gamma.
+\end{align}
+The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side, and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$.
+
+As a concrete example, we consider computing the electrostatic capacity of an object $\Omega$.
+In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$.
+Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\,\mathrm{d}s(\bx)$.
+
 Many practical problems have a significantly more complex structure and can involve block systems of integral equations.
@@
   \label{fig:triangulation}
-
 \end{figure}
 
-The first step is the \textbf{definition of the surface} $\Gamma$.
-Surfaces are represented in Bempp-cl as a triangulation into flat triangles (see \ref{fig:triangulation}).
+The first step is the \textbf{discretisation of the surface} $\Gamma$.
+Surfaces are represented in Bempp-cl as a triangulation into flat triangles (see \cref{fig:triangulation}).
 The triangulation is internally represented as an array of node coordinates and an associated connectivity array of node indices that define each triangle.
-In this step also topology data is computed.
-In particular, for each triangle the neighboring triangles and the type of intersection (i.e.
-via joint node or edge) is computed.
+In this step, topology data is also computed: in particular, for each triangle, we compute the neighboring triangles and the type of intersection (i.e.
+are they connected by an edge or a vertex).
 
 Once a triangulation is given we need to define the necessary data structures for the discretisation.
-Bempp-cl uses a Galerkin discretisation.
-Consider \eqref{eq:bnd_integral}.
-We represent the solution $\phi$ as $\phi=\sum_{j=}^N x_j\phi_j$, where the $\phi_i$ are basis functions of a \textbf{trial space}, defined in the most simple case as constant $1$ on the triangle $\tau_j$ and $0$ everywhere also.
-Several other types of discretisation spaces are available in Bempp.
-Moreover, we require a \textbf{test (or dual) space} for the discretisation.
-Let $\psi=\sum_{i=1}^Ny_i \psi_i$, where each $\psi_i$ is a basis function of a space of test functions.
-For simplicity, here we will assume that the $\psi_i$ are also piecewise constant.
-The discrete representation of the above problem then takes the form
+Bempp-cl uses a Galerkin discretisation: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function.
+We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N x_j\phi_j$.
+In the most simple case, we can define the function \(\phi_j\) to be equal to 1 on the triangle $\tau_j$ and 0 everywhere else.
+Other spaces are commonly defined to be piecewise polynomials on each triangle.
 
+To discretise \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
 $$
@@
 with
-$$
-A_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)ds(\by)ds(\bx)
-$$
-and $f_i = \int_{\Gamma}\psi_i(\bx)f(\bx)ds(\bx)$.
-In the case of piecewise constant trial and test functions the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)ds(\by)ds(\bx)$.
+\begin{align*}
+A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)\\
+b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(\bx).
+\end{align*}
+In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)$.
 
-In Bempp-cl an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
+In Bempp-cl, an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
 the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
-products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{Grid function object}, which
-consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)ds(x)$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
+products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
+consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(x)$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
 
-Once the grid, the space objects, and the operator are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $A_{ij}$.
+Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $A_{ij}$.
 For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $A_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights.
@@
 
-In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as Gmres.
-The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.
+In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRes.
+The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualisation.
 
-In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp
+In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp:
 \begin{enumerate}

### 2021-04-06 intro.tex
edititing first section

@@
-\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
-The original release came with a simple Python wrapper to the C++ interface.
-Over time more and more functionality was moved into the Python interface.
-However, computationally intensive routines and the main data structures remained in C++.
+\chapterinitial{The Bempp boundary element library} (originally named BEM++ \cite{BEM++}) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
+The original release came with a simple Python wrapper to the C++ library.
+Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
+
 This proved a burden when efforts started to modernise the library to be able to make better use of Single-Instruction-Multiple-Data (SIMD) optimisation on CPUs and offload computation to GPUs.
-In 2018 we decided to rewrite the computational core from scratch.
+In 2018, we decided to rewrite the computational core from scratch.
 The aims were to support explicit SIMD optimisation on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
-These aims naturally lead to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.
-The name of the new library was Bempp-cl.
-At the end of 2019 we released an initial version 0.1 that we used to test the library in a number of real-world scenarios.
+These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.
+
+At the end of 2019, we released the first version (0.1) of Bempp-cl \cite{Bempp-cl}.
 This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use.
-Since then we have used Bempp-cl in a number of practical application and many of our users are slowly migrating to it from the old C++ based Bempp.
-In this article we want to discuss the design choices behind Bempp-cl and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.
+Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp.
+In this article, we want to discuss the design choices behind Bempp-cl and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.
 

### 2021-04-06 bempp_overview.tex
Make tikz version of figure

@@
 \begin{figure}
-	\center
-	\includegraphics[width=7cm]{img/bempp_overview}
+	\centering
+    \input{img/bempp_overview}
 	\caption{The layout of Bempp-cl with its computational backends.}
@@
 
-The main user visible component of Bempp-cl is the module \textbf{bempp.api}, which defines all user interface functions and other high-level routines.
-In particular, it contains the definition of the main object types: \textbf{Grids}, \textbf{Spaces}, \textbf{Gridfunctions}, \textbf{Operators}.
-All computational backend routines are part of \textbf{bempp.core}.
+The main user-visible component of Bempp-cl is the module \texttt{bempp.api}, which defines all user interface functions and other high-level routines.
+In particular, it contains the definition of the main object types: \texttt{Grid}, \texttt{Space}, \texttt{GridFunction}, \texttt{Operator}.
+All computational backend routines are part of \texttt{bempp.core}.
 We are currently supporting a full Numba backend and an OpenCL backend.
-An overview of this structure is provided in Figure \ref{fig:overview}.
+An overview of this structure is provided in \cref{fig:overview}.
 
-Outside the operator discretisation Numba is used in the following contexts.
+Outside the operator discretisation, Numba is used in the following contexts:
 \begin{itemize}
@@
 
-The main computational complexity in discretising the boundary integral equation \eqref{eq:bnd_integral} is coming from the discretisation of the left-hand side integral operator.
+The main computational complexity in discretising the boundary integral equation \cref{eq:bnd_integral} is coming from the discretisation of the left-hand side integral operator.
 Using dense methods this has quadratic complexity in the number of surface triangles.
@@
 If the support of basis functions is larger than a single triangle different threads may need to sum into the same global degree of freedom.
-We discuss this in more detail in Section \ref{sec:coloring}.
+We discuss this in more detail in \cref{sec:coloring}.

### 2021-04-06 img/bempp_overview.tex
Make tikz version of figure

\begin{tikzpicture}[x=1.45cm,y=-1.45cm]

\draw[line width=1pt] (0.5,0.5) -- (0.5,0.7);
\draw[line width=1pt] (1.7,0.5) -- (1.7,0.7);
\draw[line width=1pt] (2.9,0.5) -- (2.9,0.7);
\draw[line width=1pt] (4.1,0.5) -- (4.1,0.7);
\draw[line width=1pt] (4.1,1.2) -- (4.1,1.4);

\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (0,0) rectangle (4.6,0.5);
\node at (2.3,0.25) {\texttt{bempp.api}};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (0,0.7) rectangle (1,1.2);
\node at (0.5,0.95) {\footnotesize grids};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (1.2,0.7) rectangle (2.2,1.2);
\node at (1.7,0.95) {\footnotesize spaces};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (2.4,0.7) rectangle (3.4,1.2);
\node at (2.9,0.95) [text width=1.45cm,align=center] {\footnotesize grid\\[-1.2mm]functions};
\draw[line width=1pt,line join=round,line cap=round,bempporange] (3.6,0.7) rectangle (4.6,1.2);
\node at (4.1,0.95) {\footnotesize operators};
\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.5,1.4) rectangle (4.7,1.9);
\node at (4.1,1.65) {\footnotesize\texttt{bempp.core}};

\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (0,1.3) -- (3.4,1.3);
\node[anchor=north] at (1.7,1.35) {\footnotesize Numba};

\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (3.5,2) -- (4.7,2);
\node[anchor=north] at (4.1,2.05) {\footnotesize Numba / OpenCL};
\end{tikzpicture}

### 2021-04-06 bempp_overview.tex
make ise, isation, etc consistent

@@
 
-Once the user defines an operator with its associated discretisation spaces and calls the \textbf{weak\_form} method to discretize it the code calls a regular integrator to assemble all the interactions between non-adjacent elements and a singular integrator to compute the interactions between adjacent triangles (only necessary if trial and test space are defined on the same grid).
+Once the user defines an operator with its associated discretisation spaces and calls the \textbf{weak\_form} method to discretise it the code calls a regular integrator to assemble all the interactions between non-adjacent elements and a singular integrator to compute the interactions between adjacent triangles (only necessary if trial and test space are defined on the same grid).
 The result is stored internally as a dense Numpy array.
@@
 If the computational backend is Numba, the call is forwarded to the corresponding Numba discretisation routines and executed.
-For simple piecewise constant function spaces or other spaces, where the support of each basis function is localized to a single triangle, only one call to the computational routines is necessary.
+For simple piecewise constant function spaces or other spaces, where the support of each basis function is localised to a single triangle, only one call to the computational routines is necessary.
 If the support of basis functions is larger than a single triangle different threads may need to sum into the same global degree of freedom.

### 2021-04-06 opencl.tex
make ise, isation, etc consistent

@@
 Work items are grouped together into work-groups, which share local memory.
-Barrier synchronization is only allowed within a work-group.
+Barrier synchronisation is only allowed within a work-group.
 All work-items are uniquely indexed by a one, two, or three dimensional index space, called NDRange.
@@
 \begin{itemize}
-	\item Initialize buffers on the compute device.
+	\item Initialise buffers on the compute device.
 	\item Compute relevant data from host to the compute device.
@@
 \end{itemize}
-An excellent feature of OpenCL is its very good support for vectorized operations.
+An excellent feature of OpenCL is its very good support for vectorised operations.
 OpenCL provides vector data types, such as \textbf{double4}, that can hold four double values in a SIMD register and defines a number of standard operations for these vector types.
@@
 
-Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner, which automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.
+Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner, which automates much of the initialisation of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.
 
@@
 
-\subsection{SIMD optimized kernels}
+\subsection{SIMD optimised kernels}
 
 The above description was valid for kernels that do not take advantage of CPU SIMD operations are kernels that run directly on a GPU device.
-For SIMD optimization on CPUs we proceed slightly differently.
+For SIMD optimisation on CPUs we proceed slightly differently.
 The corresponding kernel in principle works as described above, but batches together the interaction on one test triangle with X trial triangles, where X is either 4, 8, or 16, depending on the number of available SIMD lanes.
-This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
-If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorized kernel.
+This strategy allows us to optimise almost all floating point operations within a kernel run for SIMD operation.
+If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorised kernel.
 
 Each kernel definition is stored in two variants, one with the ending \textit{\_novec.cl} and another one with the ending \textit{\_vec.cl}.
-The vectorized variant is configured via preprocessor directives for the desired number of vector lanes.
+The vectorised variant is configured via preprocessor directives for the desired number of vector lanes.
 Having to develop two OpenCL kernel codes for each operator is a certain amount of overhead.
-If we write a new kernel, we first implement the non-vectorized version.
-Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.
+If we write a new kernel, we first implement the non-vectorised version.
+Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorised or not, it is usually a matter of an hour or two to convert the non-vectorised kernel into a vectorised version.
 
-Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorize kernels by batching together work items on SIMD lanes, similarly to what we do manually.
+Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorise kernels by batching together work items on SIMD lanes, similarly to what we do manually.
 In our experience this works well for very simple kernels but often fails for more complex OpenCL kernels.
@@
 First, it only works well if the number of quadrature points is a multiple of the available SIMD lanes.
-Second, other operations such as the geometry calculations for each element then can not be SIMD optimized as these are only performed once per test/triangle pair.
+Second, other operations such as the geometry calculations for each element then can not be SIMD optimised as these are only performed once per test/triangle pair.

### 2021-04-06 img/sphere.py
add tikz sphere

@@ -1,0 +1,18 @@
+import bempp.api
+
+g = bempp.api.shapes.sphere(h=0.3)
+
+
+with open("sphere.tex", "w") as f:
+
+    f.write("\\begin{tikzpicture}[x={(1.732cm,-1cm)},y={(1.732cm,1cm)},z={(0,2cm)}]\n")
+
+    for i, j in enumerate(g.vertices.T):
+        f.write(f"\\coordinate (v{i}) at ({j[0]},{j[1]},{j[2]});\n")
+
+    triangles = [(sum(g.vertices.T[i] for i in t) / 3, t) for t in g.elements.T]
+
+    triangles.sort(key=lambda x: x[0][0] + x[0][2] - x[0][1])
+    for _, (i, j, k) in triangles:
+        f.write(f"\\draw[fill=white, fill opacity=0.8] (v{i}) -- (v{j}) -- (v{k}) -- cycle;\n")
+    f.write("\\end{tikzpicture}")

### 2021-04-06 opencl.tex
Further work

--- 2021-03-29 opencl.tex
@@
 Second, other operations such as the geometry calculations for each element then can not be SIMD optimized as these are only performed once per test/triangle pair.
+
+\subsection{Assembling the singular part of integral operators}
+
+The assembly of the singular part of an integral operator works a bit differently.
+Remember that the singular part consists of triangle parts which are adjacent to each other or identical.
+There are $\mathcal{O}(N)$ of such pairs.
+We are using fully numerical quadrature rules for these integrals that are based on subdividing the four-dimensional integration domain and using transformation techniques to remove the singularities.
+This gives highly accurate evaluations of these integration pairs but requires a large number (typically over $1000$) quadrature points per triangle pair.
+Hence, we create one workgroup for each triangle pair and inside this workgroup have a number of work-items that are evaluating the quadrature rules and then at the end sum up the result together.
+Depending on how two triangles are in relationship to each other different types of singular quadrature rules are needed.
+We solve this pre-loading all possible quadrature rules onto the device and also store for each triangle pair an index into the required quadrature rule so that the kernel function can select the correct rule to evaluate.
+For the singular quadrature rules we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above.
+At the end, the singular integral values are either summed into the overall result matrix, or if desired by the user, stored as separate sparse matrix.
+
+\section{Numba assembly of integral operators}
+
+The Numba assembly of integral operators is structured differently from the OpenCL assembly.
+In OpenCL we create a large number of work-items that are executed in parallel on the compute device by the runtime system with the details of the parallel execution left to the device driver.
+For Numba we use classical loop based parallelism.
+Each loop iteration is the assembly of one test element with all trial elements.
+Within each loop we partially optimized for SIMD instruction, though this depends on the autovectorization of the just-in-time compiler and cannot be explicitly targeted in Numba.
+Again, we separate regular and singular interactions.
+In the next section we give some more details on SIMD optimization in Numba at the example of the Laplace kernel.
+
+\section{SIMD optimized evaluation of the Laplace kernel with Numba}

### 2021-04-06 bem.tex
make ise, isation, etc consistent
add tikz sphere
add command for \ds

@@
   \label{eq:bnd_integral}
-  \int_{\Gamma} g(\bx, \by)\phi(\by)\,\mathrm{d}s(\by) &= f(\bx),&&\bx\in\Gamma.
+  \int_{\Gamma} g(\bx, \by)\phi(\by)\ds[\by] &= f(\bx),&&\bx\in\Gamma.
 \end{align}
@@
 In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$.
-Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\,\mathrm{d}s(\bx)$.
+Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\ds[\bx]$.
 
@@
 \begin{figure}
-  \center
-  \includegraphics[width=5cm]{img/sphere}
+  \centering
+  \input{img/sphere}
   \caption{Discretisation of a sphere into flat surface triangles.}
@@
 \begin{align*}
-A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)\\
-b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(\bx).
+A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\ds[\by]\ds[\bx]\\
+b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx].
 \end{align*}
-In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\,\mathrm{d}s(\by)\,\mathrm{d}s(\bx)$.
+In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\ds[\by]\ds[\bx]$.
 
@@
 products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
-consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\,\mathrm{d}s(x)$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
+consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
 
@@
 	\item Definition of the trial space, test space and range space.
-	\item Discretization into a matrix problem $Ax=b$.
+	\item Discretisation into a matrix problem $Ax=b$.
 	\item Solution of the matrix problem by either a direct or iterative solver.
-	\item Evaluation of domain potential operators for visualization and post-processing.
+	\item Evaluation of domain potential operators for visualisation and post-processing.
 \end{enumerate}

### 2021-04-06 img/bempp_overview.tex
add command for \ds

@@
 \node at (4.1,0.95) {\footnotesize operators};
-\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.5,1.4) rectangle (4.7,1.9);
+\draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.45,1.4) rectangle (4.75,1.9);
 \node at (4.1,1.65) {\footnotesize\texttt{bempp.core}};
@@
 
-\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (3.5,2) -- (4.7,2);
+\draw[line width=1pt,decorate,decoration={brace,amplitude=3pt,mirror}] (3.45,2) -- (4.75,2);
 \node[anchor=north] at (4.1,2.05) {\footnotesize Numba / OpenCL};

### 2021-04-06 img/bempp_overview.tex
working on bempp section

@@
 
-\draw[line width=1pt] (0.5,0.5) -- (0.5,0.7);
-\draw[line width=1pt] (1.7,0.5) -- (1.7,0.7);
-\draw[line width=1pt] (2.9,0.5) -- (2.9,0.7);
+\draw[line width=1pt] (0.35,0.5) -- (0.35,0.7);
+\draw[line width=1pt] (1.3,0.5) -- (1.3,0.7);
+\draw[line width=1pt] (2.65,0.5) -- (2.65,0.7);
 \draw[line width=1pt] (4.1,0.5) -- (4.1,0.7);
@@
 \draw[line width=1pt,line join=round,line cap=round,bemppgrey] (0,0) rectangle (4.6,0.5);
-\node at (2.3,0.25) {\texttt{bempp.api}};
-\draw[line width=1pt,line join=round,line cap=round,bempporange] (0,0.7) rectangle (1,1.2);
-\node at (0.5,0.95) {\footnotesize grids};
-\draw[line width=1pt,line join=round,line cap=round,bempporange] (1.2,0.7) rectangle (2.2,1.2);
-\node at (1.7,0.95) {\footnotesize spaces};
-\draw[line width=1pt,line join=round,line cap=round,bempporange] (2.4,0.7) rectangle (3.4,1.2);
-\node at (2.9,0.95) [text width=1.45cm,align=center] {\footnotesize grid\\[-1.2mm]functions};
+\node at (2.3,0.25) {\pyth{bempp.api}};
+\draw[line width=1pt,line join=round,line cap=round,bempporange] (0,0.7) rectangle (0.7,1.2);
+\node at (0.35,0.95) {\footnotesize \pyth{Grid}};
+\draw[line width=1pt,line join=round,line cap=round,bempporange] (0.9,0.7) rectangle (1.7,1.2);
+\node at (1.3,0.95) {\footnotesize \pyth{Space}};
+\draw[line width=1pt,line join=round,line cap=round,bempporange] (1.9,0.7) rectangle (3.4,1.2);
+\node at (2.65,0.95) {\footnotesize \pyth{GridFunction}};
 \draw[line width=1pt,line join=round,line cap=round,bempporange] (3.6,0.7) rectangle (4.6,1.2);
-\node at (4.1,0.95) {\footnotesize operators};
+\node at (4.1,0.95) {\footnotesize \pyth{Operator}};
 \draw[line width=1pt,line join=round,line cap=round,bemppgrey] (3.45,1.4) rectangle (4.75,1.9);
-\node at (4.1,1.65) {\footnotesize\texttt{bempp.core}};
+\node at (4.1,1.65) {\footnotesize\pyth{bempp.core}};
 

### 2021-04-06 bem.tex

@@
 Bempp-cl uses a Galerkin discretisation: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function.
-We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N x_j\phi_j$.
+We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients.
 In the most simple case, we can define the function \(\phi_j\) to be equal to 1 on the triangle $\tau_j$ and 0 everywhere else.
@@
 $$
-Ax = b
+\dmat{A}\dvec{x} = \dvec{b}
 $$
@@
 \begin{align*}
-A_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\ds[\by]\ds[\bx]\\
-b_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx].
+\dmat{A}_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\bx)\ds[\by]\ds[\bx]\\
+\dvec{b}_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx].
 \end{align*}
-In the case of piecewise constant trial and test functions, the definition of $A_{ij}$ simplifies to $A_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\ds[\by]\ds[\bx]$.
+In the case of piecewise constant trial and test functions, the definition of $\dmat{A}_{ij}$ simplifies to $\dmat{A}_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\ds[\by]\ds[\bx]$.
 
@@
 products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
-consists of either the dual representation in the forms of the integrals $b_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
+consists of either the dual representation in the forms of the integrals $\dvec{b}_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
 
-Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $A_{ij}$.
-For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $A_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights.
+Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$.
+For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights.
 In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical, corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.
 
-The values $b_i$ of the right-hand side vector $b$ are similarly computed through a numerical quadrature rule.
+The values $\dvec{b}_i$ of the right-hand side vector $b$ are similarly computed through a numerical quadrature rule.
 
-In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRes.
+In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRES.
 The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualisation.
@@
 	\item Definition of the trial space, test space and range space.
-	\item Discretisation into a matrix problem $Ax=b$.
+	\item Discretisation into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
 	\item Solution of the matrix problem by either a direct or iterative solver.

### 2021-04-06 bempp_overview.tex
add command for \ds

@@
 
-The main user-visible component of Bempp-cl is the module \texttt{bempp.api}, which defines all user interface functions and other high-level routines.
-In particular, it contains the definition of the main object types: \texttt{Grid}, \texttt{Space}, \texttt{GridFunction}, \texttt{Operator}.
-All computational backend routines are part of \texttt{bempp.core}.
-We are currently supporting a full Numba backend and an OpenCL backend.
+The main user-visible component of Bempp-cl is the module \pyth{bempp.api}, which defines all user interface functions and other high-level routines.
+In particular, it contains the definition of the main object types: \pyth{Grid}, \pyth{Space}, \pyth{GridFunction}, \pyth{Operator}.
+The computational backend routines are contained in the module \pyth{bempp.core}.
+Currently, we support a Numba and OpenCL backends.
 An overview of this structure is provided in \cref{fig:overview}.
+
+The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles (although for larger problems this cost can be reduced through the use of hierarchical matrices or fast multipole methods).
+Great care is therefore taken to ensure that the operator assembly routines perform as highly as possible.
+
+Once a user has defined in operator using Bempp-cl, the discretisation can be computed by calling the \pyth{weak_form} method.
+Upon calling this method, a regular integrator will be used to assemble all the interactions between non-adjacent elements, and a singular integrator will be used to compute the interactions between adjacent triangles (if the trial and test spaces are defined on different grids, this second integrator is not needed).
+Depending on the user's preferences, these integrators will internally use computational routines defined using either Numba or OpenCL.
+
+For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), andwhether the discretisation should proceed in single or double precision.
+The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.
+If the computational backend is Numba, the call is forwarded to the corresponding Numba discretisation routines and executed.
+
+For simple piecewise constant function spaces or other spaces, where the support of each basis function is localised to a single triangle, only one call to the computational routines is necessary.
+If the support of basis functions is larger than a single triangle, different threads may need to sum into the same global degree of freedom.
+We discuss this in more detail in \cref{sec:coloring}.
 
@@
 \begin{itemize}
-	\item Topological computations for the grid.
-Finding out all neighbour relationships between triangles requires parses through the grid data structure with linear complexity, which is accelerated with Numba.
-	\item Definition of local-to-global maps for function spaces. Again, this requires traversal through the grid and assigning relationships between global and local indices.
-	\item Grid Functions.
-A right-hand side function $f$ can be defined as a Python callable.
-This is just-in-time compiled via Numba and then the product with the corresponding basis functions integrated in each triangle via numerical quadrature, again via Numba accelerated routines.
-	\item Sparse matrices. Sparse surface matrices are assembled through Numba accelerated routines.
+  \item Computing the grid topology: this involves iterating through the grid to compute the neighbour relataionships between triangles.
+  \item Definition of local-to-global maps for function spaces: again, this requires traversal through the grid and assigning relationships between global and local indices.
+  \item Grid functions: a right-hand side function $f$ can be defined as a Python callable.
+This is just-in-time compiled via Numba and then the product with the corresponding basis functions is integrated in each triangle via numerical quadrature, again via Numba accelerated routines.
+  \item Computing sparse matrices.
 \end{itemize}
 
-The main computational complexity in discretising the boundary integral equation \cref{eq:bnd_integral} is coming from the discretisation of the left-hand side integral operator.
-Using dense methods this has quadratic complexity in the number of surface triangles.
-
-Once the user defines an operator with its associated discretisation spaces and calls the \textbf{weak\_form} method to discretise it the code calls a regular integrator to assemble all the interactions between non-adjacent elements and a singular integrator to compute the interactions between adjacent triangles (only necessary if trial and test space are defined on the same grid).
-The result is stored internally as a dense Numpy array.
-The corresponding discretisation routines are proxies that forward to computational routines in either Numba or OpenCL, depending on the user preferences.
-For OpenCL assembly the code checks additional parameters, such as the default vector length for SIMD operators (e.g.
-4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), and whether the discretisation should proceed in single or double precision.
-The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.
-If the computational backend is Numba, the call is forwarded to the corresponding Numba discretisation routines and executed.
-For simple piecewise constant function spaces or other spaces, where the support of each basis function is localised to a single triangle, only one call to the computational routines is necessary.
-If the support of basis functions is larger than a single triangle different threads may need to sum into the same global degree of freedom.
-We discuss this in more detail in \cref{sec:coloring}.
+As the cost of each of these processes is in general much smaller than the cost of operator discretisation, these can be permformed using Numba without any need to consider the use of OpenCL for potential further speed up.

### 2021-04-06 intro.tex

@@
-\chapterinitial{The Bempp boundary element library} (originally named BEM++ \cite{BEM++}) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
+\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
 The original release came with a simple Python wrapper to the C++ library.
@@
 
-This proved a burden when efforts started to modernise the library to be able to make better use of Single-Instruction-Multiple-Data (SIMD) optimisation on CPUs and offload computation to GPUs.
+This proved a burden when efforts started to modernise the library to be able to make better use of single-instruction-multiple-data (SIMD) optimisation on CPUs and offload computation to GPUs.
 In 2018, we decided to rewrite the computational core from scratch.

### 2021-04-06 opencl.tex

@@
 The above description was valid for kernels that do not take advantage of CPU SIMD operations are kernels that run directly on a GPU device.
-For SIMD optimization on CPUs we proceed slightly differently.
+For SIMD optimisation on CPUs we proceed slightly differently.
 The corresponding kernel in principle works as described above, but batches together the interaction on one test triangle with X trial triangles, where X is either 4, 8, or 16, depending on the number of available SIMD lanes.
-This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
-If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorized kernel.
+This strategy allows us to optimise almost all floating point operations within a kernel run for SIMD operation.
+If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorised kernel.
 
 Each kernel definition is stored in two variants, one with the ending \textit{\_novec.cl} and another one with the ending \textit{\_vec.cl}.
-The vectorized variant is configured via preprocessor directives for the desired number of vector lanes.
+The vectorised variant is configured via preprocessor directives for the desired number of vector lanes.
 Having to develop two OpenCL kernel codes for each operator is a certain amount of overhead.
-If we write a new kernel, we first implement the non-vectorized version.
-Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.
+If we write a new kernel, we first implement the non-vectorised version.
+Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorised or not, it is usually a matter of an hour or two to convert the non-vectorised kernel into a vectorised version.
 
-Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorize kernels by batching together work items on SIMD lanes, similarly to what we do manually.
+Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorise kernels by batching together work items on SIMD lanes, similarly to what we do manually.
 In our experience this works well for very simple kernels but often fails for more complex OpenCL kernels.

### 2021-04-06 img/kernel_header.tex
OpenCL highlighting

\begin{opencl}
#include "bempp_base_types.h"
#include "bempp_helpers.h"
#include "bempp_spaces.h"
#include "kernels.h"

__kernel void kernel_function(
    __global uint* testIndices, __global uint* trialIndices,
    __global int* testNormalSigns, __global int* trialNormalSigns,
    __global REALTYPE* testGrid, __global REALTYPE* trialGrid,
    __global uint* testConnectivity, __global uint* trialConnectivity,
    __global uint* testLocal2Global, __global uint* trialLocal2Global,
    __global REALTYPE* testLocalMultipliers, __global REALTYPE* trialLocalMultipliers,
    __constant REALTYPE* quadPoints, __constant REALTYPE* quadWeights,
    __global REALTYPE* globalResult,
    __global REALTYPE* kernel_parameters,
    int nTest, int nTrial, char gridsAreDisjoint)
{
  /* */
}
\end{opencl}

### 2021-04-06 bem.tex
american spelling

@@
 \section{Boundary element methods with Bempp}
-In this section, we provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretisation and solution.
+In this section, we provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretization and solution.
 
@@
 In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$.
-Once $\phi$ has been found, the normalised capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\ds[\bx]$.
+Once $\phi$ has been found, the normalized capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\ds[\bx]$.
 
@@
   \input{img/sphere}
-  \caption{Discretisation of a sphere into flat surface triangles.}
+  \caption{Discretization of a sphere into flat surface triangles.}
   \label{fig:triangulation}
@@
 
-The first step is the \textbf{discretisation of the surface} $\Gamma$.
+The first step is the \textbf{discretization of the surface} $\Gamma$.
 Surfaces are represented in Bempp-cl as a triangulation into flat triangles (see \cref{fig:triangulation}).
@@
 
-Once a triangulation is given we need to define the necessary data structures for the discretisation.
-Bempp-cl uses a Galerkin discretisation: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function.
+Once a triangulation is given we need to define the necessary data structures for the discretization.
+Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function.
 We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients.
@@
 
-To discretise \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
+To discretize \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
 $$
@@
 In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRES.
-The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualisation.
+The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.
 
@@
 	\item Definition of the trial space, test space and range space.
-	\item Discretisation into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
+	\item Discretization into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
 	\item Solution of the matrix problem by either a direct or iterative solver.
-	\item Evaluation of domain potential operators for visualisation and post-processing.
+	\item Evaluation of domain potential operators for visualization and post-processing.
 \end{enumerate}

### 2021-04-06 bempp_overview.tex
american spelling

@@
 
-Once a user has defined in operator using Bempp-cl, the discretisation can be computed by calling the \pyth{weak_form} method.
+Once a user has defined in operator using Bempp-cl, the discretization can be computed by calling the \pyth{weak_form} method.
 Upon calling this method, a regular integrator will be used to assemble all the interactions between non-adjacent elements, and a singular integrator will be used to compute the interactions between adjacent triangles (if the trial and test spaces are defined on different grids, this second integrator is not needed).
@@
 
-For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), andwhether the discretisation should proceed in single or double precision.
+For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), andwhether the discretization should proceed in single or double precision.
 The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.
-If the computational backend is Numba, the call is forwarded to the corresponding Numba discretisation routines and executed.
+If the computational backend is Numba, the call is forwarded to the corresponding Numba discretization routines and executed.
 
-For simple piecewise constant function spaces or other spaces, where the support of each basis function is localised to a single triangle, only one call to the computational routines is necessary.
+For simple piecewise constant function spaces or other spaces, where the support of each basis function is localized to a single triangle, only one call to the computational routines is necessary.
 If the support of basis functions is larger than a single triangle, different threads may need to sum into the same global degree of freedom.
@@
 
-Outside the operator discretisation, Numba is used in the following contexts:
+Outside the operator discretization, Numba is used in the following contexts:
 \begin{itemize}
@@
 
-As the cost of each of these processes is in general much smaller than the cost of operator discretisation, these can be permformed using Numba without any need to consider the use of OpenCL for potential further speed up.
+As the cost of each of these processes is in general much smaller than the cost of operator discretization, these can be permformed using Numba without any need to consider the use of OpenCL for potential further speed up.

### 2021-04-06 intro.tex
american spelling

@@
 
-This proved a burden when efforts started to modernise the library to be able to make better use of single-instruction-multiple-data (SIMD) optimisation on CPUs and offload computation to GPUs.
+This proved a burden when efforts started to modernize the library to be able to make better use of single-instruction-multiple-data (SIMD) optimization on CPUs and offload computation to GPUs.
 In 2018, we decided to rewrite the computational core from scratch.
-The aims were to support explicit SIMD optimisation on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
+The aims were to support explicit SIMD optimization on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
 These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.

### 2021-04-06 bem.tex
start working through opencl section

@@
-\section{Boundary element methods with Bempp}
+\section{BOUNDARY ELEMENT METHODS WITH BEMPP}
 In this section, we provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretization and solution.

### 2021-04-06 main.tex
Finished most of the OpenCL section.
edititing first section
Make tikz version of figure
add command for \ds
working on bempp section
american spelling
start working through opencl section

@@
 
+\usepackage{xcolor}
+\definecolor{bemppgrey}{HTML}{555555}
+\definecolor{bempporange}{HTML}{FF7A1A}
+
+\usepackage{pythonhighlight}
+
+\usepackage{tikz}
+\usetikzlibrary{arrows,decorations.pathreplacing}
+
+\usepackage{ifthen}
+
 \usepackage{upmath}
+\usepackage{upgreek}
 \usepackage{amsmath, amssymb}
 
-\def\bx{\mathbf{x}}
-\def\by{\mathbf{y}}
+\usepackage{cleveref}
+\crefname{equation}{Equation}{Equations}
+\Crefname{equation}{Equation}{Equations}
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{section}{Section}{Sections}
+\Crefname{section}{Section}{Sections}
+
+% Bold vector (for points in R^3)
+\newcommand{\bvec}[1]{\boldsymbol{#1}}
+\def\bx{\bvec{x}}
+\def\by{\bvec{y}}
+\newcommand{\ds}[1][]{\,\mathrm{d}s\ifthenelse{\equal{#1}{}}{}{_{#1}}}
+
+% Discrete vector and discrete matrix
+\newcommand{\dmat}[1]{\mathbf{#1}}
+\newcommand{\dvec}[1]{\mathbf{#1}}
 
@@
 \jname{IT Professional}
-\pubyear{2019}
+\pubyear{2021}
 \newtheorem{theorem}{Theorem}
@@
 Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba.
-The OpenCL backend implements kernels for GPUs and SIMD optimised for CPUs.
+The OpenCL backend implements kernels for GPUs and SIMD optimized for CPUs.
 In this paper we will discuss the design of Bempp-cl, provide performance comparisons on different compute devices and discuss the advantages and disadvantages of OpenCL as compared to Numba.
@@
 \input{bem}
+\input{bempp_overview}
+\input{opencl}
+
+\bibliographystyle{IEEEtran}
+\begin{thebibliography}{1}
+
+\bibitem{Bempp-cl}
+T. Betcke and M. W. Scroggs, ``Bempp-cl: A fast Python based just-in-time compiling boundary element library,''
+{\it Journal of Open Source Software}, vol. 6, no. 59, p. 2879, 2021.
+\textsc{doi}: \href{https://dx.doi.org/10.21105/joss.02879}{10.21105/joss.02879}
+\end{thebibliography}
 

### 2021-04-06 opencl.tex
OpenCL highlighting
american spelling
start working through opencl section

--- 2021-04-06 opencl.tex
@@
-\section{Assembling boundary integral operators with OpenCL}
+\section{ASSEMBLING BOUNDARY INTEGRAL OPERATORS WITH OPENCL}
 
-In this section we discuss in more detail the assembly of boundary integral operators with OpenCL
+In this section, we discuss in more detail the assembly of boundary integral operators with OpenCL
 and how we integrated this into our Python workflow. We start with a brief introduction to OpenCL and then
@@
 OpenCL (\url{https://www.khronos.org/opencl/}) is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
-At its core OpenCL executes compute kernels that can be written in C99, or more recently, in C++.
+At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++.
 The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.
-OpenCL splits the computational tasks into work-items, each of which represents a single execution of a compute kernel.
-Work items are grouped together into work-groups, which share local memory.
-Barrier synchronization is only allowed within a work-group.
-All work-items are uniquely indexed by a one, two, or three dimensional index space, called NDRange.
-Kernels are launched onto a compute device (e.g.
-CPU or GPU) from a host device.
+
+OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel.
+Work items are grouped together into work groups, which share local memory.
+Barrier synchronization is only allowed within a work group.
+All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange.
+Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device.
 OpenCL allows kernels to be loaded as strings and compiled on the fly for a given device, making it well suited for launching from high-productivity languages.
-To launch an OpenCL kernel the following steps need to be performed.
+
+To launch an OpenCL kernel the following steps need to be performed:
 \begin{itemize}
@@
 	\center
-	\includegraphics[width=12cm]{img/kernel_header}
+	\input{img/kernel_header}
 	\caption{Definition of the OpenCL compute kernel for scalar integral equations.}
@@
 These include the names of the test and trial space, the name of the function that evaluates the integral kernel, choosing between single and double precision types, and for SIMD enhanced kernel the vector length of the SIMD types.
-For example, in Figure \ref{fig:kernel_definition} all floating point types have the name REALTYPE. This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.
+For example, in Figure \ref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}.
+This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.
 
@@
 
-\section{Numba assembly of integral operators}
+\section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}
 
@@
 
-\section{SIMD optimized evaluation of the Laplace kernel with Numba}
+\section{SIMD OPTIMIZED EVALUATION OF THE LAPLACE KERNEL WITH NUMBA}

### 2021-04-07 opencl.tex
working through opencl section

@@
 
-To launch an OpenCL kernel the following steps need to be performed:
-\begin{itemize}
-	\item Initialize buffers on the compute device.
-	\item Compute relevant data from host to the compute device.
-	\item Load a kernel string and just-in-time compile it for the device.
-	\item Execute the kernel for the given buffers and chosen index space.
-	\item Copy results back to the host.
-\end{itemize}
-An excellent feature of OpenCL is its very good support for vectorized operations.
-OpenCL provides vector data types, such as \textbf{double4}, that can hold four double values in a SIMD register and defines a number of standard operations for these vector types.
+To launch an OpenCL kernel the user must first initialize buffers on the compute device, then copy any relevant data from host to the compute device.
+A kernel string can then be loaded and just-in-time compiled for the device.
+The kernel is then run, and the results can be copied back to the host.
+
+OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types.
+For example, the type \ocl{double4} will allow four double values to be held in a SIMD register.
 This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.
 
-Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner, which automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.
+Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner \cite{pyopencl}.
+PyOpenCL automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.
 
@@
 All operators have the same interface and are launched in the same way.
-In the first step the relevant data will need to be copied to the compute device.
+In the first step, the relevant data will need to be copied to the compute device.
 This data consists of:
@@
 \end{itemize}
-The actual kernel definition is shown in Figure \ref{fig:kernel_definition}.
+The actual kernel definition is shown in \cref{fig:kernel_definition}.
 Before the kernel can be launched, it needs to be configured and just-in-time compiled.

### 2021-04-09 numba.tex
Started performance section

\section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}

The main focus of Numba within Bempp-cl is to provide accelerated implementations for linear complexity routines, such as grid iterations, integration of functions over grids or assembly of certain sparse matrices. However, we do also provide a fall-back implementation of the OpenCL dense operators in Numba. Here, we use classic loop parallelism. Each loop iteration is the assembly of one test element with all trial elements. Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations. However, a much smaller fraction of operations is SIMD optimized due to the lack of targeted SIMD constructs in Numba.

### 2021-04-09 opencl.tex
Started performance section

@@
 
-\section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}
-
-The Numba assembly of integral operators is structured differently from the OpenCL assembly.
-In OpenCL we create a large number of work-items that are executed in parallel on the compute device by the runtime system with the details of the parallel execution left to the device driver.
-For Numba we use classical loop based parallelism.
-Each loop iteration is the assembly of one test element with all trial elements.
-Within each loop we partially optimized for SIMD instruction, though this depends on the autovectorization of the just-in-time compiler and cannot be explicitly targeted in Numba.
-Again, we separate regular and singular interactions.
-In the next section we give some more details on SIMD optimization in Numba at the example of the Laplace kernel.
-
-\section{SIMD OPTIMIZED EVALUATION OF THE LAPLACE KERNEL WITH NUMBA}
+\subsection{Avoiding data races in the global assembly}

### 2021-04-09 performance.tex
Started performance section

\section{PERFORMANCE BENCHMARKS}

In this section we want to provide a number of performance benchmarks. The tests are running on an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz. The CPU supports AVX, AVX2, and AVX-512.
The tests are running on a Dell Precision 7740 Workstation Laptop, meaning the burst clock can not be maintained due to cooling. As GPU we use an Nvidia RTX 3000 GPU. All benchmarks are running within Linux. For OpenCL on the GPU we use the Nvidia GPU drivers and as CPU runtime we compare the open-source POCL driver against the Intel CPU OpenCL runtime environment.

\subsection{Benchmarking a simple Laplace single-layer boundary operator}

\subsection{Complex arithmetic benchmark - A Helmholtz Kernel}

\subsection{CPU vs GPU for evaluating domain potential operators}

\subsection{Solving a complete electromagnetic problem}

### 2021-04-13 bem.tex
Added triangles tikz image

@@
 
+\begin{figure*}
+  \centering
+  \input{img/triangles}
+  \caption{The four types of intersection of two triangles: the triangles can be (L to R) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle.
+In the first case, standard quadrature is used.
+In the other three cases, singular quadrature rules must be used.}
+  \label{fig:triangles}
+\end{figure*}
+
 The first step is the \textbf{discretization of the surface} $\Gamma$.
@@
 In this step, topology data is also computed: in particular, for each triangle, we compute the neighboring triangles and the type of intersection (i.e.
-are they connected by an edge or a vertex).
+are they connected by an edge or a vertex, see \cref{fig:triangles}).
 

### 2021-04-13 img/triangles.tex
Added triangles tikz image

\input{img/triangles_command}
%
\begin{tikzpicture}
\clip (-1.5,-1.5) rectangle (1.5,1.5);
\drawtriangles
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (0.2723649506759587,1.2015357806708773) -- (0.01759915424630132,0.686970401916699) -- (0.6676122663150503,0.7464322913822609) -- cycle;
\end{tikzpicture}
%
\hfill
%
\begin{tikzpicture}
\clip (-1.5,-1.5) rectangle (1.5,1.5);
\drawtriangles
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (0.36044208773959274,0.13283689310102198) -- (0.01759915424630132,0.686970401916699) -- (-0.283541653228761,0.14673508308776784) -- cycle;
\end{tikzpicture}
%
\hfill
%
\begin{tikzpicture}
\clip (-1.5,-1.5) rectangle (1.5,1.5);
\drawtriangles
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.9219481904610344,0.1627856205506708) -- (-0.283541653228761,0.14673508308776784) -- (-0.617476093555047,0.7029236883312638) -- cycle;
\end{tikzpicture}
%
\hfill
%
\begin{tikzpicture}
\clip (-1.5,-1.5) rectangle (1.5,1.5);
\drawtriangles
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\begin{scope}[shift={(0.02,0.02)}]
\draw[line width=1pt,bempporange,fill=bempporange,fill opacity=0.3] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\end{scope}
\end{tikzpicture}

### 2021-04-13 img/sphere.py

@@ -4,7 +4,6 @@ g = bempp.api.shapes.sphere(h=0.3)
 
 
 with open("sphere.tex", "w") as f:
-
     f.write("\\begin{tikzpicture}[x={(1.732cm,-1cm)},y={(1.732cm,1cm)},z={(0,2cm)}]\n")
 
     for i, j in enumerate(g.vertices.T):
@@ -16,3 +15,27 @@ with open("sphere.tex", "w") as f:
     for _, (i, j, k) in triangles:
         f.write(f"\\draw[fill=white, fill opacity=0.8] (v{i}) -- (v{j}) -- (v{k}) -- cycle;\n")
     f.write("\\end{tikzpicture}")
+
+
+def to2d(i, j, k):
+    return 1.732 * i + 1.732 * j, -i + j + 2 * k
+
+
+with open("triangles_command.tex", "w") as f:
+    f.write("\\newcommand{\\drawtriangles}{\n")
+
+    triangles = [(sum(g.vertices.T[i] for i in t) / 3, t) for t in g.elements.T]
+
+    triangles.sort(key=lambda x: x[0][0] + x[0][2] - x[0][1])
+    for _, (i, j, k) in triangles:
+        if _[0] + _[2] - _[1] > 0.6:
+            p1 = to2d(*g.vertices.T[i])
+            p2 = to2d(*g.vertices.T[j])
+            p3 = to2d(*g.vertices.T[k])
+            if -1.5 > max(p1[0], p2[0], p3[0]) or 1.5 < min(p1[0], p2[0], p3[0]):
+                continue
+            if -1.5 > max(p1[1], p2[1], p3[1]) or 1.5 < min(p1[1], p2[1], p3[1]):
+                continue
+
+            f.write(f"\\draw ({p1[0]},{p1[1]}) -- ({p2[0]},{p2[1]}) -- ({p3[0]},{p3[1]}) -- cycle;\n")
+    f.write("}")

### 2021-04-13 bem.tex
added to intro

@@
   \input{img/triangles}
-  \caption{The four types of intersection of two triangles: the triangles can be (L to R) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle.
+  \caption{The four types of intersection of two triangles: the triangles can be (left to right) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle.
 In the first case, standard quadrature is used.

### 2021-04-13 intro.tex
added to intro

@@
 [To be continued with some more details about content and overview of the Sections]
+
+\subsection{The three language problem}
+The original BEM++ was troubled by what is often called the two language problem.
+It is common for programming language to be either easy for humans to write (e.g., Python) or easy for computers to run and acheive high performance (e.g., C++).
+It is not common, however, for a language to do both of these.
+Due to this, it is common in scientific computing libraries to write a library in a fast low-level language such as C++, while providing a user interface in a higher-level language such as Python.
+
+This was the model followed by BEM++, but the problem would perhaps better be described as a three language problem: as well as the Python and C++ code contained in libraries like this, it is common to also include a significant amount of code in a third interfacing language.
+In the case of BEM++, this third language was Cython, an extension of Python with C data types that can be compiled to include functionality from C++ libraries.
+
+Due to the three languages involved, making changes to the library would often mean having to duplicate changed in three places, with many class structures duplicated in all three languages.
+This made seemingly simple changes in onorous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.
+
+\subsection{Delegating computations with PyOpenCL}
+Prompted by a desire to simplify the library as well as to be able to run on a wide range of CPU and GPU devices, we began a full re-write in 2018, which led to Bempp-cl.
+The aims of this rewrite were to support explicit single-instruction-multiple-data (SIMD) optimization on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
+These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.
+
+In addition to the performance benefits, the library redesign has greatly improved the issues related to the three language problem.
+Both OpenCL and Numba are use to compile functions.
+Each function is provided all the data is needs as inputs, so there is no need to duplicate any class structure outside Python.
+The need for a third interfacing language is removed, as the interfacing to the OpenCL kernels is handled by PyOpenCL.
+
+Boundary element methods are particularly suited to this library model, as the main performance critical task is the computation of discrete operators.
+Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great detriment to performance.

### 2021-04-13 intro.tex
finish expanding intro

@@
 
-This proved a burden when efforts started to modernize the library to be able to make better use of single-instruction-multiple-data (SIMD) optimization on CPUs and offload computation to GPUs.
-In 2018, we decided to rewrite the computational core from scratch.
-The aims were to support explicit SIMD optimization on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
-These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.
-
-At the end of 2019, we released the first version (0.1) of Bempp-cl \cite{Bempp-cl}.
+At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl \cite{Bempp-cl}.
 This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use.
 Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp.
-In this article, we want to discuss the design choices behind Bempp-cl and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.
-
-[To be continued with some more details about content and overview of the Sections]
+In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.
 
@@
 Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great detriment to performance.
+
+\bigskip
+
+We begin this article by giving an overview of the boundary element method, and looking at how Bempp can be used to implement such problems.
+Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.

### 2021-04-13 main.tex
working through opencl section
Started performance section
editing up to simd section

@@
 \input{opencl}
+\input{numba}
 
@@
 \bibitem{Bempp-cl}
-T. Betcke and M. W. Scroggs, ``Bempp-cl: A fast Python based just-in-time compiling boundary element library,''
+T. Betcke and M. W. Scroggs,
+``Bempp-cl: A fast Python based just-in-time compiling boundary element library,''
 {\it Journal of Open Source Software}, vol. 6, no. 59, p. 2879, 2021.
-\textsc{doi}: \href{https://dx.doi.org/10.21105/joss.02879}{10.21105/joss.02879}
+\textsc{doi}: \href{https://dx.doi.org/10.21105/joss.02879}{10.21105/joss.02879}.
+
+\bibitem{pyopencl}
+A. Kloeckner, N. Pinto, Y. Lee, B. Catanzaro, P. Ivanov, and A. Fasih,
+``PyCUDA and PyOpenCL: A scripting-based approach to GPU run-time code generation,''
+{\it Parallel Computing}, vol. 38, no. 3, pp. 157--174, 2012.
+\textsc{doi}: \href{https://dx.doi.org/10.1016/j.parco.2011.09.001}{10.1016/j.parco.2011.09.001}.
+
+\bibitem{opencl}
+OpenCL, \url{https://www.khronos.org/opencl/}.
 \end{thebibliography}

### 2021-04-13 opencl.tex
editing up to simd section

@@
 
-OpenCL (\url{https://www.khronos.org/opencl/}) is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
+OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
 At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++.
@@
 \end{itemize}
-The actual kernel definition is shown in \cref{fig:kernel_definition}.
+
+An example kernel definition is shown in \cref{fig:kernel_definition}.
 Before the kernel can be launched, it needs to be configured and just-in-time compiled.
 Kernel configuration happens through C-style preprocessor definitions that are passed through the just-in-time compiler.
-These include the names of the test and trial space, the name of the function that evaluates the integral kernel, choosing between single and double precision types, and for SIMD enhanced kernel the vector length of the SIMD types.
-For example, in Figure \ref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}.
-This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.
+These include the names of the test and trial space, the name of the function that evaluates the Green's function, whether we are using single or double precision types, and (for SIMD enhanced kernels) the vector length of the SIMD types.
+For example, in \cref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}. This is substituted with either \ocl{float} or \ocl{double} during just-in-time compilation.
 
-Each work-item computes via numerical quadrature all interactions via the integral kernel of basis functions on the trial element with basis functions on the test element.
-Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangle are adjoint or identical.
-For this it simply checks if at at least one of the node indices of the test triangle is identical to one of the node indices of the trial triangle.
-If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer.
-The reason is that for adjoint triangles a separate singular quadrature rule is used.
+Each work-item computes (using numerical quadrature) all interactions of basis functions on the trial element with basis functions on the test element.
+Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangles are adjacent or identical (see \cref{fig:triangles}).
+To do this, it simply checks if at at least one of the node indices of the test triangle is equal to one of the node indices of the trial triangle.
+If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer: for these triangles, separate singular quadrature rules need to be used.
 The effect is that a few work-items do work that is discarded.
-However, in a grid with $N$ elements the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of interactions is $N^2$.
+However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of interactions is $N^2$.
 Hence, only a tiny fraction of work-items are discarded.

### 2021-04-13 opencl.tex
work through rest of opencl section

@@
 \subsection{SIMD optimized kernels}
-
-The above description was valid for kernels that do not take advantage of CPU SIMD operations are kernels that run directly on a GPU device.
-For SIMD optimization on CPUs we proceed slightly differently.
-The corresponding kernel in principle works as described above, but batches together the interaction on one test triangle with X trial triangles, where X is either 4, 8, or 16, depending on the number of available SIMD lanes.
+When we are running on a CPU and want to take advantage of available SIMD optimizations, we need to make a few modifications to our approach.
+The corresponding kernel works similarly to described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes).
 This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
@@
 
-Each kernel definition is stored in two variants, one with the ending \textit{\_novec.cl} and another one with the ending \textit{\_vec.cl}.
+Each kernel definition is stored in two variants, one with the ending \texttt{\_novec.cl} and another one with the ending \texttt{\_vec.cl}.
 The vectorized variant is configured via preprocessor directives for the desired number of vector lanes.
-Having to develop two OpenCL kernel codes for each operator is a certain amount of overhead.
-If we write a new kernel, we first implement the non-vectorized version.
-Then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.
+Having to develop two OpenCL kernel codes for each operator creates a certain amount of overhead, but once we have implemented the non-vectorized version then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually only a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.
 
-Alternatively, some CPU OpenCL runtime environments have the option to try and auto-vectorize kernels by batching together work items on SIMD lanes, similarly to what we do manually.
-In our experience this works well for very simple kernels but often fails for more complex OpenCL kernels.
+Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work items on SIMD lanes, similar to what we do manually.
+In our experience, this works well for very simple kernels but often fails for more complex OpenCL kernels.
 This is why we decided to implement this strategy manually.
 
-A completely different SIMD strategy is based on batching together quadrature evaluations within a single test/triangle pair.
-There are two disadvantages to this approach.
-First, it only works well if the number of quadrature points is a multiple of the available SIMD lanes.
-Second, other operations such as the geometry calculations for each element then can not be SIMD optimized as these are only performed once per test/triangle pair.
+A completely different SIMD strategy could be taken by batching together quadrature evaluations within a single test/trial pair.
+There are two disadvantages to this approach: first, it only works well if the number of quadrature points is a multiple of the available SIMD lanes.
+Second, other operations such as the geometry calculations for each element then cannot be SIMD optimized as these are only performed once per test/trial pair.
 
 \subsection{Assembling the singular part of integral operators}
+The assembly of the singular part of an integral operator works a bit differently.
+Remember that the singular part consists of triangle parts which are adjacent to each other or identical (the three later cases in \cref{fig:triangles}): there are $\mathcal{O}(N)$ such pairs.
+We are using fully numerical quadrature rules for these integrals that are based on subdividing the four-dimensional integration domain and using transformation techniques to remove the singularities.
+This gives highly accurate evaluation of these integration pairs but requires a large number (typically over $1000$) quadrature points per triangle pair.
 
-The assembly of the singular part of an integral operator works a bit differently.
-Remember that the singular part consists of triangle parts which are adjacent to each other or identical.
-There are $\mathcal{O}(N)$ of such pairs.
-We are using fully numerical quadrature rules for these integrals that are based on subdividing the four-dimensional integration domain and using transformation techniques to remove the singularities.
-This gives highly accurate evaluations of these integration pairs but requires a large number (typically over $1000$) quadrature points per triangle pair.
-Hence, we create one workgroup for each triangle pair and inside this workgroup have a number of work-items that are evaluating the quadrature rules and then at the end sum up the result together.
-Depending on how two triangles are in relationship to each other different types of singular quadrature rules are needed.
-We solve this pre-loading all possible quadrature rules onto the device and also store for each triangle pair an index into the required quadrature rule so that the kernel function can select the correct rule to evaluate.
-For the singular quadrature rules we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above.
-At the end, the singular integral values are either summed into the overall result matrix, or if desired by the user, stored as separate sparse matrix.
+For this assembly, we create one workgroup for each triangle pair.
+Inside this workgroup, we have a number of work-items that evaluate the quadrature rules then sum up the results.
+Depending on how two triangles are related to each other, different types of singular quadrature rule are needed.
+We solve this by pre-loading all possible quadrature rules onto the device, and also store for each triangle pair an index pointing to the required quadrature rule so that the kernel function can select the correct rule to evaluate.
+For the singular quadrature rules, we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above.
+At the end, the singular integral values are either summed into the overall result matrix, or (if desired by the user) stored as separate sparse matrix.
 

### 2021-04-13 opencl.tex
First draft finished.

--- 2021-04-09 opencl.tex
@@
 \subsection{Avoiding data races in the global assembly}
+
+Data races in global assembly routines are a problem whenever different elements need to sum into the same global degree of freedom.
+To solve this we use standard color coding techniques to split up the computations into chunks that access different data regions.
+To this end, we define two elements as neighbors if they share at least one global degree of freedom.
+Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.
+This information is then later used to run the compute kernels color by color.
+OpenCL parallelises over test elements and trial elements.
+Here we therefore have to iterate over the product space of possible color combinations.
+In Numba we only parallelise over test elements.
+Here it is sufficient to iterate over all possible colors in the test space.

### 2021-04-13 performance.tex
First draft finished.

@@
 The CPU supports AVX, AVX2, and AVX-512.
-The tests are running on a Dell Precision 7740 Workstation Laptop, meaning the burst clock can not be maintained due to cooling.
+The tests are running on a Dell Precision 7740 Workstation Laptop with 64GB RAM.
 As GPU we use an Nvidia RTX 3000 GPU.
-All benchmarks are running within Linux.
+All benchmark tests were performed in Linux.
 For OpenCL on the GPU we use the Nvidia GPU drivers and as CPU runtime we compare the open-source POCL driver against the Intel CPU OpenCL runtime environment.
+All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT Compiler did not skew the results.
+Though hardly noticeable by users, for smaller experiments the JIT phase typically takes longer than the actual computation.
 
-\subsection{Benchmarking a simple Laplace single-layer boundary operator}
+\subsection{Dense Operator assembly}
 
-\subsection{Complex arithmetic benchmark - A Helmholtz Kernel}
+Let us start with benchmarking the dense operator assembly. We assemble the matrix $\dmat{A}$ defined by
+$$
+\dmat{A}_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\pi |\bx -\by|}(\bx)\ds[\by]\ds[\bx]\\
+$$
+with $\Gamma$ being the unit sphere.
+For the basis functions $\phi_j$ and test functions $\psi_i$ we compare two cases, piecewise constant functions for both (P0 functions), or nodal, piecewise linear, and globally continuous functions for both (P1 functions).
+In the P0 case each triangle is associated with just one piecewise constant function.
+In the P1 case each triangle is associated with $3$ linear basis functions, one for each node of the triangle.
 
-\subsection{CPU vs GPU for evaluating domain potential operators}
+We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the POCL OpenCL runtime driver.
+We run the tests in single precision and double precision.
+The native vector width for both drivers in single precision is $8$, and and double precision $4$, which corresponds to AVX instructions.
+In Bempp-cl this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
+Within the assembly almost all floating point instructions are manually vectorized to take advantage of this.
+We should hence see up a factor 2 speed-up between single and double precision assembly.
 
-\subsection{Solving a complete electromagnetic problem}
+\begin{figure}
+	\center
+	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp}
+	\caption{Comparison of the performance of POCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
+	\label{fig:intel_pocl_laplace_cmp}
+\end{figure}
+
+The results for a grid with 32,768 elements are shown in Figure \ref{fig:intel_pocl_laplace_cmp}.
+We see the expected speed-up between single precision and double precision evaluation.
+What is interesting to note is the difference between the Intel and the POCL runtime environment.
+For P0 basis functions the Pocl driver (black bar) significantly outperforms the Intel driver (blue bar) in both, single and double precision.
+For P1 basis functions the Intel driver gives better performance.
+The kernel source code is the same in both cases and manual vectorization of floating point operations is performed in the same way.
+However, in the P1 case significantly more data needs to be copied between a work-item and the global memory since the result of the interaction of a test triangle with a trial triangle is a local $3\times 3$ matrix containing all interactions between the three test basis functions on a triangle and the three trial basis functions.
+
+In Figure \ref{fig:pocl_single_layer} we compare specifically the performance of single precision and double precision evaluation for the POCL driver for various grid sizes in the case of a P0 basis.
+We can see that the speed-up for larger grid sizes is even slightly faster than just a factor two.
+The final data point in this graph corresponds to the grid size used for Figure \ref{fig:intel_pocl_laplace_cmp}.
+
+\begin{figure}
+	\center
+	\includegraphics[width=7cm]{img/pocl_single_layer}
+	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using Pocl and P0 basis functions.}
+	\label{fig:pocl_single_layer}
+\end{figure}
+
+In Bempp-cl we can easily switch between CPU Assembly, GPU Assembly and Numba Assembly.
+In Figure \ref{fig:cpu_gpu_numba_compare} we present a table for a grid size 2048 elements to compare these modes.
+\begin{figure}
+\begin{center}
+\begin{tabular}{l|c|c}
+	        &   single      &    double\\
+	        \hline
+	 POCL   &   0.05s       &    0.08s\\
+	 Numba  &   0.25s       &    0.25s\\
+	 GPU    &   0.11s       &    3s\\
+\end{tabular}
+\end{center}
+\caption{Comparison of POCL, Numba and GPU Assembly for a grid with 2048 elements.}
+\label{fig:cpu_gpu_numba_compare}
+\end{figure}
+The speed differences are striking.
+GPU assembly in single precision is a factor two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision).
+Numba is five times slower for single precision and still around three times slower for double precision than POCL.
+The GPU behaviour is due to the data transfer.
+We have done numerous experiments with GPU assembly of dense boundary integral operators.
+While the GPU kernels themselves are extremely fast, data transfer over the bus is destroying performance.
+But even for medium sized problems we have to transfer the data back to the main memory as GPU RAM is too limited to keep a number of dense matrices of dimension a few ten thousand on the device.
+
+A different solution is to compute not the matrix but the matrix vector product $Ax$ on the device by recomputing all matrix elements in each-matvec on-the-fly but not storing them.
+We have done internal experiments with this and it gives significant speed-ups as compared to CPU evaluation as we now only need to transfer single vectors over the bus.
+But it is not competitive for larger problems compared to accelerated methods, such as Fast Multipole Methods due to the quadratic complexity of direct evaluation of the matvec as compared to linear complexity in the FMM.
+For smaller problem sizes it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver.
+Hence, there is only limited practical relevance for such on-the-fly GPU evaluation of boundary integral operators.
+There is though very significant practical relevance of on-the-fly evaluation for the evaluation of domain potential operators for visualization and post-processing as we will see in the next section.
+
+The Numba performance difference is interesting.
+The main issue, we think, is significantly less usage of AVX vectorized instructions.
+We have taken care to optimize the Green's function evaluation for vectorized evaluation.
+But the loops over integration points and other operations auto-vectorize very badly by just looping over all trial triangle for a single test triangle, as we currently do.
+We could tune our code for better auto-vectorization in Numba.
+However, this would be work with little benefit as we have highly optimized hand-tuned OpenCL kernels already.
+We therefore recommend Numba for operator assembly only as a fallback if no OpenCL runtime is available.
+
+\subsection{Evaluating domain potentials for post-processing of electromagnetic problems}
+
+Once an integral equation is solved one is usually interested in evaluating the solution not only at the boundary but also at points away from it. Hence, we want to evaluate
+$$
+f(\bx) = \int_{\Gamma}g(\bx), \by)\phi(\by) \ds[\by]
+$$
+for many points $\bx$ away from the boundary $\Gamma$.
+For example, if we want to visualize a solution the points $\bx$ would arise from some plotting grid.
+Typical of this operation that we usually only want to do it once or a few times at the end of a calculation.
+Discretising this operation into a dense matrix and then evaluating the dense matrix-vector product is not practical for larger sizes.
+For very large problems with hundreds of thousands of elements we use Fast Multipole or other accelerated approximate methods.
+However, for problem sizes of a few ten thousand elements up to around a hundred thousand elements (depending on the problem at hand), direct evaluation of this integral for every value $\bx$ is highly efficient.
+We won't go into the details of the corresponding OpenCL kernels here but want to show some results that demonstrate the performance of CPU vs GPU.
+While for the dense assembly of boundary integral operators the bus transfer of the dense matrix limited performance, here we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$ and then back to the host the values at the points $x$.
+
+For this section we used the evaluation of the electric potential operator, defined by
+\begin{align*}
+\mathcal{E\mathbf{p}}(\bx) &= ik\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\nonumber\\
+&-\frac{1}{ik}\nabla_{\bx}\int_{\Gamma}\text{div}\,\mathbf{p}(\by)g(\bx, \by)\ds[\by]
+\end{align*}
+with $g(\bx, \by) = \frac{e^{ik|\bx - \by|}}{4\pi|\bx - \by}$ the Helmholtz Green's function, and
+$k$ the wavenumber of the problem.
+The function $\mathbf{p}(\by)\in\mathbb{R}^3$ is a vectorial function, leading to an overall vectorial solution at each point $\bx$.
+For details about the implementation of electromagnetic problems in Bempp see XXX.
+Here, we again use a grid with 32,768 elements and so-called (Rao, Wilton, and Glisson) edge based basis functions.
+We use 50,000 random evaluation points in the exterior of the unit sphere.
+We set $k=1.0$, though the value is not relevant for the performance.
+
+In Figure \ref{fig:efield_domain_potential} we compare the performance of GPU evaluation with that of the POCL driver.
+In single-precision the GPU significantly outperforms the CPU, and even in double-precision the Quadro RTX is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
+
+\begin{figure}
+	\center
+	\includegraphics[width=7cm]{img/efield_domain_potential}
+	\caption{Evaluation of an electric field potential operator on CPU via POCL vs GPU}
+	\label{fig:efield_domain_potential}
+\end{figure}
+
+\section{SUMMARY}
+
+With Bempp-cl we have created a Python platform that achieves high-performance through use of modern JIT technologies.
+Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
+Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate into other applications with Python interfaces.
+An example is the recent work of integrating Bempp-cl and the Exafmm library for electrostatic virus-scale simulations [XXX].
+Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends.
+We currently support Numba and OpenCL as compute backends.
+Other technologies could be added easily.
+OpenCL itself has proved a valuable choice for the purpose of this library.
+We have all kernels available in CPU and GPU optimized versions and the user can easily move between CPU and GPU based compute devices by a simple parameter change, allowing us to support a wide range of hardware.
+Here, we have demonstrated Nvidia benchmarks.
+But it would have made no difference to use AMD or Intel GPUs.
+
+A disadvantage of our approach is that using OpenCL kernels is mixed-language development due to the OpenCL kernels written in C99.
+Using Numba throughout would be a much more native Python experience.
+However, while Numba is constantly improving, achieving optimal performance for complex operations is difficult.
+OpenCL shines here.
+It makes explicit SIMD operations very easy through dedicated constructs.
+Moreoever, OpenCL kernels are completely stack based functions, allowing much better optimisations while Numba needs to create every object dynamically, even very small arrays for objects such as coordinate vectors.
+
+Also, one has to consider for this approach the type of operations that are required to be accelerated.
+For the dense assembly of integral operators we have very simple data structures that can easily be passed to compute kernels.
+More complex data structures with larger mixture of data movement operations and computations (e.g.
+tree-based algorithms), are much harder to accelerate since the Python layer imposes limits here on the performance.
+
+Overall, with the model of mixed Python/OpenCL/Numba development we have created a flexible and easy to extend platform for integral equation computations and it was an effort that has payed back in terms of simple development without sacrificing performance.
+Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.
+
+The current focus of further developments is on MPI implementations for cluster computing via \verb@mpi4py@.
+We have also made big steps forward for large problems by creating a black-box FMM interface that is currently implemented for Exafmm, which has allowed us to solve problems with 10 million elements on a single workstation.
+This shows that Python focused development (with some native routines in lower-level languages) is a scalable model and we are aiming to exploit this scalable further as we move from single workstation computations to large cluster problems.

### 2021-04-13 intro.tex

@@
 Boundary element methods are particularly suited to this library model, as the main performance critical task is the computation of discrete operators.
-Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great detriment to performance.
+Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great decrease in performance.
 

### 2021-04-13 performance.tex
Updated image files with the right font

@@
 	\center
-	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp}
+	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp.pdf}
 	\caption{Comparison of the performance of POCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
@@
 	\center
-	\includegraphics[width=7cm]{img/pocl_single_layer}
+	\includegraphics[width=7cm]{img/pocl_single_layer.pdf}
 	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using Pocl and P0 basis functions.}
@@
 	\center
-	\includegraphics[width=7cm]{img/efield_domain_potential}
+	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
 	\caption{Evaluation of an electric field potential operator on CPU via POCL vs GPU}

### 2021-04-14 bem.tex
Proof reading

@@
 \section{BOUNDARY ELEMENT METHODS WITH BEMPP}
-In this section, we provide a brief introduction to boundary element methods and describe the necessary steps for their numerical discretization and solution.
+In this section, we provide a brief introduction to boundary element methods (BEM) and describe the necessary steps for their numerical discretization and solution.
 
@@
 
-Many practical problems have a significantly more complex structure and can involve block systems of integral equations.
-Nevertheless, the fundamental structure of what Bempp-cl does is well described by the above problem and will be described in the following.
+Many practical problems have a significantly more complex structure and can involve block systems of integral equations or even coupling with finite element (FEM) codes.
+Nevertheless, the fundamental structure of what Bempp-cl does is well described using the above simple problem.
 

### 2021-04-14 bempp_overview.tex
start working through opencl section
remove reference to section, as sections are not numbered
Proof reading

@@
-\section{A high level overview of Bempp-cl}
+\section{A HIGH-LEVEL OVERVIEW OF BEMPP-CL}
 
@@
 
-The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles (although for larger problems this cost can be reduced through the use of hierarchical matrices or fast multipole methods).
-Great care is therefore taken to ensure that the operator assembly routines perform as highly as possible.
+The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles.
+For larger problems this cost can be reduced through the use of approximate hierarchical matrices or fast multipole methods (FMM), where the price of the improved complexity is significantly more involved data structures and additional approximation errors.
+Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems.
+Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory.
+Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.
 
-Once a user has defined in operator using Bempp-cl, the discretization can be computed by calling the \pyth{weak_form} method.
+Once a user has defined an operator using Bempp-cl, the discretization can be computed by calling the \pyth{weak_form} method.
 Upon calling this method, a regular integrator will be used to assemble all the interactions between non-adjacent elements, and a singular integrator will be used to compute the interactions between adjacent triangles (if the trial and test spaces are defined on different grids, this second integrator is not needed).
@@
 
-For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), andwhether the discretization should proceed in single or double precision.
+For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision.
 The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.
@@
 If the support of basis functions is larger than a single triangle, different threads may need to sum into the same global degree of freedom.
-We discuss this in more detail in \cref{sec:coloring}.
 
@@
 This is just-in-time compiled via Numba and then the product with the corresponding basis functions is integrated in each triangle via numerical quadrature, again via Numba accelerated routines.
-  \item Computing sparse matrices.
+  \item Computing sparse matrix entries, such as for mass matrices that are required to translate between representations of grid functions through basis coefficients or projections, or when we want to evaluate operator products.
 \end{itemize}
 
-As the cost of each of these processes is in general much smaller than the cost of operator discretization, these can be permformed using Numba without any need to consider the use of OpenCL for potential further speed up.
+As the cost of each of these processes is in general much smaller than the cost of operator discretization, these can be performed using Numba without any need to consider the use of OpenCL for potential further speed up.

### 2021-04-14 intro.tex
Proof reading

@@
 \chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
-The original release came with a simple Python wrapper to the C++ library.
+The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}.
 Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
@@
 The original BEM++ was troubled by what is often called the two language problem.
-It is common for programming language to be either easy for humans to write (e.g., Python) or easy for computers to run and acheive high performance (e.g., C++).
+It is common for programming language to be either easy for humans to write (e.g., Python) or easy for computers to run and achieve high performance (e.g., C++).
 It is not common, however, for a language to do both of these.
@@
 This was the model followed by BEM++, but the problem would perhaps better be described as a three language problem: as well as the Python and C++ code contained in libraries like this, it is common to also include a significant amount of code in a third interfacing language.
-In the case of BEM++, this third language was Cython, an extension of Python with C data types that can be compiled to include functionality from C++ libraries.
+In the case of BEM++, this third language was first Swig and later Cython, an extension of Python with C data types that can be compiled to include functionality from C++ libraries.
 
-Due to the three languages involved, making changes to the library would often mean having to duplicate changed in three places, with many class structures duplicated in all three languages.
-This made seemingly simple changes in onorous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.
+Due to the three languages involved, making changes to the library would often mean having to duplicate changes in three places, with many class structures duplicated in all three languages.
+This made seemingly simple changes an onerous task, and provided a barrier to new members of the community looking to contribute to the open-source project.
 
@@
 In addition to the performance benefits, the library redesign has greatly improved the issues related to the three language problem.
-Both OpenCL and Numba are use to compile functions.
-Each function is provided all the data is needs as inputs, so there is no need to duplicate any class structure outside Python.
+Both OpenCL and Numba are used to compile functions.
+Each function is provided all the data it needs as inputs, so there is no need to duplicate any class structure outside Python.
 The need for a third interfacing language is removed, as the interfacing to the OpenCL kernels is handled by PyOpenCL.
@@
 Boundary element methods are particularly suited to this library model, as the main performance critical task is the computation of discrete operators.
-Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great decrease in performance.
+Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any significant decrease in performance.
 
@@
 Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.
+We conclude with some thoughts on the advantages and disadvantages of OpenCL and Numba.

### 2021-04-14 main.tex
First draft finished.
Proof reading

@@
 \input{numba}
+\input{performance}
 
@@
 OpenCL, \url{https://www.khronos.org/opencl/}.
+
+\bibitem{bempp_maxwell}
+M. W. Scroggs, T. Betcke, E. Burman, Wojciech \'{S}migaj, and E. van 't Wout, ``Software frameworks for integral equations in electromagnetic scattering based on Cader\'{o}n identities''
+{\it Computers \& Mathematics with Applications}, vol. 74, no. 11, pp. 2897--2914, 2017.
+{\textsc{doi}: \href{https://doi.org/10.1016/j.camwa.2017.07.049}{10.1016/j.camwa.2017.07.049}.
+
+\bibitem{bempp_orig}
+W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
+``Solving boundary integral problems with BEM++,''
+{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, Article 6, 2015.
+\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}
+
+\bibitem{bempp_exafmm}
+T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba, ``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-Exafmm''
+{\it arXiv preprint} arXiv:103.01048, 2021.
 \end{thebibliography}

### 2021-04-14 opencl.tex
Proof reading

@@
 When we are running on a CPU and want to take advantage of available SIMD optimizations, we need to make a few modifications to our approach.
-The corresponding kernel works similarly to described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes).
+The corresponding kernel works similarly to what is described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes).
 This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
@@
 
-Data races in global assembly routines are a problem whenever different elements need to sum into the same global degree of freedom.
+Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.
 To solve this we use standard color coding techniques to split up the computations into chunks that access different data regions.
-To this end, we define two elements as neighbors if they share at least one global degree of freedom.
+To this end, we define two triangles as neighbors if they share at least one global degree of freedom.
 Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.

### 2021-04-14 performance.tex
Proof reading

@@
 In this section we want to provide a number of performance benchmarks.
-The tests are running on an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz.
+The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB Ram.
+Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz.
 The CPU supports AVX, AVX2, and AVX-512.
-The tests are running on a Dell Precision 7740 Workstation Laptop with 64GB RAM.
 As GPU we use an Nvidia RTX 3000 GPU.
 All benchmark tests were performed in Linux.
-For OpenCL on the GPU we use the Nvidia GPU drivers and as CPU runtime we compare the open-source POCL driver against the Intel CPU OpenCL runtime environment.
+For OpenCL on the GPU we use the Nvidia GPU drivers and as CPU runtime we compare the open-source Pocl driver against the Intel CPU OpenCL runtime environment.
 All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT Compiler did not skew the results.
@@
 $$
-\dmat{A}_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\pi |\bx -\by|}(\bx)\ds[\by]\ds[\bx]\\
+\dmat{A}_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\pi |\bx -\by|}\ds[\by]\ds[\bx]\\
 $$
@@
 
-We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the POCL OpenCL runtime driver.
+We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the Pocl OpenCL runtime driver.
 We run the tests in single precision and double precision.
-The native vector width for both drivers in single precision is $8$, and and double precision $4$, which corresponds to AVX instructions.
+The native vector width for both drivers in single precision is $8$, and in double precision $4$, which corresponds to AVX instructions.
 In Bempp-cl this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
@@
 	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp.pdf}
-	\caption{Comparison of the performance of POCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
+	\caption{Comparison of the performance of Pocl and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
 	\label{fig:intel_pocl_laplace_cmp}
@@
 We see the expected speed-up between single precision and double precision evaluation.
-What is interesting to note is the difference between the Intel and the POCL runtime environment.
+What is interesting to note is the difference between the Intel and the Pocl runtime environment.
 For P0 basis functions the Pocl driver (black bar) significantly outperforms the Intel driver (blue bar) in both, single and double precision.
@@
 
-In Figure \ref{fig:pocl_single_layer} we compare specifically the performance of single precision and double precision evaluation for the POCL driver for various grid sizes in the case of a P0 basis.
+In Figure \ref{fig:pocl_single_layer} we compare specifically the performance of single precision and double precision evaluation for the Pocl driver for various grid sizes in the case of a P0 basis.
 We can see that the speed-up for larger grid sizes is even slightly faster than just a factor two.
@@
 	        \hline
-	 POCL   &   0.05s       &    0.08s\\
+	 Pocl   &   0.05s       &    0.08s\\
 	 Numba  &   0.25s       &    0.25s\\
@@
 GPU assembly in single precision is a factor two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision).
-Numba is five times slower for single precision and still around three times slower for double precision than POCL.
+Numba is five times slower for single precision and still around three times slower for double precision than Pocl.
 The GPU behaviour is due to the data transfer.
@@
 The main issue, we think, is significantly less usage of AVX vectorized instructions.
-We have taken care to optimize the Green's function evaluation for vectorized evaluation.
+We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba.
 But the loops over integration points and other operations auto-vectorize very badly by just looping over all trial triangle for a single test triangle, as we currently do.
@@
 We won't go into the details of the corresponding OpenCL kernels here but want to show some results that demonstrate the performance of CPU vs GPU.
-While for the dense assembly of boundary integral operators the bus transfer of the dense matrix limited performance, here we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$ and then back to the host the values at the points $x$.
+While for the dense assembly of boundary integral operators the bus transfer of the dense matrix limited performance, here we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$ and then back to the host the values at the points $\bx$.
 
-For this section we used the evaluation of the electric potential operator, defined by
+For this section we use the evaluation of the electric potential operator, defined by
 \begin{align*}
-\mathcal{E\mathbf{p}}(\bx) &= ik\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\nonumber\\
+\left(\mathcal{E\mathbf{p}}\right)(\bx) &= ik\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\nonumber\\
 &-\frac{1}{ik}\nabla_{\bx}\int_{\Gamma}\text{div}\,\mathbf{p}(\by)g(\bx, \by)\ds[\by]
 \end{align*}
-with $g(\bx, \by) = \frac{e^{ik|\bx - \by|}}{4\pi|\bx - \by}$ the Helmholtz Green's function, and
+with $g(\bx, \by) = \frac{e^{ik|\bx - \by|}}{4\pi|\bx - \by|}$ the Helmholtz Green's function, and
 $k$ the wavenumber of the problem.
 The function $\mathbf{p}(\by)\in\mathbb{R}^3$ is a vectorial function, leading to an overall vectorial solution at each point $\bx$.
-For details about the implementation of electromagnetic problems in Bempp see XXX.
-Here, we again use a grid with 32,768 elements and so-called (Rao, Wilton, and Glisson) edge based basis functions.
+For details about the implementation of electromagnetic problems in Bempp see \cite{bempp_maxwell}.
+Here, we again use a grid with 32,768 elements and so-called RWG (Rao, Wilton, and Glisson) edge based basis functions.
 We use 50,000 random evaluation points in the exterior of the unit sphere.
@@
 
-In Figure \ref{fig:efield_domain_potential} we compare the performance of GPU evaluation with that of the POCL driver.
+In Figure \ref{fig:efield_domain_potential} we compare the performance of GPU evaluation with that of the Pocl CPU driver.
 In single-precision the GPU significantly outperforms the CPU, and even in double-precision the Quadro RTX is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
@@
 
-With Bempp-cl we have created a Python platform that achieves high-performance through use of modern JIT technologies.
+With Bempp-cl we have created a Python platform that achieves high-performance through use of modern just-in-time compilation technologies.
 Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
 Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate into other applications with Python interfaces.
-An example is the recent work of integrating Bempp-cl and the Exafmm library for electrostatic virus-scale simulations [XXX].
+An example is the recent work of integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
 Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends.
@@
 It makes explicit SIMD operations very easy through dedicated constructs.
-Moreoever, OpenCL kernels are completely stack based functions, allowing much better optimisations while Numba needs to create every object dynamically, even very small arrays for objects such as coordinate vectors.
+Moreoever, OpenCL kernels are completely stack based functions, allowing much better compiler optimisations while Numba needs to create every object dynamically, even very small arrays for objects such as coordinate vectors.
 
-Also, one has to consider for this approach the type of operations that are required to be accelerated.
+Also, one has to consider the type of operations that are required to be accelerated.
 For the dense assembly of integral operators we have very simple data structures that can easily be passed to compute kernels.
@@
 
-The current focus of further developments is on MPI implementations for cluster computing via \verb@mpi4py@.
+The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \verb@mpi4py@ MPI bindings for Python.
 We have also made big steps forward for large problems by creating a black-box FMM interface that is currently implemented for Exafmm, which has allowed us to solve problems with 10 million elements on a single workstation.
-This shows that Python focused development (with some native routines in lower-level languages) is a scalable model and we are aiming to exploit this scalable further as we move from single workstation computations to large cluster problems.
+This shows that Python focused development (potentially with some native routines in lower-level languages) is a scalable model and we are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.
+
+\section{Acknowledgements}
+The work of Timo Betcke is supported by Engineering and Physical Sciences Research Council Grant EP/V001531/1.

### 2021-04-14 img/colouring_triangles.tex
add colouring diagram

\draw[fill=col1] (0.03522373721510963,-0.40803916808742735) -- (0.36044208773959274,0.13283689310102198) -- (-0.283541653228761,0.14673508308776784) -- cycle;
\fill[white,fill opacity=0.65] (0.03737472390864712,-0.042822397299545845) circle (1pt);
\draw[fill=col2] (0.36044208773959274,0.13283689310102198) -- (0.01759915424630132,0.686970401916699) -- (-0.283541653228761,0.14673508308776784) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.03149986291904435,0.3221807927018296) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.03149986291904435,0.3221807927018296) circle (1pt);\end{scope}
\draw[fill=col3] (0.03522373721510963,-0.40803916808742735) -- (-0.283541653228761,0.14673508308776784) -- (-0.6036260907137695,-0.3766161010593425) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.28398133557580696,-0.21264006201966734) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-0.28398133557580696,-0.21264006201966734) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.28398133557580696,-0.21264006201966734) circle (1pt);\end{scope}
\draw[fill=col4] (0.03522373721510963,-0.40803916808742735) -- (0.6403684745446391,-0.38579170221426184) -- (0.36044208773959274,0.13283689310102198) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.34534476649978046,-0.22033132573355574) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.34534476649978046,-0.22033132573355574) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.34534476649978046,-0.22033132573355574) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.34534476649978046,-0.22033132573355574) circle (1pt);\end{scope}
\draw[fill=col4] (-0.283541653228761,0.14673508308776784) -- (0.01759915424630132,0.686970401916699) -- (-0.617476093555047,0.7029236883312638) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.2944728641791689,0.5122097244452436) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.2944728641791689,0.5122097244452436) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.2944728641791689,0.5122097244452436) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.2944728641791689,0.5122097244452436) circle (1pt);\end{scope}
\draw[fill=col5] (-0.6036260907137695,-0.3766161010593425) -- (-0.283541653228761,0.14673508308776784) -- (-0.9219481904610344,0.1627856205506708) -- cycle;
\fill[white,fill opacity=0.65] (-0.6030386448011883,-0.022365132473634624) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6030386448011883,-0.022365132473634624) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6030386448011883,-0.022365132473634624) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6030386448011883,-0.022365132473634624) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6030386448011883,-0.022365132473634624) circle (1pt);\end{scope}
\draw[fill=col2] (0.03522373721510963,-0.40803916808742735) -- (-0.6036260907137695,-0.3766161010593425) -- (-0.3052595746679132,-0.9153944499673305) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.291220642722191,-0.5666832397047001) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.291220642722191,-0.5666832397047001) circle (1pt);\end{scope}
\draw[fill=col3] (0.6676122663150503,0.7464322913822609) -- (0.01759915424630132,0.686970401916699) -- (0.36044208773959274,0.13283689310102198) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3485511694336481,0.5220798621333272) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (0.3485511694336481,0.5220798621333272) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3485511694336481,0.5220798621333272) circle (1pt);\end{scope}
\draw[fill=col5] (0.36044208773959274,0.13283689310102198) -- (0.6403684745446391,-0.38579170221426184) -- (0.9483026476725608,0.187832500997976) -- cycle;
\fill[white,fill opacity=0.65] (0.6497044033189309,-0.02170743603842129) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6497044033189309,-0.02170743603842129) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6497044033189309,-0.02170743603842129) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6497044033189309,-0.02170743603842129) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6497044033189309,-0.02170743603842129) circle (1pt);\end{scope}
\draw[fill=col6] (0.32742755947762947,-0.9192184282730179) -- (0.6403684745446391,-0.38579170221426184) -- (0.03522373721510963,-0.40803916808742735) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (0.3343399237457927,-0.5710164328582358) circle (1pt);\end{scope}
\draw[fill=col6] (-0.9219481904610344,0.1627856205506708) -- (-0.283541653228761,0.14673508308776784) -- (-0.617476093555047,0.7029236883312638) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.6076553124149474,0.3374814639899008) circle (1pt);\end{scope}
\draw[fill=col5] (0.32742755947762947,-0.9192184282730179) -- (0.03522373721510963,-0.40803916808742735) -- (-0.3052595746679132,-0.9153944499673305) -- cycle;
\fill[white,fill opacity=0.65] (0.019130574008275298,-0.7475506821092587) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.019130574008275298,-0.7475506821092587) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.019130574008275298,-0.7475506821092587) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.019130574008275298,-0.7475506821092587) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.019130574008275298,-0.7475506821092587) circle (1pt);\end{scope}
\draw[fill=col6] (0.6676122663150503,0.7464322913822609) -- (0.36044208773959274,0.13283689310102198) -- (0.9483026476725608,0.187832500997976) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (0.6587856672424013,0.3557005618270863) circle (1pt);\end{scope}
\draw[fill=col1] (0.01759915424630132,0.686970401916699) -- (-0.32059734486270186,1.1911423443882259) -- (-0.617476093555047,0.7029236883312638) -- cycle;
\fill[white,fill opacity=0.65] (-0.3068247613904825,0.8603454782120629) circle (1pt);
\draw[fill=col1] (-0.6036260907137695,-0.3766161010593425) -- (-0.9219481904610344,0.1627856205506708) -- (-1.1766044663247621,-0.37681628548939927) -- cycle;
\fill[white,fill opacity=0.65] (-0.9007262491665221,-0.19688225533269033) circle (1pt);
\draw[fill=col5] (0.2723649506759587,1.2015357806708773) -- (0.01759915424630132,0.686970401916699) -- (0.6676122663150503,0.7464322913822609) -- cycle;
\fill[white,fill opacity=0.65] (0.3191921237457701,0.8783128246566124) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3191921237457701,0.8783128246566124) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3191921237457701,0.8783128246566124) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3191921237457701,0.8783128246566124) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3191921237457701,0.8783128246566124) circle (1pt);\end{scope}
\draw[fill=col4] (-0.6036260907137695,-0.3766161010593425) -- (-0.8999338933384341,-0.8661822216119329) -- (-0.3052595746679132,-0.9153944499673305) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6029398529067056,-0.7193975908795354) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6029398529067056,-0.7193975908795354) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6029398529067056,-0.7193975908795354) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6029398529067056,-0.7193975908795354) circle (1pt);\end{scope}
\draw[fill=col1] (0.6403684745446391,-0.38579170221426184) -- (1.2002626395553422,-0.32694233121632177) -- (0.9483026476725608,0.187832500997976) -- cycle;
\fill[white,fill opacity=0.65] (0.929644587257514,-0.17496717747753587) circle (1pt);
\draw[fill=col2] (0.9029440075044141,-0.8554271671726799) -- (0.6403684745446391,-0.38579170221426184) -- (0.32742755947762947,-0.9192184282730179) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6235800138422275,-0.7201457658866532) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6235800138422275,-0.7201457658866532) circle (1pt);\end{scope}
\draw[fill=col6] (0.2723649506759587,1.2015357806708773) -- (-0.32059734486270186,1.1911423443882259) -- (0.01759915424630132,0.686970401916699) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.01021107998014729,1.026549508991934) circle (1pt);\end{scope}
\draw[fill=col6] (-1.1766044663247621,-0.37681628548939927) -- (-0.8999338933384341,-0.8661822216119329) -- (-0.6036260907137695,-0.3766161010593425) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.8933881501256553,-0.5398715360535582) circle (1pt);\end{scope}
\draw[fill=col3] (0.9029440075044141,-0.8554271671726799) -- (1.2002626395553422,-0.32694233121632177) -- (0.6403684745446391,-0.38579170221426184) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.9145250405347984,-0.5227204002010878) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (0.9145250405347984,-0.5227204002010878) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.9145250405347984,-0.5227204002010878) circle (1pt);\end{scope}
\draw[fill=col2] (-1.2247089422269528,0.7071067860159023) -- (-0.9219481904610344,0.1627856205506708) -- (-0.617476093555047,0.7029236883312638) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.9213777420810114,0.5242720316326123) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.9213777420810114,0.5242720316326123) circle (1pt);\end{scope}
\draw[fill=col1] (5.842836880987079e-09,-1.4142135623730951) -- (0.32742755947762947,-0.9192184282730179) -- (-0.3052595746679132,-0.9153944499673305) -- cycle;
\fill[white,fill opacity=0.65] (0.007389330217517716,-1.082942146871148) circle (1pt);
\draw[fill=col2] (1.2247089422269528,0.7071067860159023) -- (0.6676122663150503,0.7464322913822609) -- (0.9483026476725608,0.187832500997976) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.946874618738188,0.5471238594653797) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.946874618738188,0.5471238594653797) circle (1pt);\end{scope}
\draw[fill=col3] (-0.617476093555047,0.7029236883312638) -- (-0.32059734486270186,1.1911423443882259) -- (-0.8659999976476163,1.2320508104953682) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6013578120217883,1.042038947738286) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-0.6013578120217883,1.042038947738286) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6013578120217883,1.042038947738286) circle (1pt);\end{scope}
\draw[fill=col3] (-1.1766044663247621,-0.37681628548939927) -- (-0.9219481904610344,0.1627856205506708) -- (-1.4999559980990318,0.13397459945181844) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.1995028849616094,-0.026685355162303342) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-1.1995028849616094,-0.026685355162303342) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.1995028849616094,-0.026685355162303342) circle (1pt);\end{scope}
\draw[fill=col4] (0.9483026476725608,0.187832500997976) -- (1.2002626395553422,-0.32694233121632177) -- (1.4999559980990318,0.13397459945181844) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.216173761775645,-0.0017117435888424442) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.216173761775645,-0.0017117435888424442) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.216173761775645,-0.0017117435888424442) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.216173761775645,-0.0017117435888424442) circle (1pt);\end{scope}
\draw[fill=col3] (-0.3052595746679132,-0.9153944499673305) -- (-0.8999338933384341,-0.8661822216119329) -- (-0.6339559956388627,-1.3660254043592894) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.61304982121507,-1.0492006919795178) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-0.61304982121507,-1.0492006919795178) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.61304982121507,-1.0492006919795178) circle (1pt);\end{scope}
\draw[fill=col4] (0.9029440075044141,-0.8554271671726799) -- (0.32742755947762947,-0.9192184282730179) -- (0.6339560031270253,-1.3660254032008328) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6214425233696895,-1.0468903328821768) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6214425233696895,-1.0468903328821768) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6214425233696895,-1.0468903328821768) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6214425233696895,-1.0468903328821768) circle (1pt);\end{scope}
\draw[fill=col1] (0.2723649506759587,1.2015357806708773) -- (0.6676122663150503,0.7464322913822609) -- (0.8659999976476163,1.2320508104953682) -- cycle;
\fill[white,fill opacity=0.65] (0.6019924048795418,1.0600062941828354) circle (1pt);
\draw[fill=col5] (-1.2247089422269528,0.7071067860159023) -- (-0.617476093555047,0.7029236883312638) -- (-0.8659999976476163,1.2320508104953682) -- cycle;
\fill[white,fill opacity=0.65] (-0.9027283444765386,0.8806937616141782) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.9027283444765386,0.8806937616141782) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.9027283444765386,0.8806937616141782) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.9027283444765386,0.8806937616141782) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.9027283444765386,0.8806937616141782) circle (1pt);\end{scope}
\draw[fill=col4] (-1.4999559980990318,0.13397459945181844) -- (-0.9219481904610344,0.1627856205506708) -- (-1.2247089422269528,0.7071067860159023) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.2155377102623397,0.3346223353394639) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.2155377102623397,0.3346223353394639) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.2155377102623397,0.3346223353394639) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.2155377102623397,0.3346223353394639) circle (1pt);\end{scope}
\draw[fill=col6] (5.842836880987079e-09,-1.4142135623730951) -- (-0.3052595746679132,-0.9153944499673305) -- (-0.6339559956388627,-1.3660254043592894) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-0.313071854821313,-1.2318778055665718) circle (1pt);\end{scope}
\draw[fill=col3] (1.2247089422269528,0.7071067860159023) -- (0.9483026476725608,0.187832500997976) -- (1.4999559980990318,0.13397459945181844) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.2243225293328484,0.34297129548856553) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (1.2243225293328484,0.34297129548856553) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.2243225293328484,0.34297129548856553) circle (1pt);\end{scope}
\draw[fill=col3] (0.6339560031270253,-1.3660254032008328) -- (0.32742755947762947,-0.9192184282730179) -- (5.842836880987079e-09,-1.4142135623730951) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.3204611894824972,-1.2331524646156486) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (0.3204611894824972,-1.2331524646156486) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.3204611894824972,-1.2331524646156486) circle (1pt);\end{scope}
\draw[fill=col4] (0.8659999976476163,1.2320508104953682) -- (0.6676122663150503,0.7464322913822609) -- (1.2247089422269528,0.7071067860159023) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.9194404020632065,0.8951966292978438) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.9194404020632065,0.8951966292978438) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.9194404020632065,0.8951966292978438) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.9194404020632065,0.8951966292978438) circle (1pt);\end{scope}
\draw[fill=col2] (0.2723649506759587,1.2015357806708773) -- (0.006491410749332627,1.577780961043636) -- (-0.32059734486270186,1.1911423443882259) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.013913661145803521,1.3234863620342463) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.013913661145803521,1.3234863620342463) circle (1pt);\end{scope}
\draw[fill=col2] (-1.1766044663247621,-0.37681628548939927) -- (-1.3596833219976368,-0.7979501947311224) -- (-0.8999338933384341,-0.8661822216119329) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.1454072272202778,-0.6803162339441515) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.1454072272202778,-0.6803162339441515) circle (1pt);\end{scope}
\draw[fill=col5] (0.9029440075044141,-0.8554271671726799) -- (1.3836131459129406,-0.7602490131863429) -- (1.2002626395553422,-0.32694233121632177) -- cycle;
\fill[white,fill opacity=0.65] (1.1622732643242324,-0.6475395038584482) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1622732643242324,-0.6475395038584482) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1622732643242324,-0.6475395038584482) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1622732643242324,-0.6475395038584482) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1622732643242324,-0.6475395038584482) circle (1pt);\end{scope}
\draw[fill=col3] (0.4482745847469731,1.6730326086910265) -- (0.2723649506759587,1.2015357806708773) -- (0.8659999976476163,1.2320508104953682) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.5288798443568493,1.3688730666190907) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (0.5288798443568493,1.3688730666190907) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.5288798443568493,1.3688730666190907) circle (1pt);\end{scope}
\draw[fill=col4] (-0.8659999976476163,1.2320508104953682) -- (-0.32059734486270186,1.1911423443882259) -- (-0.4482745847469731,1.6730326086910265) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.5449573090857638,1.365408587858207) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.5449573090857638,1.365408587858207) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.5449573090857638,1.365408587858207) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.5449573090857638,1.365408587858207) circle (1pt);\end{scope}
\draw[fill=col5] (-1.6729835308248417,-0.44828773457972093) -- (-1.1766044663247621,-0.37681628548939927) -- (-1.4999559980990318,0.13397459945181844) -- cycle;
\fill[white,fill opacity=0.65] (-1.449847998416212,-0.2303764735391006) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.449847998416212,-0.2303764735391006) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.449847998416212,-0.2303764735391006) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.449847998416212,-0.2303764735391006) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.449847998416212,-0.2303764735391006) circle (1pt);\end{scope}
\draw[fill=col1] (1.2247089467157768,-1.2247448708246802) -- (0.9029440075044141,-0.8554271671726799) -- (0.6339560031270253,-1.3660254032008328) -- cycle;
\fill[white,fill opacity=0.65] (0.9205363191157386,-1.1487324803993977) circle (1pt);
\draw[fill=col2] (1.4999559980990318,0.13397459945181844) -- (1.2002626395553422,-0.32694233121632177) -- (1.6729835308248417,-0.44828773457972093) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4577340561597385,-0.21375182211474142) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4577340561597385,-0.21375182211474142) circle (1pt);\end{scope}
\draw[fill=col1] (-0.6339559956388627,-1.3660254043592894) -- (-0.8999338933384341,-0.8661822216119329) -- (-1.2247089435189604,-1.2247448718903169) -- cycle;
\fill[white,fill opacity=0.65] (-0.9195329441654191,-1.1523174992871796) circle (1pt);
\draw[fill=col4] (0.4482745847469731,1.6730326086910265) -- (0.006491410749332627,1.577780961043636) -- (0.2723649506759587,1.2015357806708773) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.24237698205742145,1.48411645013518) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.24237698205742145,1.48411645013518) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.24237698205742145,1.48411645013518) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.24237698205742145,1.48411645013518) circle (1pt);\end{scope}
\draw[fill=col5] (-0.32059734486270186,1.1911423443882259) -- (0.006491410749332627,1.577780961043636) -- (-0.4482745847469731,1.6730326086910265) -- cycle;
\fill[white,fill opacity=0.65] (-0.2541268396201141,1.4806519713742963) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.2541268396201141,1.4806519713742963) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.2541268396201141,1.4806519713742963) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.2541268396201141,1.4806519713742963) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.2541268396201141,1.4806519713742963) circle (1pt);\end{scope}
\draw[fill=col4] (-1.6729835308248417,-0.44828773457972093) -- (-1.3596833219976368,-0.7979501947311224) -- (-1.1766044663247621,-0.37681628548939927) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4030904397157469,-0.5410180716000809) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4030904397157469,-0.5410180716000809) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4030904397157469,-0.5410180716000809) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4030904397157469,-0.5410180716000809) circle (1pt);\end{scope}
\draw[fill=col6] (1.2247089467157768,-1.2247448708246802) -- (1.3836131459129406,-0.7602490131863429) -- (0.9029440075044141,-0.8554271671726799) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.1704220333777104,-0.9468070170612343) circle (1pt);\end{scope}
\draw[fill=col7] (1.2002626395553422,-0.32694233121632177) -- (1.3836131459129406,-0.7602490131863429) -- (1.6729835308248417,-0.44828773457972093) -- cycle;
\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.4189531054310416,-0.5118263596607951) circle (1pt);\end{scope}
\draw[fill=col5] (-0.8999338933384341,-0.8661822216119329) -- (-1.3596833219976368,-0.7979501947311224) -- (-1.2247089435189604,-1.2247448718903169) -- cycle;
\fill[white,fill opacity=0.65] (-1.1614420529516771,-0.9629590960777907) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.1614420529516771,-0.9629590960777907) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.1614420529516771,-0.9629590960777907) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.1614420529516771,-0.9629590960777907) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.1614420529516771,-0.9629590960777907) circle (1pt);\end{scope}
\draw[fill=col1] (1.4999559980990318,0.13397459945181844) -- (1.7244819355588035,0.6435009302712996) -- (1.2247089422269528,0.7071067860159023) -- cycle;
\fill[white,fill opacity=0.65] (1.483048958628263,0.49486077191300676) circle (1pt);
\draw[fill=col5] (1.2247089422269528,0.7071067860159023) -- (1.4019120630812156,1.1964328621053317) -- (0.8659999976476163,1.2320508104953682) -- cycle;
\fill[white,fill opacity=0.65] (1.1642070009852616,1.0451968195388674) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1642070009852616,1.0451968195388674) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1642070009852616,1.0451968195388674) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1642070009852616,1.0451968195388674) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1642070009852616,1.0451968195388674) circle (1pt);\end{scope}
\draw[fill=col2] (-0.6339559956388627,-1.3660254043592894) -- (-0.3120738877317131,-1.8242063273717561) -- (5.842836880987079e-09,-1.4142135623730951) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.3153432925092463,-1.5348150980347135) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.3153432925092463,-1.5348150980347135) circle (1pt);\end{scope}
\draw[fill=col1] (-0.8659999976476163,1.2320508104953682) -- (-1.4337988048201304,1.1717130557144757) -- (-1.2247089422269528,0.7071067860159023) -- cycle;
\fill[white,fill opacity=0.65] (-1.1748359148982332,1.0369568840752488) circle (1pt);
\draw[fill=col5] (5.842836880987079e-09,-1.4142135623730951) -- (0.30132891135355866,-1.8297405547125698) -- (0.6339560031270253,-1.3660254032008328) -- cycle;
\fill[white,fill opacity=0.65] (0.31176164010780694,-1.5366598400954992) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.31176164010780694,-1.5366598400954992) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.31176164010780694,-1.5366598400954992) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.31176164010780694,-1.5366598400954992) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.31176164010780694,-1.5366598400954992) circle (1pt);\end{scope}
\draw[fill=col6] (-1.2247089422269528,0.7071067860159023) -- (-1.776217984883259,0.6644369119418589) -- (-1.4999559980990318,0.13397459945181844) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.500294308403081,0.50183943246986) circle (1pt);\end{scope}
\draw[fill=col5] (1.6729835308248417,-0.44828773457972093) -- (1.8335853359395162,0.15791651270858353) -- (1.4999559980990318,0.13397459945181844) -- cycle;
\fill[white,fill opacity=0.65] (1.66884162162113,-0.05213220747310632) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.66884162162113,-0.05213220747310632) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.66884162162113,-0.05213220747310632) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.66884162162113,-0.05213220747310632) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.66884162162113,-0.05213220747310632) circle (1pt);\end{scope}
\draw[fill=col6] (1.7244819355588035,0.6435009302712996) -- (1.4019120630812156,1.1964328621053317) -- (1.2247089422269528,0.7071067860159023) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.4503676469556572,0.8490135261308445) circle (1pt);\end{scope}
\draw[fill=col4] (-0.3120738877317131,-1.8242063273717561) -- (0.30132891135355866,-1.8297405547125698) -- (5.842836880987079e-09,-1.4142135623730951) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.003581656845105844,-1.6893868148191402) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.003581656845105844,-1.6893868148191402) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.003581656845105844,-1.6893868148191402) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.003581656845105844,-1.6893868148191402) circle (1pt);\end{scope}
\draw[fill=col1] (-1.4999559980990318,0.13397459945181844) -- (-1.9114100886604803,0.09754555687332445) -- (-1.6729835308248417,-0.44828773457972093) -- cycle;
\fill[white,fill opacity=0.65] (-1.6947832058614514,-0.07225585941819268) circle (1pt);
\draw[fill=col4] (-1.2247089435189604,-1.2247448718903169) -- (-0.901561078474667,-1.6980737352629498) -- (-0.6339559956388627,-1.3660254043592894) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.92007533921083,-1.4296146705041854) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.92007533921083,-1.4296146705041854) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.92007533921083,-1.4296146705041854) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.92007533921083,-1.4296146705041854) circle (1pt);\end{scope}
\draw[fill=col2] (0.8659999976476163,1.2320508104953682) -- (1.0052430855783654,1.652055275477378) -- (0.4482745847469731,1.6730326086910265) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.7731725559909849,1.519046231554591) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.7731725559909849,1.519046231554591) circle (1pt);\end{scope}
\draw[fill=col7] (1.4999559980990318,0.13397459945181844) -- (1.8335853359395162,0.15791651270858353) -- (1.7244819355588035,0.6435009302712996) -- cycle;
\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.6860077565324507,0.31179734747723387) circle (1pt);\end{scope}
\draw[fill=col3] (-1.2247089422269528,0.7071067860159023) -- (-1.4337988048201304,1.1717130557144757) -- (-1.776217984883259,0.6644369119418589) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4782419106434475,0.847752251224079) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-1.4782419106434475,0.847752251224079) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4782419106434475,0.847752251224079) circle (1pt);\end{scope}
\draw[fill=col2] (0.6339560031270253,-1.3660254032008328) -- (0.9132374716234245,-1.723800491533479) -- (1.2247089467157768,-1.2247448708246802) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.9239674738220754,-1.4381902551863306) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.9239674738220754,-1.4381902551863306) circle (1pt);\end{scope}
\draw[fill=col2] (-0.4482745847469731,1.6730326086910265) -- (-1.036192489880894,1.6528102962492381) -- (-0.8659999976476163,1.2320508104953682) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.7834890240918279,1.519297905145211) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.7834890240918279,1.519297905145211) circle (1pt);\end{scope}
\draw[fill=col5] (-0.6339559956388627,-1.3660254043592894) -- (-0.901561078474667,-1.6980737352629498) -- (-0.3120738877317131,-1.8242063273717561) -- cycle;
\fill[white,fill opacity=0.65] (-0.6158636539484142,-1.6294351556646651) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6158636539484142,-1.6294351556646651) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6158636539484142,-1.6294351556646651) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-0.6158636539484142,-1.6294351556646651) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-0.6158636539484142,-1.6294351556646651) circle (1pt);\end{scope}
\draw[fill=col7] (0.8659999976476163,1.2320508104953682) -- (1.4019120630812156,1.1964328621053317) -- (1.0052430855783654,1.652055275477378) -- cycle;
\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.0910517154357324,1.3601796493593594) circle (1pt);\end{scope}
\draw[fill=col3] (-1.2247089435189604,-1.2247448718903169) -- (-1.3596833219976368,-0.7979501947311224) -- (-1.732,-1.0) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.438797421838866,-1.0075650222071464) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-1.438797421838866,-1.0075650222071464) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.438797421838866,-1.0075650222071464) circle (1pt);\end{scope}
\draw[fill=col6] (-1.732,-1.0) -- (-1.3596833219976368,-0.7979501947311224) -- (-1.6729835308248417,-0.44828773457972093) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.5882222842741596,-0.7487459764369477) circle (1pt);\end{scope}
\draw[fill=col1] (1.6729835308248417,-0.44828773457972093) -- (1.3836131459129406,-0.7602490131863429) -- (1.732,-1.0) -- cycle;
\fill[white,fill opacity=0.65] (1.5961988922459274,-0.7361789159220212) circle (1pt);
\draw[fill=col3] (1.732,-1.0) -- (1.3836131459129406,-0.7602490131863429) -- (1.2247089467157768,-1.2247448708246802) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4467740308762391,-0.9949979613370076) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (1.4467740308762391,-0.9949979613370076) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4467740308762391,-0.9949979613370076) circle (1pt);\end{scope}
\draw[fill=col6] (-1.036192489880894,1.6528102962492381) -- (-1.4337988048201304,1.1717130557144757) -- (-0.8659999976476163,1.2320508104953682) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.111997097449547,1.3521913874863607) circle (1pt);\end{scope}
\draw[fill=col6] (0.6339560031270253,-1.3660254032008328) -- (0.30132891135355866,-1.8297405547125698) -- (0.9132374716234245,-1.723800491533479) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (0.6161741287013361,-1.6398554831489605) circle (1pt);\end{scope}
\draw[fill=col2] (-1.776217984883259,0.6644369119418589) -- (-1.9114100886604803,0.09754555687332445) -- (-1.4999559980990318,0.13397459945181844) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.7291946905475903,0.2986523560890006) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.7291946905475903,0.2986523560890006) circle (1pt);\end{scope}
\draw[fill=col2] (-1.2247089435189604,-1.2247448718903169) -- (-1.3618422790127012,-1.5005158798690261) -- (-0.901561078474667,-1.6980737352629498) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.1627041003354428,-1.474444829007431) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.1627041003354428,-1.474444829007431) circle (1pt);\end{scope}
\draw[fill=col4] (0.9132374716234245,-1.723800491533479) -- (1.3841875442815668,-1.5157210596388775) -- (1.2247089467157768,-1.2247448708246802) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1740446542069227,-1.4880888073323455) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1740446542069227,-1.4880888073323455) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.1740446542069227,-1.4880888073323455) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.1740446542069227,-1.4880888073323455) circle (1pt);\end{scope}
\draw[fill=col7] (-1.732,-1.0) -- (-1.3618422790127012,-1.5005158798690261) -- (-1.2247089435189604,-1.2247448718903169) -- cycle;
\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.4395170741772205,-1.241753583919781) circle (1pt);\end{scope}
\draw[fill=col5] (1.2247089467157768,-1.2247448708246802) -- (1.3841875442815668,-1.5157210596388775) -- (1.732,-1.0) -- cycle;
\fill[white,fill opacity=0.65] (1.4469654969991146,-1.246821976821186) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4469654969991146,-1.246821976821186) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4469654969991146,-1.246821976821186) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4469654969991146,-1.246821976821186) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4469654969991146,-1.246821976821186) circle (1pt);\end{scope}
\draw[fill=col2] (1.7244819355588035,0.6435009302712996) -- (1.8486907406397926,1.0813196020269937) -- (1.4019120630812156,1.1964328621053317) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.6583615797599371,0.973751131467875) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.6583615797599371,0.973751131467875) circle (1pt);\end{scope}
\draw[fill=col3] (1.4019120630812156,1.1964328621053317) -- (1.4742937693835385,1.5937503271738678) -- (1.0052430855783654,1.652055275477378) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.2938163060143733,1.4807461549188592) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (1.2938163060143733,1.4807461549188592) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.2938163060143733,1.4807461549188592) circle (1pt);\end{scope}
\draw[fill=col4] (-1.4337988048201304,1.1717130557144757) -- (-1.903696259741285,1.0889518583943874) -- (-1.776217984883259,0.6644369119418589) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.7045710164815582,0.9750339420169073) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.7045710164815582,0.9750339420169073) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.7045710164815582,0.9750339420169073) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.7045710164815582,0.9750339420169073) circle (1pt);\end{scope}
\draw[fill=col5] (-1.5186506450853736,1.5852744704976554) -- (-1.4337988048201304,1.1717130557144757) -- (-1.036192489880894,1.6528102962492381) -- cycle;
\fill[white,fill opacity=0.65] (-1.3295473132621327,1.469932607487123) circle (1pt);
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.3295473132621327,1.469932607487123) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.3295473132621327,1.469932607487123) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.3295473132621327,1.469932607487123) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.3295473132621327,1.469932607487123) circle (1pt);\end{scope}
\draw[fill=col1] (1.8486907406397926,1.0813196020269937) -- (1.4742937693835385,1.5937503271738678) -- (1.4019120630812156,1.1964328621053317) -- cycle;
\fill[white,fill opacity=0.65] (1.5749655243681822,1.2905009304353978) circle (1pt);
\draw[fill=col4] (-1.6729835314920616,-1.4835639151527893) -- (-1.3618422790127012,-1.5005158798690261) -- (-1.732,-1.0) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.588941936834921,-1.3280265983406052) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.588941936834921,-1.3280265983406052) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.588941936834921,-1.3280265983406052) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.588941936834921,-1.3280265983406052) circle (1pt);\end{scope}
\draw[fill=col2] (-1.5186506450853736,1.5852744704976554) -- (-1.903696259741285,1.0889518583943874) -- (-1.4337988048201304,1.1717130557144757) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.6187152365489297,1.2819797948688394) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.6187152365489297,1.2819797948688394) circle (1pt);\end{scope}
\draw[fill=col6] (1.732,-1.0) -- (1.3841875442815668,-1.5157210596388775) -- (1.6729835314920616,-1.4835639151527893) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (1.5963903585912094,-1.3330949915972223) circle (1pt);\end{scope}
\draw[fill=col3] (-1.185537765755746,-1.8617424377913965) -- (-1.3618422790127012,-1.5005158798690261) -- (-1.6729835314920616,-1.4835639151527893) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4067878587535028,-1.6152740776044041) circle (1pt);\end{scope}
\fill[white,fill opacity=0.65] (-1.4067878587535028,-1.6152740776044041) circle (1pt);
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4067878587535028,-1.6152740776044041) circle (1pt);\end{scope}
\draw[fill=col2] (1.6729835314920616,-1.4835639151527893) -- (1.3841875442815668,-1.5157210596388775) -- (1.1910008946773103,-1.8839989784629743) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.4160573234836462,-1.627761317751547) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.4160573234836462,-1.627761317751547) circle (1pt);\end{scope}
\draw[fill=col4] (1.865497566342666,1.4310098979201906) -- (1.4742937693835385,1.5937503271738678) -- (1.8486907406397926,1.0813196020269937) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.7294940254553322,1.3686932757070174) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (1.7294940254553322,1.3686932757070174) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (1.7294940254553322,1.3686932757070174) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (1.7294940254553322,1.3686932757070174) circle (1pt);\end{scope}
\draw[fill=col6] (-1.4999560007354338,-1.8660254018200062) -- (-1.185537765755746,-1.8617424377913965) -- (-1.6729835314920616,-1.4835639151527893) -- cycle;
\begin{scope}[shift={(2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,2pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,-2pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}
\begin{scope}[shift={(2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}
\begin{scope}[shift={(-2pt,0pt)}]\fill[white,fill opacity=0.65] (-1.4528257659944137,-1.7371105849213972) circle (1pt);\end{scope}

### 2021-04-14 opencl.tex
add colouring diagram

@@
 
+\begin{figure}
+	\center
+	\input{img/colouring}
+	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighbouring triangles are the same color.
+Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors.
+[The dots on each cell are included as a visual aid for anyone who prints this article in black and white.]}
+	\label{fig:colouring}
+\end{figure}
+
 Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.

### 2021-04-14 img/sphere.py

@@ -39,3 +39,71 @@ with open("triangles_command.tex", "w") as f:
 
             f.write(f"\\draw ({p1[0]},{p1[1]}) -- ({p2[0]},{p2[1]}) -- ({p3[0]},{p3[1]}) -- cycle;\n")
     f.write("}")
+
+
+with open("colouring_triangles.tex", "w") as f:
+    triangles = [(sum(g.vertices.T[i] for i in t) / 3, t) for t in g.elements.T]
+
+    triangles.sort(key=lambda x: x[0][0] + x[0][2] - x[0][1])
+    triangles = [(i, j) for i, j in triangles if i[0] + i[2] - i[1] > 0.6][::-1]
+
+    at_vertex = {}
+    numbers = []
+    for i, j in triangles:
+        banned = []
+        for v in j:
+            if v not in at_vertex:
+                at_vertex[v] = []
+            banned += at_vertex[v]
+        n = min([i for i in range(1, 50) if i not in banned])
+        numbers.append(n)
+        for v in j:
+            at_vertex[v].append(n)
+
+    for n, (_, (i, j, k)) in zip(numbers, triangles):
+        if _[0] + _[2] - _[1] > 0.6:
+            p1 = to2d(*g.vertices.T[i])
+            p2 = to2d(*g.vertices.T[j])
+            p3 = to2d(*g.vertices.T[k])
+            if -1.5 > max(p1[0], p2[0], p3[0]) or 1.5 < min(p1[0], p2[0], p3[0]):
+                continue
+            if -1.5 > max(p1[1], p2[1], p3[1]) or 1.5 < min(p1[1], p2[1], p3[1]):
+                continue
+
+            f.write(f"\\draw[fill=col{n}] ({p1[0]},{p1[1]}) -- ({p2[0]},{p2[1]}) -- ({p3[0]},{p3[1]}) -- cycle;\n")
+            mid = ((p1[0] + p2[0] + p3[0]) / 3, (p1[1] + p2[1] + p3[1]) / 3)
+            if n == 1:
+                f.write(f"\\fill[white,fill opacity=0.65] ({mid[0]},{mid[1]}) circle (1pt);\n")
+            if n == 2:
[…]

### 2021-04-14 numba.tex
Created benchmark figures
up to summary

@@
 \section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}
+The main focus of Numba within Bempp-cl is to provide accelerated implementations of routines with linear complexity, such as grid iterations, integration of functions over grids, or assembly of certain sparse matrices.
+However, we also provide a fall-back implementation of the OpenCL dense operator assembly in Numba.
 
-The main focus of Numba within Bempp-cl is to provide accelerated implementations for linear complexity routines, such as grid iterations, integration of functions over grids or assembly of certain sparse matrices.
-However, we do also provide a fall-back implementation of the OpenCL dense operators in Numba.
-Here, we use classic loop parallelism.
-Each loop iteration is the assembly of one test element with all trial elements.
+With Numba, we use classic loop parallelism: each loop iteration is the assembly of one test element with all trial elements.
 Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations.

### 2021-04-14 opencl.tex
up to summary

@@
 
-\subsection{Avoiding data races in the global assembly}
-
 \begin{figure}
@@
 	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighbouring triangles are the same color.
-Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors.
+Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors, so do not share any degrees of freedom.
 [The dots on each cell are included as a visual aid for anyone who prints this article in black and white.]}
@@
 
+\subsection{Avoiding data races in the global assembly}
 Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.
@@
 Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.
-This information is then later used to run the compute kernels color by color.
-OpenCL parallelises over test elements and trial elements.
-Here we therefore have to iterate over the product space of possible color combinations.
-In Numba we only parallelise over test elements.
-Here it is sufficient to iterate over all possible colors in the test space.
+An example coloring is shown in \cref{fig:colouring}.
+
+The compute kernels can then be run color-by-color.
+OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible color combinations.
+In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colors in the test space.

### 2021-04-14 performance.tex
up to summary

@@
 
-In this section we want to provide a number of performance benchmarks.
-The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB Ram.
+In this section, we provide a number of performance benchmarks.
+The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB RAM.
 Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz.
 The CPU supports AVX, AVX2, and AVX-512.
-As GPU we use an Nvidia RTX 3000 GPU.
+As a GPU we use an Nvidia RTX 3000 GPU.
 All benchmark tests were performed in Linux.
-For OpenCL on the GPU we use the Nvidia GPU drivers and as CPU runtime we compare the open-source Pocl driver against the Intel CPU OpenCL runtime environment.
-All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT Compiler did not skew the results.
+For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment.
+All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT compilers did not skew the results.
 Though hardly noticeable by users, for smaller experiments the JIT phase typically takes longer than the actual computation.
@@
 \subsection{Dense Operator assembly}
-
-Let us start with benchmarking the dense operator assembly. We assemble the matrix $\dmat{A}$ defined by
+We start by benchmarking the dense operator assembly. We assemble the matrix $\dmat{A}$ defined by
 $$
-\dmat{A}_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\pi |\bx -\by|}\ds[\by]\ds[\bx]\\
+\dmat{A}_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\uppi |\bx -\by|}\ds[\by]\ds[\bx],
 $$
 with $\Gamma$ being the unit sphere.
-For the basis functions $\phi_j$ and test functions $\psi_i$ we compare two cases, piecewise constant functions for both (P0 functions), or nodal, piecewise linear, and globally continuous functions for both (P1 functions).
-In the P0 case each triangle is associated with just one piecewise constant function.
-In the P1 case each triangle is associated with $3$ linear basis functions, one for each node of the triangle.
+For the basis functions $\phi_j$ and test functions $\psi_i$, we compare two cases: piecewise constant functions for both (P0 functions); and nodal, piecewise linear, and globally continuous functions for both (P1 functions).
+In the P0 case, each triangle is associated with just one piecewise constant function.
+In the P1 case, each triangle is associated with $3$ linear basis functions, one for each node of the triangle.
 
-We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the Pocl OpenCL runtime driver.
+We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the PoCL OpenCL runtime driver.
 We run the tests in single precision and double precision.
-The native vector width for both drivers in single precision is $8$, and in double precision $4$, which corresponds to AVX instructions.
-In Bempp-cl this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
-Within the assembly almost all floating point instructions are manually vectorized to take advantage of this.
+The native vector width for both drivers in single precision is $8$, and in double precision is $4$, which corresponds to AVX instructions.
+In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
+Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this.
 We should hence see up a factor 2 speed-up between single and double precision assembly.
@@
 	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp.pdf}
-	\caption{Comparison of the performance of Pocl and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
+	\caption{Comparison of the performance of PoCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
 	\label{fig:intel_pocl_laplace_cmp}
@@
 
-The results for a grid with 32,768 elements are shown in Figure \ref{fig:intel_pocl_laplace_cmp}.
+The timings for a grid with 32,768 elements are shown in \cref{fig:intel_pocl_laplace_cmp}.
 We see the expected speed-up between single precision and double precision evaluation.
-What is interesting to note is the difference between the Intel and the Pocl runtime environment.
-For P0 basis functions the Pocl driver (black bar) significantly outperforms the Intel driver (blue bar) in both, single and double precision.
-For P1 basis functions the Intel driver gives better performance.
+It is interesting to note the difference between the Intel and the PoCL runtime environment.
+For P0 basis functions, the PoCL driver (gray bar) significantly outperforms the Intel driver (orange bar) in both single and double precision.
+For P1 basis functions, however, the Intel driver gives better performance.
 The kernel source code is the same in both cases and manual vectorization of floating point operations is performed in the same way.
-However, in the P1 case significantly more data needs to be copied between a work-item and the global memory since the result of the interaction of a test triangle with a trial triangle is a local $3\times 3$ matrix containing all interactions between the three test basis functions on a triangle and the three trial basis functions.
+In the P1 case, significantly more data needs to be copied between a work-item and the global memory since the result of the interaction of a test triangle with a trial triangle is a local $3\times 3$ matrix containing all interactions between the three test basis functions on a triangle and the three trial basis functions.
+{\color{red}(?)}
 
-In Figure \ref{fig:pocl_single_layer} we compare specifically the performance of single precision and double precision evaluation for the Pocl driver for various grid sizes in the case of a P0 basis.
-We can see that the speed-up for larger grid sizes is even slightly faster than just a factor two.
-The final data point in this graph corresponds to the grid size used for Figure \ref{fig:intel_pocl_laplace_cmp}.
+In \cref{fig:pocl_single_layer}, we compare specifically the performance of single precision and double precision evaluation for the PoCL driver for various grid sizes in the case of a P0 basis.
+We can see that the speed-up for larger grid sizes is slightly more than just a factor of two.
+The final data point in this graph corresponds to the grid size used for \cref{fig:intel_pocl_laplace_cmp}.
 
@@
 	\includegraphics[width=7cm]{img/pocl_single_layer.pdf}
-	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using Pocl and P0 basis functions.}
+	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using PoCL and P0 basis functions.}
 	\label{fig:pocl_single_layer}
@@
 
-In Bempp-cl we can easily switch between CPU Assembly, GPU Assembly and Numba Assembly.
-In Figure \ref{fig:cpu_gpu_numba_compare} we present a table for a grid size 2048 elements to compare these modes.
 \begin{figure}
@@
 	        \hline
-	 Pocl   &   0.05s       &    0.08s\\
+	 PoCL   &   0.05s       &    0.08s\\
 	 Numba  &   0.25s       &    0.25s\\
-	 GPU    &   0.11s       &    3s\\
+	 GPU    &   0.11s       &    3.00s\\
 \end{tabular}
 \end{center}
-\caption{Comparison of POCL, Numba and GPU Assembly for a grid with 2048 elements.}
+\caption{Comparison of PoCL, Numba and GPU Assembly for a grid with 2048 elements.}
 \label{fig:cpu_gpu_numba_compare}
 \end{figure}
+In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly.
+\Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements.
 The speed differences are striking.
-GPU assembly in single precision is a factor two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision).
-Numba is five times slower for single precision and still around three times slower for double precision than Pocl.
-The GPU behaviour is due to the data transfer.
-We have done numerous experiments with GPU assembly of dense boundary integral operators.
-While the GPU kernels themselves are extremely fast, data transfer over the bus is destroying performance.
-But even for medium sized problems we have to transfer the data back to the main memory as GPU RAM is too limited to keep a number of dense matrices of dimension a few ten thousand on the device.
+GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision).
+Numba is five times slower for single precision and still around three times slower for double precision than PoCL.
+The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance.
+Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with a tens of thousands or rows and columns on the device.
 
-A different solution is to compute not the matrix but the matrix vector product $Ax$ on the device by recomputing all matrix elements in each-matvec on-the-fly but not storing them.
-We have done internal experiments with this and it gives significant speed-ups as compared to CPU evaluation as we now only need to transfer single vectors over the bus.
-But it is not competitive for larger problems compared to accelerated methods, such as Fast Multipole Methods due to the quadratic complexity of direct evaluation of the matvec as compared to linear complexity in the FMM.
-For smaller problem sizes it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver.
-Hence, there is only limited practical relevance for such on-the-fly GPU evaluation of boundary integral operators.
-There is though very significant practical relevance of on-the-fly evaluation for the evaluation of domain potential operators for visualization and post-processing as we will see in the next section.
+An alternative method is to compute the matrix-vector product (matvec) $\dmat{A}\dvec{x}$ on the device without first computing the dense matrix $\dmat{A}$.
+This can be done by recomputing all matrix elements during each matvec calculation on-the-fly and not storing them.
+We have done experiments with this, and observed significant speed-ups compared to CPU evaluation as we now only need to transfer single vectors over the bus.
+For larger problems, however, it is not competitive compared to accelerated methods such as fast multipole methods (FMM), due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM.
+For smaller problems, it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver.
+Hence, we can conclude that there is only limited practical relevance for on-the-fly GPU evaluation of boundary integral operators.
+There are, however, very significant practical advantages of on-the-fly evaluation of the evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.
 
-The Numba performance difference is interesting.
-The main issue, we think, is significantly less usage of AVX vectorized instructions.
-We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba.
-But the loops over integration points and other operations auto-vectorize very badly by just looping over all trial triangle for a single test triangle, as we currently do.
-We could tune our code for better auto-vectorization in Numba.
-However, this would be work with little benefit as we have highly optimized hand-tuned OpenCL kernels already.
-We therefore recommend Numba for operator assembly only as a fallback if no OpenCL runtime is available.
+The Numba performance difference seen in \cref{fig:cpu_gpu_numba_compare} is interesting.
+The main reason for this, we believe, is significantly lower usage of AVX vectorized instructions.
+We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangle for a single test triangle, as we currently do.
+We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already.
+We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available.
 
 \subsection{Evaluating domain potentials for post-processing of electromagnetic problems}
+Once an integral equation is solved, one is usually interested in evaluating the solution not only at the boundary but also at points away from it. To do this, we evaluate the integral
+$$
+f(\bx) = \int_{\Gamma}g(\bx, \by)\phi(\by) \ds[\by]
+$$
+for many points $\bx$ away from the boundary $\Gamma$. For example, if we want to visualize a solution, we take the points $\bx$ to be a regular grid of points.
 
-Once an integral equation is solved one is usually interested in evaluating the solution not only at the boundary but also at points away from it. Hence, we want to evaluate
-$$
-f(\bx) = \int_{\Gamma}g(\bx), \by)\phi(\by) \ds[\by]
-$$
-for many points $\bx$ away from the boundary $\Gamma$.
-For example, if we want to visualize a solution the points $\bx$ would arise from some plotting grid.
-Typical of this operation that we usually only want to do it once or a few times at the end of a calculation.
+Typically, we want to do only a small number of potential evaluations at the end of a calculation.
 Discretising this operation into a dense matrix and then evaluating the dense matrix-vector product is not practical for larger sizes.
-For very large problems with hundreds of thousands of elements we use Fast Multipole or other accelerated approximate methods.
-However, for problem sizes of a few ten thousand elements up to around a hundred thousand elements (depending on the problem at hand), direct evaluation of this integral for every value $\bx$ is highly efficient.
-We won't go into the details of the corresponding OpenCL kernels here but want to show some results that demonstrate the performance of CPU vs GPU.
-While for the dense assembly of boundary integral operators the bus transfer of the dense matrix limited performance, here we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$ and then back to the host the values at the points $\bx$.
+For very large problems with hundreds of thousands of elements, we use FMM or other accelerated approximate methods.
+For moderately sized problems with a few ten thousand elements up to around a hundred thousand elements (depending on the problem at hand), direct evaluation of this integral for every point $\bx$ is highly efficient.
+We won't go into the details of the corresponding OpenCL kernels here, but we show some results that demonstrate the relative performance on CPUs and GPUs.
 
-For this section we use the evaluation of the electric potential operator, defined by
-\begin{align*}
-\left(\mathcal{E\mathbf{p}}\right)(\bx) &= ik\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\nonumber\\
-&-\frac{1}{ik}\nabla_{\bx}\int_{\Gamma}\text{div}\,\mathbf{p}(\by)g(\bx, \by)\ds[\by]
-\end{align*}
-with $g(\bx, \by) = \frac{e^{ik|\bx - \by|}}{4\pi|\bx - \by|}$ the Helmholtz Green's function, and
+For the dense assembly of boundary integral operators, the performance was limited by the bus transfer of the dense matrix.
+For the evaluation of domain potentials, however, we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$, and then transfer back to the host the values at the points $\bx$.
+
+In this section, we consider the evaluation of the electric potential operator, defined by
+\begin{multline}
+\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
+-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
+\end{multline}
+where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\pi|\bx - \by|}$ is the Helmholtz Green's function, and
 $k$ the wavenumber of the problem.
-The function $\mathbf{p}(\by)\in\mathbb{R}^3$ is a vectorial function, leading to an overall vectorial solution at each point $\bx$.
-For details about the implementation of electromagnetic problems in Bempp see \cite{bempp_maxwell}.
-Here, we again use a grid with 32,768 elements and so-called RWG (Rao, Wilton, and Glisson) edge based basis functions.
-We use 50,000 random evaluation points in the exterior of the unit sphere.
-We set $k=1.0$, though the value is not relevant for the performance.
+The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$.
+The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}.
+For these experiments, we again use a grid with 32,768 elements and RWG (Rao, Wilton, and Glisson) edge-based basis functions.
+For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere.
+We set $k=1.0$, although the wavenumber have little to no effect on the performance.
+{\color{red}[If $k$ is very high, surely it becomes relevant?]}
 
-In Figure \ref{fig:efield_domain_potential} we compare the performance of GPU evaluation with that of the Pocl CPU driver.
-In single-precision the GPU significantly outperforms the CPU, and even in double-precision the Quadro RTX is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
+In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver.
+In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Quadro {\color{red}[Quadro?
+Nvidia?]} RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
 
@@
 	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
-	\caption{Evaluation of an electric field potential operator on CPU via POCL vs GPU}
+	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs GPU}
 	\label{fig:efield_domain_potential}

### 2021-04-14 performance.tex
summary

@@
 \section{SUMMARY}
+With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies.
+Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
+Basing development in Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
 
-With Bempp-cl we have created a Python platform that achieves high-performance through use of modern just-in-time compilation technologies.
-Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
-Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate into other applications with Python interfaces.
-An example is the recent work of integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
-Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends.
-We currently support Numba and OpenCL as compute backends.
-Other technologies could be added easily.
-OpenCL itself has proved a valuable choice for the purpose of this library.
-We have all kernels available in CPU and GPU optimized versions and the user can easily move between CPU and GPU based compute devices by a simple parameter change, allowing us to support a wide range of hardware.
-Here, we have demonstrated Nvidia benchmarks.
-But it would have made no difference to use AMD or Intel GPUs.
+Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models.
+OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change.
+In this article, we demonstrated Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs (for example).
 
-A disadvantage of our approach is that using OpenCL kernels is mixed-language development due to the OpenCL kernels written in C99.
-Using Numba throughout would be a much more native Python experience.
-However, while Numba is constantly improving, achieving optimal performance for complex operations is difficult.
-OpenCL shines here.
-It makes explicit SIMD operations very easy through dedicated constructs.
+A disadvantage of our approach is that using OpenCL kernels introduce a second language (C99) to the library.
+Using Numba throughout would give a much more native Python experience, but, while Numba is constantly improving, it is currently difficult to achieve optimal performance for complex operations.
+OpenCL really shines here, as it makes explicit SIMD operations very easy through dedicated constructs.
 Moreoever, OpenCL kernels are completely stack based functions, allowing much better compiler optimisations while Numba needs to create every object dynamically, even very small arrays for objects such as coordinate vectors.
 
-Also, one has to consider the type of operations that are required to be accelerated.
-For the dense assembly of integral operators we have very simple data structures that can easily be passed to compute kernels.
-More complex data structures with larger mixture of data movement operations and computations (e.g.
-tree-based algorithms), are much harder to accelerate since the Python layer imposes limits here on the performance.
+Additionally, one has to consider the type of operations that need to be accelerated.
+For the dense assembly of integral operators, we have very simple data structures that can easily be passed to compute kernels.
+More complex data structures with larger mixture of data movement operations and computations (e.g., tree-based algorithms), are much harder to accelerate since the Python layer imposes limits here on the performance.
 
-Overall, with the model of mixed Python/OpenCL/Numba development we have created a flexible and easy to extend platform for integral equation computations and it was an effort that has payed back in terms of simple development without sacrificing performance.
+Overall, with the model of mixed Python/OpenCL/Numba development, we have created a flexible and easy to extend platform for integral equation computations.
+The initial re-implementation efforts are really paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance.
 Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.
 
-The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \verb@mpi4py@ MPI bindings for Python.
-We have also made big steps forward for large problems by creating a black-box FMM interface that is currently implemented for Exafmm, which has allowed us to solve problems with 10 million elements on a single workstation.
-This shows that Python focused development (potentially with some native routines in lower-level languages) is a scalable model and we are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.
+The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \pyth{mpi4py} MPI bindings for Python.
+We have also made big steps forward for large problems by creating a black-box FMM interface that currently intenfaces to Exafmm: this has allowed us to solve problems with 10 million elements on a single workstation.
+We believe that Python-focused development (with some native routines in lower-level languages) is a scalable model and are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.
 

### 2021-04-14 performance.tex
Small corrections

@@
 The CPU supports AVX, AVX2, and AVX-512.
-As a GPU we use an Nvidia RTX 3000 GPU.
+As a GPU we use an Nvidia Quadro RTX 3000 GPU.
 All benchmark tests were performed in Linux.
@@
 For P1 basis functions, however, the Intel driver gives better performance.
-The kernel source code is the same in both cases and manual vectorization of floating point operations is performed in the same way.
-In the P1 case, significantly more data needs to be copied between a work-item and the global memory since the result of the interaction of a test triangle with a trial triangle is a local $3\times 3$ matrix containing all interactions between the three test basis functions on a triangle and the three trial basis functions.
-{\color{red}(?)}
 
@@
 For these experiments, we again use a grid with 32,768 elements and RWG (Rao, Wilton, and Glisson) edge-based basis functions.
-For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere.
-We set $k=1.0$, although the wavenumber have little to no effect on the performance.
-{\color{red}[If $k$ is very high, surely it becomes relevant?]}
+For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.
 
 In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver.
-In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Quadro {\color{red}[Quadro?
-Nvidia?]} RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
+In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
 

### 2021-04-14 bem.tex
Finished small corrections.

@@
 Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$.
-For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights.
-In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical, corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.
+For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights.
+In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see Figure \ref{fig:triangles}), corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.
 
-The values $\dvec{b}_i$ of the right-hand side vector $b$ are similarly computed through a numerical quadrature rule.
+The values $\dvec{b}_i$ of the right-hand side vector $\dvec{b}$ are similarly computed through a numerical quadrature rule.
 
-In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRES.
+In the final step, \textbf{Bempp-cl solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers.
 The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.
 
-In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp:
+In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp-cl:
 \begin{enumerate}

### 2021-04-14 bempp_overview.tex
Finished small corrections.

@@
 The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles.
-For larger problems this cost can be reduced through the use of approximate hierarchical matrices or fast multipole methods (FMM), where the price of the improved complexity is significantly more involved data structures and additional approximation errors.
+For larger problems this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity.
+The price of the improved complexity is significantly more involved data structures and additional approximation errors.
 Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems.
@@
 
-For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision.
+For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision.
 The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.

### 2021-04-14 intro.tex
Finished small corrections.

@@
 Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
-
-At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl \cite{Bempp-cl}.
-This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use.
+At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl .
+This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}.
 Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp.
@@
 We begin this article by giving an overview of the boundary element method, and looking at how Bempp can be used to implement such problems.
-Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs.
+Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs and Nvidia GPUs.
 We conclude with some thoughts on the advantages and disadvantages of OpenCL and Numba.

### 2021-04-14 opencl.tex
Finished small corrections.

@@
 OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
-At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++.
+At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++, with some restrictions on the allowed operations.
 The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.
@@
 OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel.
-Work items are grouped together into work groups, which share local memory.
-Barrier synchronization is only allowed within a work group.
+Work-items are grouped together into work-groups, which share local memory.
+Barrier synchronization is only allowed within a work-group.
 All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange.
 Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device.
-OpenCL allows kernels to be loaded as strings and compiled on the fly for a given device, making it well suited for launching from high-productivity languages.
+OpenCL allows kernels to be loaded as strings and compiled on-the-fly for a given device, making it well suited for launching from high-productivity languages.
 
-To launch an OpenCL kernel the user must first initialize buffers on the compute device, then copy any relevant data from host to the compute device.
+To launch an OpenCL kernel the user must provide relevant data as buffers, which are transferred from host to the corresponding compute device.
 A kernel string can then be loaded and just-in-time compiled for the device.
@@
 The effect is that a few work-items do work that is discarded.
-However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of interactions is $N^2$.
+However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of triangle interactions is $N^2$.
 Hence, only a tiny fraction of work-items are discarded.
@@
 
-For this assembly, we create one workgroup for each triangle pair.
-Inside this workgroup, we have a number of work-items that evaluate the quadrature rules then sum up the results.
+For this assembly, we create one work-group for each singular triangle pair.
+Inside this work-group, we have a number of work-items that evaluate the quadrature rules then sum up the results.
 Depending on how two triangles are related to each other, different types of singular quadrature rule are needed.

### 2021-04-14 performance.tex
Finished small corrections.

@@
 For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment.
-All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT compilers did not skew the results.
-Though hardly noticeable by users, for smaller experiments the JIT phase typically takes longer than the actual computation.
+All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba just-in-time compilers did not skew the results.
+Though hardly noticeable by users, for smaller experiments the compilation phase typically takes longer than the actual computation.
 
@@
 We run the tests in single precision and double precision.
-The native vector width for both drivers in single precision is $8$, and in double precision is $4$, which corresponds to AVX instructions.
+The native vector width for both drivers in single precision is $8$, and in double precision is $4$, corresponding to AVX instructions.
 In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
 Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this.
-We should hence see up a factor 2 speed-up between single and double precision assembly.
+We should hence see up to a factor 2 speed-up between single and double precision assembly.
 
@@
 The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance.
-Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with a tens of thousands or rows and columns on the device.
+Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with tens of thousands of rows and columns on the device.
 
@@
 We have done experiments with this, and observed significant speed-ups compared to CPU evaluation as we now only need to transfer single vectors over the bus.
-For larger problems, however, it is not competitive compared to accelerated methods such as fast multipole methods (FMM), due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM.
+For larger problems, however, it is not competitive compared to accelerated methods such as FMM, due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM.
 For smaller problems, it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver.
 Hence, we can conclude that there is only limited practical relevance for on-the-fly GPU evaluation of boundary integral operators.
-There are, however, very significant practical advantages of on-the-fly evaluation of the evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.
+There are, however, very significant practical advantages of on-the-fly evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.
 
@@
 The main reason for this, we believe, is significantly lower usage of AVX vectorized instructions.
-We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangle for a single test triangle, as we currently do.
+We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangles for a single test triangle, as we currently do.
 We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already.
@@
 The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}.
-For these experiments, we again use a grid with 32,768 elements and RWG (Rao, Wilton, and Glisson) edge-based basis functions.
+For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao, Wilton, and Glisson) edge-based basis functions.
 For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.
@@
 In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver.
-In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.
+In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double precision operations.
 
@@
 Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
-Basing development in Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
+Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
 
@@
 OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change.
-In this article, we demonstrated Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs (for example).
+In this article, we demonstrate Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs.
 
@@
 Overall, with the model of mixed Python/OpenCL/Numba development, we have created a flexible and easy to extend platform for integral equation computations.
-The initial re-implementation efforts are really paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance.
+The initial re-implementation efforts by abandoning our old C++ code base are paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance.
 Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.

### 2021-04-14 main.tex
add colouring diagram
up to summary
remove template

@@
 \usetikzlibrary{arrows,decorations.pathreplacing}
-
 \usepackage{ifthen}
@@
 \newcommand{\dvec}[1]{\mathbf{#1}}
+
+\newcommand{\ii}{\mathrm{i}}
+\newcommand{\ee}{\mathrm{e}}
 
@@
 {\it Computers \& Mathematics with Applications}, vol. 74, no. 11, pp. 2897--2914, 2017.
-{\textsc{doi}: \href{https://doi.org/10.1016/j.camwa.2017.07.049}{10.1016/j.camwa.2017.07.049}.
+\textsc{doi}: \href{https://doi.org/10.1016/j.camwa.2017.07.049}{10.1016/j.camwa.2017.07.049}.
 
@@
 \end{document}
-
-\section{COPYRIGHT AND CLEARANCE}
-
-All CS Magazine authors must obtain clearance from IEEE Computer Society before submitting the final manuscript.
-The ``\href{https://apps.na.collabserv.com/wikis/home?lang=en-us#!/wiki/W18e544042a85_4b63_915a_1d1ed2cf8338/page/Publication\%20clearance}{Publication Clearance}'' wiki provides details about the procedure.
-Computer Society employees must use the \href{https://mc.manuscriptcentral.com/cs-ieee}{ScholarOne  Manuscripts Clearance} System to obtain publication approval.
-
-\section{SECTIONS}
-
-Sections following the introduction should present your results and findings.
-The body of the paper should be approximately 6,000 words.
-The manuscript should evolve so that each sentence, equation, figure, and table flow smoothly and logically from whatever precedes it.
-Relevant work by others, as well as relevant products from other companies, should be adequately and accurately cited.
-Sufficient support should be provided (or cited) for the assertions made and conclusions drawn.
-
-Headings may be numbered or unnumbered (``1 Introduction'' and ``1.2 Numbered level 2 head''), with no ending punctuation.
-As demonstrated in this document, the initial paragraph after a headingis not indented.
-
-\section{JOURNAL STYLE}
-
-Use American English when writing your paper.
-The serial comma should be used (``a, b, and c'' not ``a, b and c'').
-In American English, periods and commas are within quotation marks, like ``this period.'' Other punctuation is ``outside''!
-The use of technical jargon, slang, and vague or informal English should be avoided.
-Generic technical terms should instead be used.
-
-\subsection{Acronyms and abbreviations}
-
-All acronyms should be defined at first mention in the abstract and in the main text.
-Define in figures, tables, and footnotes only if not defined in the discussion of the figure/table.
-Acronyms consist of capital letters (except where salted with lowercase), but the terms they represent need not be given initial caps unless a proper name is involved (``central processing unit'' [CPU] but ``Fourier transform'' [FT]).
-Use of ``e.g.'' and ``i.e.'' okay, but refrain from using ``etc.'' It is preferable to use these abbreviations only in parentheses (e.g., like this).
-
-Abbreviate units of time (s, min, hr, day, mo, yr) only in virgule constructions (10~$\umu$g/hr) and in artwork; otherwise, spell out, e.g., 10 days, 3 months, 25 minutes.
-Units of measure (Kb, MB, kWh, etc.) should always be abbreviated when used with a numeral.
-If used alone, spell out (``16 MB of RAM'' but ``these values are measured in micrometers'').
-
-\subsection{Numbers}
-
-Spell out numerals that have no unit of measure or time (one, two, $\ldots$ ten), but always use numerals with units of time and measure.
-Some examples are as follows: 11 through 999; 1,000; 10,000; twentieth century; twofold, tenfold, 20-fold; 2 times; 0.2 cm; $p = .001$; 25\%; 10\% to 25\%.
-
-\section{MATH AND EQUATIONS}
-
-Scalar {\it variables} and {\it physical constants} should be italicized, and a bold (non-italics) font should be used for {\bf vectors} and {\bf matrices}.
-Do not italicize subscripts unless they are variables.
-
-Equations should be either display (with a number in parentheses) or inline.
-
-Display equations should be flush left and numbered consecutively, with equation numbers in parentheses and flush right.
-
-Be sure the symbols in your equation have been defined before the equation appears or immediately following. Please refer to ``Equation (1),'' not ``Eq. (1)'' or ``equation (1).''
-
-Punctuate display equations when they are part of the sentence preceding it, as in
-\begin{equation}
-A=\pi r^2.
-\end{equation}
-In addition, if the text following the equation flows logically as a part of the display equation, 
-\begin{equation}
-A=\pi r^2,
-\end{equation}
-use ending punctuation (comma) after the display equation.
-
-\section{LISTS}
-
-Avoid using lists. Instead, use full sentences and flowing paragraphs. If you absolutely must use a list, use them rarely and keep them short:
-\begin{itemize}
-\item {\it Style for bulleted lists}---This is the style that should be used for bulleted lists.
-
-\item {\it Punctuation in lists}---Each item in the list should end with a period, regardless of whether full sentences are used.
-\end{itemize}
-
-\section{GRAPHICAL ABSTRACTS}
-
-This journal accepts graphical abstracts, and they must be peer reviewed, which means the graphical abstract must be submitted with the full paper.
-graphical abstracts and their specifications.
-Please read the additional information provided by \href{http://www.ieee.org/publications_standards/publications/graphical_abstract.pdf}{\underline{IEEE about graphical abstracts}}.
-
-%\begin{figure}
-%\centerline{\includegraphics[width=18.5pc]{fig1.png}}
-%\caption{Note that ``Figure'' is spelled out.
-There is a period after the figure number, followed by one space.
-It is good practice to briefly explain the significance of the figure in the caption.
-(Used, with permission, from [4].)}
-%\end{figure}
-%
-%\begin{figure*}
-%\centerline{\includegraphics[width=26pc]{fig1.png}}
-%\caption{Note that ``Figure'' is spelled out.
-There is a period after the figure number, followed by one space.
-It is good practice to briefly explain the significance of the figure in the caption.
-(Used, with permission, from [4].)}
-%\end{figure*}
-
-\begin{table}
-\caption{Units for magnetic properties.}
-\label{table}
-\small
-\begin{tabular*}{17.5pc}{@{}|p{29pt}|p{63pt}<{\raggedright}|p{80pt}<{\raggedright}|@{}}
-\hline
-Symbol& 
-Quantity& 
-Conversion from Gaussian and  CGS EMU to SI$^{\mathrm{a}}$ \\
-\hline
-$\Phi $& 
-Magnetic flux& 
-1 Mx $\to  10^{-8}$ Wb $= 10^{-8}$ V $\cdot$ s \\
-$B$& 
-Magnetic flux density,   magnetic induction& 
-1 G $\to  10^{-4}$ T $= 10^{-4}$ Wb/m$^{2}$ \\
-$H$& 
-Magnetic field strength& 
-1 Oe $\to  10^{-3}/(4\pi )$ A/m \\
-$m$& 
-Magnetic moment& 
-1 erg/G $=$ 1 emu   $\to 10^{-3}$ A $\cdot$ m$^{2} = 10^{-3}$ J/T \\
-$M$& 
-Magnetization& 
-1 erg/(G $\cdot$ cm$^{3}) =$ 1 emu/cm$^{3}$   $\to 10^{-3}$ A/m \\
-4$\pi M$& 
-Magnetization& 
-1 G $\to  10^{-3}/(4\pi )$ A/m \\
-$\sigma $& 
-Specific magnetization& 
-1 erg/(G $\cdot$ g) $=$ 1 emu/g $\to $ 1 A $\cdot$ m$^{2}$/kg \\
-$j$& 
-Magnetic dipole   moment& 
-1 erg/G $=$ 1 emu   $\to 4\pi \times  10^{-10}$ Wb $\cdot$ m \\
-$J$& 
-Magnetic polarization& 
-1 erg/(G $\cdot$ cm$^{3}) =$ 1 emu/cm$^{3}$  $\to 4\pi \times  10^{-4}$ T \\
-$\chi , \kappa $& 
-Susceptibility& 
-1 $\to  4\pi $ \\
-$\chi_{\rho }$& 
-Mass susceptibility& 
-1 cm$^{3}$/g $\to  4\pi \times  10^{-3}$ m$^{3}$/kg \\
-$\mu $& 
-Permeability& 
-1 $\to  4\pi \times  10^{-7}$ H/m   $= 4\pi \times  10^{-7}$ Wb/(A $\cdot$ m) \\
-$\mu_{r}$& 
-Relative permeability& 
-$\mu \to \mu_{r}$ \\
-$w, W$& 
-Energy density& 
-1 erg/cm$^{3} \to  10^{-1}$ J/m$^{3}$ \\
-$N, D$& 
-Demagnetizing factor& 
-1 $\to  1/(4\pi )$ \\
-\hline
-\multicolumn{3}{@{}p{17.5pc}@{}}{Vertical lines are optional in tables. Statements that serve as captions for 
-the entire table do not need footnote letters. }\\
-\multicolumn{3}{@{}p{17.5pc}@{}}{$^{\mathrm{a}}$Gaussian units are the same as cg emu for magnetostatics; Mx 
-$=$ maxwell, G $=$ gauss, Oe $=$ oersted; Wb $=$ weber, V $=$ volt, s $=$ 
-second, T $=$ tesla, m $=$ meter, A $=$ ampere, J $=$ joule, kg $=$ 
-kilogram, H $=$ henry.}
-\end{tabular*}
-\label{tab1}
-\end{table}
-
-\section{FIGURES AND TABLES}
-
-\subsection{In-text callouts for figures and tables}
-
-Figures and tables must be cited in the running text in consecutive order.
-At first mention, the citation should be boldface ({\bf Figure 1}); subsequent mentions should be Roman type (see Figure 1 and {\bf Table 1}).
-{\bf Figure 2} shows an example of a figure spanning across two columns.
-
-Previously published figures or tables require permission to reprint.
-Please obtain permission.
-Then, add the following text to the figure/table caption: ``From [reference no.], with permission,'' or ``Adapted from [reference no.], with permission.'' {\it Carefully} explain each figure in the text.
-Each manuscript should be limited to four figures.\pagebreak
-
-\section{END SECTIONS}
-
-\subsection{Appendices}
-
-If multiple appendices are required, they should labeled ``Appendix A,'' ``Appendix B,'' etc. They appear before the ``Acknowledgment'' or the ``References'' section.
-
-\subsection{Acknowledgment}
-
-The ``Acknowledgment'' (no's) section appears immediately after the conclusion.
-If applicable, this is where you indicate funding for the work.
-The preferred spelling of the word ``acknowledgment'' in American English is without an ``e'' after the ``g.'' Avoid expressions such as ``One of us (S.B.A.) would like to thank $\ldots$ .'' Instead, write ``We thank $\ldots$.'' Sponsor and financial support acknowledgments are included in the acknowledgment section.
-For example: This work was supported in part by the U.S.
-Department of Commerce under Grant BS123456 (sponsor and financial support acknowledgment goes here).
-Researchers that contributed information or assistance to the article should also be acknowledged in this section.
-Also, if corresponding authorship is noted in your paper it will be placed in the acknowledgment section.
-Note that the acknowledgment section is placed at the end of the paper before the reference section.
-
-\subsection{References}
-
-References need not be cited in text. When they are, they appear on the 
-line, in square brackets, inside the punctuation. Multiple references are 
-each numbered with separate brackets. When citing a section in a book, 
-please give the relevant page numbers. In text, refer simply to the 
-reference number. Do not use ``Ref.'' or ``reference'' except at the 
-beginning of a sentence: ``Reference \cite{CC1} shows $\ldots$ .'' Please do not use 
-automatic endnotes in \emph{Word}, rather, type the reference list at the end of the 
-paper using the ``References'' style.
-
-Reference numbers are set flush left and form a column of their own, hanging 
-out beyond the body of the reference. The reference numbers are on the line, 
-end with period. In all references, the given name of the author 
-or editor is abbreviated to the initial only and precedes the last name. Use 
-them all; use \emph{et al.} only if names are not given. Use commas around Jr., 
-Sr., and III in names. Abbreviate conference titles. When citing IEEE 
-transactions, provide the issue number, page range, volume number, year, 
-and/or month if available. When referencing a patent, provide the day and 
-the month of issue, or application. References may not include all 
-information; please obtain and include relevant information. Do not combine 
-references. There must be only one reference with each number. If there is a 
-URL included with the print reference, it can be included at the end of the 
-reference.
-
-Other than books, capitalize only the first word in a paper title, except 
-for proper nouns and element symbols. For papers published in translation 
-journals, please give the English citation first, followed by the original 
-foreign-language citation See the end of this document for formats and 
-examples of common\break references. For a complete discussion of references and 
-their formats, see the IEEE style manual at
-\url{http://www.ieee.org/authortools}.
-
-\section{CONCLUSION}
-
-The manuscript should include a conclusion.
-In this section, summarize what was described in your paper.
-Future directions may also be included in this section.
-Authors are strongly encouraged not to reference multiple figures or tables in the conclusion; these should be referenced in the body of the paper.
-
-\section{ACKNOWLEDGMENT}
-
-We thank A, B, and C. This work was supported in part by a grant from XYZ.
-
-\begin{thebibliography}{1}
-
-\bibitem{AA1}
-G. M. Amdahl, G. A. Blaauw, and F. P. Brooks, ``Architecture of the IBM System/360,'' {\it IBM J. Res. \& Dev}., vol. 8, no. 2, pp. 87--101, 1964. (journal)
-
-\bibitem{BB1}
-IBM Corporation, IBM Knowledge Center - IBM Secure Service Container (Secure Service Container).
-[Online].
-Available: {https://www.ibm.com/support/\break knowledgecenter/en/HW11R/com.ibm.hwmca.kc\_se.doc/\break introductiontotheconsole/wn2131zaci.html} (URL)
-
-\bibitem{CC1}
-J. Williams, ``Narrow-band analyzer,'' PhD dissertation, Dept. of Electrical Eng., Harvard Univ., Cambridge, MA: 1993. (Thesis or dissertation)
-
-\bibitem{DD1}
-J.
-M.
-P.
-Martinez, R.
-B.
-Llavori, M.
-J.
-A.
-Cabo, et al., ``Integrating data warehouses with web data: A survey,'' {\it IEEE Trans.
-Knowledge and Data Eng}., preprint, Dec.
-21, 2007, doi:10.1109/TKDE.2007.190746.
-(PrePrint)
-
-\bibitem{EE1}
-W.-K. Chen, {\it Linear Networks and Systems}, Belmont, CA: Wadsworth, pp. 123--135, 1993. (book)
-
-\bibitem{FF1}
-S.
-P.
-Bingulac, ``On the compatibility of adaptive controllers,'' {\it Proc.
-Fourth Ann.
-Allerton Conf.
-Circuits and Systems Theory}, pp.
-8--16, 1994.
-(conference proceedings)\vadjust{\vfill\pagebreak}
-
-\bibitem{GG1}
-K. Elissa, ``An overview of decision theory,'' unpublished. (Unpublished manuscript)
-
-\bibitem{HH1}
-R. Nicole, ``The last word on decision theory,'' {\it J. Computer Vision}, submitted for publication. (Pending publication)
-
-\bibitem{II1}
-C. J. Smith and J. S. Smith, Rocky Mountain Research Laboratories, Boulder, CO, personal communication, 1992. (Personal communication)
-\end{thebibliography}
-
-\begin{IEEEbiography}{First A. Author}{\,}current employment (is
-currently with XYZ Corporation, New York, NY, USA).
-Author's latest degree received or which he/she is currently pursuing (He received the B.S.
-degree and the M.S.
-degree....").
-Author's research interest.
-Association with any official journals or conferences; major professional and/or academic achievements, i.e., best paper awards, research grants, etc.; any publication information (number of papers and titles of books published).
-Author membership information, e.g., is a member of the IEEE and the IEEE Computer Society, if applicable, is noted at the end of the biography.
-Biography is limited to single paragraph only (He is a member of IEEE, etc.).
-All biographies should be limited to one paragraph, five to six sentences, consisting of the following: (e.g., ``Dr.
-Author received a B.S.
-degree and an M.S.
-degree $\ldots$.''), including years achieved; association with any official journals or conferences; major professional and/or academic achievements, i.e., best paper awards, research grants, etc.; any publication information (number of papers and titles of books published); current research interests; association with any professional associations.
-Author membership information, e.g., is a member of the IEEE and the IEEE Computer Society, if applicable, is noted at the end of the biography.
-Contact him/her at faauthor@xyz.com.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}{Second B. Author, Jr.,}{\,}is with ABC Corporation, Bblingen, Germany. Ms. Author's biography appears here.  Contact him/her at sbauthor@abc.com.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}{Third C. Author, III,}{\,}is with DEF Corporation, Tokyo, Japan. Dr. Author's biography appears here.  Contact him/her at tcauthor@def.com.
-\end{IEEEbiography}

### 2021-04-14 bem.tex
final changes

@@
 Once a triangulation is given we need to define the necessary data structures for the discretization.
-Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these function.
+Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these functions.
 We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients.
@@
 
-In Bempp-cl, an operator definition consists of the type of the operator (e.g. Laplace single-layer in the above example),
+In Bempp-cl, an operator definition consists of the type of the operator (e.g., Laplace single-layer in the above example),
 the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
@@
 For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights.
-In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see Figure \ref{fig:triangles}), corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.
+In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see \cref{fig:triangles}), corresponding singular quadrature rules are used that are based on singularity-removing coordinate transformations.
 

### 2021-04-14 bempp_overview.tex
final changes

@@
 The main user-visible component of Bempp-cl is the module \pyth{bempp.api}, which defines all user interface functions and other high-level routines.
-In particular, it contains the definition of the main object types: \pyth{Grid}, \pyth{Space}, \pyth{GridFunction}, \pyth{Operator}.
+In particular, it contains the definitions of the main object types: \mbox{\pyth{Grid},} \mbox{\pyth{Space},} \mbox{\pyth{GridFunction},} and \pyth{Operator}.
 The computational backend routines are contained in the module \pyth{bempp.core}.
-Currently, we support a Numba and OpenCL backends.
+Currently, we support Numba and OpenCL backends.
 An overview of this structure is provided in \cref{fig:overview}.
@@
 The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles.
-For larger problems this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity.
+For larger problems, this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity.
 The price of the improved complexity is significantly more involved data structures and additional approximation errors.
-Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems.
+Bempp-cl provides interfaces to ExaFMM \cite{bempp_exafmm} for large problems.
 Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory.

### 2021-04-14 intro.tex
final changes

@@
 Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
-At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl .
+At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl.
 This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}.
@@
 Due to the three languages involved, making changes to the library would often mean having to duplicate changes in three places, with many class structures duplicated in all three languages.
-This made seemingly simple changes an onerous task, and provided a barrier to new members of the community looking to contribute to the open-source project.
+This made seemingly simple changes into onerous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.
 

### 2021-04-14 main.tex
final changes

@@
 Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba.
-The OpenCL backend implements kernels for GPUs and SIMD optimized for CPUs.
-In this paper we will discuss the design of Bempp-cl, provide performance comparisons on different compute devices and discuss the advantages and disadvantages of OpenCL as compared to Numba.
+The OpenCL backend implements kernels for GPUs and CPUs with SIMD optimization.
+In this paper, we discuss the design of Bempp-cl, provide performance comparisons on different compute devices, and discuss the advantages and disadvantages of OpenCL as compared to Numba.
 \end{abstract}
@@
 ``Solving boundary integral problems with BEM++,''
-{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, Article 6, 2015.
+{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, article 6, 2015.
 \textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}
@@
 \bibitem{bempp_exafmm}
-T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba, ``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-Exafmm''
-{\it arXiv preprint} arXiv:103.01048, 2021.
+T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba,
+``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-ExaFMM,''
+2021.
+arXiv (preprint): \href{https://arxiv.org/abs/2103.01048}{2103.01048}.
 \end{thebibliography}

### 2021-04-14 opencl.tex
final changes

@@
 Barrier synchronization is only allowed within a work-group.
-All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange.
+All work-items are uniquely indexed by a one, two, or three dimensional index space, called \ocl{NDRange}.
 Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device.
@@
 OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types.
-For example, the type \ocl{double4} will allow four double values to be held in a SIMD register.
+For example, the type \mbox{\ocl{double4}} will allow four double values to be held in a SIMD register.
 This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.
@@
 	\item Signs of the normal directions for the spaces.
-	\item Test and trial grid as flat floating point array, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
+	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
 	\item Test and trial connectivity information that stores the indices of each node for each triangle.
@@
 	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
-	\item Number of test and trial degrees of freedom.
+	\item The number of test and trial degrees of freedom.
 	\item A single byte that is set to one if the test and trial grids are disjoint.
@@
 This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
-If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard-non vectorized kernel.
+If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard non-vectorized kernel.
 
@@
 
-Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work items on SIMD lanes, similar to what we do manually.
+Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work-items on SIMD lanes, similar to what we do manually.
 In our experience, this works well for very simple kernels but often fails for more complex OpenCL kernels.

### 2021-04-14 performance.tex
final changes

@@
 In this section, we provide a number of performance benchmarks.
-The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB RAM.
+The tests were all run on a Dell Precision 7740 Workstation Laptop with 64 GB RAM.
 Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz.
@@
 
-We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the PoCL OpenCL runtime driver.
+We first compare the OpenCL CPU performance for the Intel OpenCL runtime and the PoCL OpenCL runtime driver.
 We run the tests in single precision and double precision.
@@
 \end{center}
-\caption{Comparison of PoCL, Numba and GPU Assembly for a grid with 2048 elements.}
+\caption{Comparison of PoCL, Numba and GPU assembly for a grid with 2048 elements.}
 \label{fig:cpu_gpu_numba_compare}
@@
 \end{multline}
-where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\pi|\bx - \by|}$ is the Helmholtz Green's function, and
+where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\uppi|\bx - \by|}$ is the Helmholtz Green's function, and
 $k$ the wavenumber of the problem.
@@
 The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}.
-For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao, Wilton, and Glisson) edge-based basis functions.
+For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao--Wilton--Glisson) edge-based basis functions are used.
 For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.
@@
 	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
-	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs GPU}
+	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs Nvidia GPU}
 	\label{fig:efield_domain_potential}
@@
 Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
-Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
+Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the ExaFMM library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
 
 Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models.
-OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change.
-In this article, we demonstrate Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs.
+OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU- and GPU-based compute devices with a simple parameter change.
+In this article, we demonstrate Nvidia benchmarks: the same benchmarking code could be used to run on AMD or Intel GPUs.
 
@@
 The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \pyth{mpi4py} MPI bindings for Python.
-We have also made big steps forward for large problems by creating a black-box FMM interface that currently intenfaces to Exafmm: this has allowed us to solve problems with 10 million elements on a single workstation.
+We have also made big steps forward for large problems by creating a black-box FMM interface that currently interfaces to ExaFMM: this has allowed us to solve problems with 10 million elements on a single workstation.
 We believe that Python-focused development (with some native routines in lower-level languages) is a scalable model and are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.

### 2021-05-10 bem.tex
Revisions done.

@@
 Once a triangulation is given we need to define the necessary data structures for the discretization.
-Bempp-cl uses a Galerkin discretization: we introduce set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these functions.
+Bempp-cl uses a Galerkin discretization: we introduce a set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these functions.
 We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients.
@@
 \end{align*}
-In the case of piecewise constant trial and test functions, the definition of $\dmat{A}_{ij}$ simplifies to $\dmat{A}_{ij} = \int_{\Gamma}\int_{\Gamma}g(\bx, \by)\ds[\by]\ds[\bx]$.
+In the case of piecewise constant trial and test functions, the definition of $\dmat{A}_{ij}$ simplifies to $\dmat{A}_{ij} = \int_{\tau_i}\int_{\tau_j}g(\bx, \by)\ds[\by]\ds[\bx]$.
 
@@
 
-Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$.
-For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights.
-In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see \cref{fig:triangles}), corresponding singular quadrature rules are used that are based on singularity-removing coordinate transformations.
+Once the grid, the spaces, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$.
+For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule such as a symmetric Gauss rule for triangles.
+In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see \cref{fig:triangles}), corresponding singular quadrature rules are used that are based on singularity-removing coordinate transformations \cite{erichsen}.
 
@@
 All of these steps are accelerated through the use of either Numba or OpenCL.
-In the following section we provide a high-level overview of the library and how these acceleration techniques are deployed before we deep dive into the design of the computational kernels.
+In the following section we provide a high-level overview of the library and how these acceleration techniques are deployed before we dive into the design of the computational kernels.

### 2021-05-10 bempp_overview.tex
Revisions done.

@@
 Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory.
-Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.
+Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as efficiently as possible.
 
@@
 \begin{itemize}
-  \item Computing the grid topology: this involves iterating through the grid to compute the neighbour relataionships between triangles.
+  \item Computing the grid topology: this involves iterating through the grid to compute the neighbour relationships between triangles.
   \item Definition of local-to-global maps for function spaces: again, this requires traversal through the grid and assigning relationships between global and local indices.

### 2021-05-10 numba.tex
Revisions done.

@@
 
-With Numba, we use classic loop parallelism: each loop iteration is the assembly of one test element with all trial elements.
+With Numba, we use loop parallelism: each loop iteration is the assembly of one test triangle with all trial triangles. We then parallelize over the test triangles.
+All such assemblies are executed through a parallel for-loop.
 Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations.
 However, a much smaller fraction of operations is SIMD optimized due to the lack of targeted SIMD constructs in Numba.
+
+We stress that while Numba provides backends not only for CPU, but also for ROCm and CUDA, we currently only use the CPU component of Numba.

### 2021-05-10 opencl.tex
Revisions done.

@@
 OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
-At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++, with some restrictions on the allowed operations.
+At its core, OpenCL executes compute kernels that can be written in OpenCL C, which is based on C99, or (more recently) in C++, with some restrictions on the allowed operations.
 The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.
@@
 
-To launch an OpenCL kernel the user must provide relevant data as buffers, which are transferred from host to the corresponding compute device.
+To launch an OpenCL kernel the user must provide relevant data as buffers, which are transferred from the host to the corresponding compute device.
 A kernel string can then be loaded and just-in-time compiled for the device.
@@
 	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
-	\item Test and trial connectivity information that stores the indices of each node for each triangle.
+	\item Test and trial connectivity, which are lists of node indices that define the corresponding elements of the test and trial grid.
 	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
-	\item Test and trial basis function multipliers for each triangle.
+	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g. in electromagnetics).
 	\item Quadrature points and quadrature weights.
@@
 Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.
-To solve this we use standard color coding techniques to split up the computations into chunks that access different data regions.
-To this end, we define two triangles as neighbors if they share at least one global degree of freedom.
-Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.
-An example coloring is shown in \cref{fig:colouring}.
+To solve this we use standard colouring techniques to split up the computations into chunks that access different data regions.
+To this end, we define two triangles as neighbours if they share at least one global degree of freedom.
+Based on this relationship we run a simple greedy colouring algorithm in the initialization phase of a function space.
+An example colouring is shown in \cref{fig:colouring}.
 
-The compute kernels can then be run color-by-color.
-OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible color combinations.
-In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colors in the test space.
+The compute kernels can then be run colour-by-colour.
+OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations.
+In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colours in the test space.

### 2021-05-10 performance.tex
Revisions done.

@@
 The speed differences are striking.
-GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision).
+GPU assembly in single precision is a factor of two slower than CPU assembly, and much slower in double precision due to the limited double precision performance of our hardware, though we note that this is not an issue for data center accelerators by Nvidia that are also optimised for double precision.
 Numba is five times slower for single precision and still around three times slower for double precision than PoCL.
@@
 
-A disadvantage of our approach is that using OpenCL kernels introduce a second language (C99) to the library.
+A disadvantage of our approach is that using OpenCL kernels introduces a second language (C99) to the library.
 Using Numba throughout would give a much more native Python experience, but, while Numba is constantly improving, it is currently difficult to achieve optimal performance for complex operations.
 OpenCL really shines here, as it makes explicit SIMD operations very easy through dedicated constructs.
-Moreoever, OpenCL kernels are completely stack based functions, allowing much better compiler optimisations while Numba needs to create every object dynamically, even very small arrays for objects such as coordinate vectors.
+Moreoever, OpenCL kernels are completely stack/register based functions, allowing much better compiler optimisations while Numba needs to create every object dynamically, even for very small arrays for objects such as coordinate vectors.
+We need to stress that we have performed very few optimisations specific to Numba, while significant optimisation has gone into the OpenCL codes.
+It is therefore well possible that the performance gap between Numba and OpenCL can be significantly reduced.
+But from other projects our own anecdotal experience is that the more Numba is optimised, the less Pythonic and more C-like Numba functions look.
+So while Numba is a very powerful tool, it requires its own techniques for optimisation, different from standard Python code.
 
-Additionally, one has to consider the type of operations that need to be accelerated.
+Another important consideration with respect to Python and just-in-time acceleration is the type of algorithms that benefit.
 For the dense assembly of integral operators, we have very simple data structures that can easily be passed to compute kernels.

### 2021-05-11 opencl.tex
american spelling and add Numba reference

@@
 	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
-	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g. in electromagnetics).
+	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g., in electromagnetics).
 	\item Quadrature points and quadrature weights.
@@
 	\input{img/colouring}
-	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighbouring triangles are the same color.
+	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighboring triangles are the same color.
 Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors, so do not share any degrees of freedom.
@@
 Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.
-To solve this we use standard colouring techniques to split up the computations into chunks that access different data regions.
-To this end, we define two triangles as neighbours if they share at least one global degree of freedom.
-Based on this relationship we run a simple greedy colouring algorithm in the initialization phase of a function space.
-An example colouring is shown in \cref{fig:colouring}.
+To solve this we use standard coloring techniques to split up the computations into chunks that access different data regions.
+To this end, we define two triangles as neighbors if they share at least one global degree of freedom.
+Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.
+An example coloring is shown in \cref{fig:colouring}.
 
-The compute kernels can then be run colour-by-colour.
+The compute kernels can then be run color-by-color.
 OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations.

### 2021-05-12 response_to_reviewers.md
Have written reviewer responses

# Responses to the Associate Editor and Reviewers

We are grateful for the careful comments by the reviewers and editor. Below you find responses to questions raised by the associate editior and reviewers. For smaller remarks regarding typos, etc., we have done corrections directly in the text without providing responses.

## Responses to the Associate Editor comments

_The technique described by the authors seems to entail a substantial duplication of code that might be avoidable using certain techniques. What trade-offs led the authors to their approach?_

The main code duplication is within OpenCL kernels. Some of this was mitigated through inline functions defined in headers and significant use of define statements. But this does not avoid all code duplication. At the end, the decision was that this was a price worth paying for the advantage of being able to encapsulate expensive computational routines in fast OpenCL kernels.

The redevelopment of kernels using Numba came after OpenCL. This was mainly driven out of curiosity of how well Numba could perform but also to give users an alternative who don't have a well working OpenCL stack on their system (especially Windows and Mac)

_Modern FEM codes tend to employ atomic operations instead of graph coloring for improved performance. (Atomic FP operations, also in 64 bit, can be realized via the classical compare-and-swap technique.)_

We are using the OpenCL 1.2 standard, which supports a compare and swap operation only for 32 bit floats. Otherwise, I agree that atomics are preferable.

_The authors state that SIMD vectorization is not used for the evaluation of singular integrals, on account of the far field representing the lion's share of the work. Under FMM acceleration, near-field and far-field should take approximately equal time. What were the obstacles to applying the optimization in this setting?_

In the preprint https://arxiv.org/pdf/2103.01048.pdf we discuss the combination of Bempp-cl and Exafmm for large Poisson-Boltzmann problems. This has figures on combined assembly time and solver times. Singular integrals were not a significant bottleneck here. We are currently preparing a more detailed paper about fast solver integration into Bempp, where this will be discussed more. Should for certain problem sizes the singular integrals become a bottleneck, we can either manually optimize or shift onto GPU with not much effort. One note though, we are not manually SIMD optimizing. But I believe that due to the structure of our singular integration routines some automatic vectorization potential may be exploited by the compiler.

_The comment on lower FP64 throughput on Nvidia hardware should be removed or rephrased as it only applies to gaming-targeted consumer hardware._

A lot of our users don't have access to HPC Centre type accelerators. So this is a relevant observation for them. We have clarified in the text though that slow double precision performance is not an issue in data center accelerators.

_The authors describe repeated host-device transfers as a particular performance bottleneck. These are likely avoidable if the data is kept on the device throughout. Why was this route not chosen? This question is particularly salient because it directly influences the authors' disqualification of GPUs for direct evaluation tasks._

The reason for this is problem sizes vs available memory in most accelerators. Dense discretisation require significant amounts of memory. Barring data centre type accelerators most cards don't have enough RAM to keep dense matrices on them for interesting problem sizes. Many practical computations do not just need a single matrix, but a number of matrices (e.g. transmission probelms), quickly eating up RAM on any but the most expensive accelerators.

But this is an option that is likely coming soon as part of our push to optimise Bempp-cl for cluster computing on HPC systems with MPI.

_Numba is described as a fallback technology, but no particular scenario is given in which the fallback would be required. The installation instructions for PyOpenCL (https://documen.tician.de/pyopencl/misc.html) appear to contain instructions to deploy a full OpenCL stack on nearly any machine._

Numba has two uses in our library, one essential, and one fallback. The essential use is for all O(n) routines that involve grid iterations. Here, we make extensive use of Numba. For operator assembly it started as experiment to see how a simple Numba implementaton fares against OpenCL. The decision to keep this in the code was mainly for Mac users. On Mac we frequently encountered problems with users having crashes in POCL, and the Apple CPU OpenCL implementation being problematic (workgroup sizes of more than 1 item caused problems for us on CPU with the Apple OpenCL driver). We do not have Macs available to properly sort out these problems, and often users don't want to use Docker images as it makes it more difficult to modify the code and link with external libraries (it is certainly possible, but most users don't have this Docker knowledge). Hence, Mac users can use this fallback quite easily. On Windows this is only relevant to some extent as the Intel OpenCL CPU driver for Windows works well.

_When the authors characterize OpenCL as a "second language", to what extent is this second language necessitated by the GPU as a "second environment"?_

Good point. Personally, I like how libraries such as Numba-Cuda hide the complexity of the second device. But this certainly comes with its own drawbacks.

_Roofline plots for cost/performance._

A dedicated paper on the dense algorithms for OpenCL is in planning. This will contain much more detailed performance comparisons than possible in the available format and time for the special issue.

## Responses to Reviewer 1

_In the left column on page 6, what are "basis function multipliers" in lines 30-31?_

Certain function spaces such as RWG spaces in Maxwell need additional multipliers that depend on the triangle. These are stored here.

_In the left column on page 8, maybe an example of how "classic loop parallelism" works in line 33 would be useful. I was wondering how assembling the combination of one test element with all trial elements could still lead to linear complexity._

Each test triangle is assembled together with all trial triangles. We then paralleize over the test triangles. This has been clarified in the text.

_In the right column on page 8, why is only AVX used in line 31? As far as I can see, the processor used for the benchmarks supports AVX512, as well._

We have used the default vector width from the OpenCL driver. AVX512 is neither in Intel nor in Pocl enabled by default as it can lead to slower performance than AVX2 due to the significant reduction in the CPU clock rate. Also, on AMD it is not supported at all. We have therefore decided to only present tests with AVX2. In practice, we noticed minor improvements with AVX512.

_In the left column on page 11, I was surprised at the works "stack based functions" in lines 29-30. As far as I know, OpenCL assigned registers directly to variables and does not use a stack. The point still stands, of course, since registers are even more efficient than stack-based local variables, as long as there are enough of them available._

I think for the CPU this depends on the implementation and cannot automatically be assumed.

_I suggest to include references to previous work on BEM with GPUs, e.g., the paper "GCA-H² matrix compression for electrostatic simulations" by S. Börm and S. Christophersen or the papers "Algorithmic patterns for H-matrices on many-core processors" and "A scalable H-matrix approach for the solution of boundary integral equations on multi-GPU clusters" by P. Zaspel._

These are excellent paper of which we are aware. We have not excluded these and other similar papers (e.g. the work by Rio Yokota's group on FMM and H-Matrices on GPUs) since the focus of this publication is on software design for dense matrix assembly with Python and OpenCL. We are working on publications related to the coupling of Bempp-cl and Exafmm, in which these references are more suitable.

_In the left column on page 6, what happens if the global assembled matrix does not fit into, e.g., graphics memory when using a GPU in line 33? Is it possible to split the global matrix into submatrices that fit into graphics memory, and then transfer them to main memory? Given the high computational complexity, would it be possible to overlap memory transfers and computation?_

Right now, in this case a memory exception is created and returned to the Python interpreter. Submatrix assembly and swapping would be possible. However, efficient shadowing of memory transfer by computations requires that the computations are at least as long as the memory transfers. From the experiments we have done with simple Laplace operators this was not the case. Assembly is blazingly fast and memory transfer much slower. This may be different with more complex operators such as those from Maxwell.

The reason why we have not focused further on this is that there is not so much relevance for practical applications. On modern CPUs the assembly for everything that can feasibly done dense is fast enough. For anything larger we then want to use FMM or H/H^2 matrix techniques.

## Responses to Reviewer 2

_We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand- tuned OpenCL kernels already." To me this statement casts a shadow on the interpretation of the benchmarking results since it may not be "fair" to compare highly optimized OpenCL kernels to less well-optimized Numba kernels. I do understand that Numba is meant to be a backup implementation in Bempp, but then the authors should be careful to put these results in context. I ask that they provide some additional discussion on the "fairness" of this comparison and how the results should be interpreted._

This is a fair point and we have emphasised this more in the text. On the last page we have added the following part:

We need to stress that we have performed very few optimisations specific to Numba, while significant optimisation has gone into the OpenCL codes. It is therefore well possible that the performance gap between Numba and OpenCL can be significantly reduced. But from other projects our own anecdotal experience is that the more Numba is optimised, the less Pythonic and more C-like Numba functions look. So while Numba is a very powerful tool, it requires its own techniques for optimisation, different from standard Python code.

_Since Numba provides support for both CUDA and ROCm, the authors should explicitly state that are using Numba CUDA in the abstract and introduction. It's clear from the context since they are using an NVIDIA GPU but they need to say this to avoid confusion._

Just as clarification. We are not using the GPU functionality of Numba. All our Numba code is CPU code. We have added a corresponding comment to the Numba Assembly section. We have considered a Numba Cuda backend. However, currently this would bring no advantage to the existing OpenCL backend for GPUs as they work on Nvidia and ROCm devices.

_Having briefly experimented with PyOpenCL/OpenCL myself, I was surprised at the authors' characterization that using it was easy. I struggled to get it configured properly with my NVIDIA vendor driver, for example. Maybe the authors did not have this experience but since OpenCL has a general reputation for being harder to use than CUDA, it would be useful for the authors to acknowledge this and state whether their experience was similar or different._

There are two things here to distinguish. 1.) Installation difficulties of OpenCL. 2.) Complexity of using OpenCL. With regards to OpenCL installation. I am using a Linux laptop with the Nvidia close source drivers and never had any issues. The only small caveat is that if PyOpenCL is installed from conda, one needs to symlink the Nvidia icd from /etc/OpenCL/vendors to the corresponding /etc/ directory in the virtual environment for PyOpenCL to see it. Also, on Windows I did not encounter issues. Indeed, we have users that have used Bempp on Windows with an Nvidia card.

With regards to the complexity of OpenCL itself, here the big advantages of Python bindings come in. While the OpenCL C library is very verbose, the PyOpenCL bindings automate most of the work and allow to launch kernels with just a few lines of code. Since we only ever used OpenCL from Python the verbosity of the C interface to OpenCL was never an issue for us.

Debugging OpenCL can be more of a problem. But I am very used to debugging kernels with simple printf statements. Over time one gets quite proficient at it.

_This is a more minor comment but since OpenCL was chosen for portability, I would have been very interested to see benchmarking results on AMD GPUs too. If this is future work it is worth mentioning._

We do not have access to modern AMD hardware. I have a workstation with two AMD Polaris generation GPUs. But my Nvidia card is a modern Quadro RTX 3000, which in all our tests significantly outperforms the older AMD Polaris generation. It would therefore be unfair to bring such comparison benchmarks here. We would only demonstrate the age difference between these GPU generations from different vendors.

We are working on a dedicated paper on the OpenCL algorithms which will contain much more detailed benchmarks on different cards with respect to their peak performance. The focus of the current paper was on demonstrating the potential of mixed Python/OpenCL development and less on detailed performance bencmarking of specific algorithms.

## Responses to Reviewer 3

_It needs to introduce the PDEs and range of applications at the beginning better, as well as more welcoming notation for the general reader. I was also hoping for more insight into the results (fraction of peak perf, fraction of time on host-to-device transfers, etc), and take-home messages about mistakes made and lessons learned._

I understand this sentiment. There is a strict page limit for this article. Within this limit we decided to focus on the software design and less on mathematical and application aspects.

With regards to more detailed performance measurements, we are preparing a more dedicated paper on the OpenCL algorithms which will have more detail on this than we have scope in this paper.

With regards to take-home message about lessons learned we again would love to talk much more than is possible to fit on a one page summary at the end. Should the referee wish I'd be happy to discuss this topic much more in person than is possible through the refereeing format. There are a number of aspects of our journey from C++ to Python that I could not discuss here, and the journey is continuing. While the development model that we use in the paper works very well for Bempp-cl, we encountered limits of the Python focused model for other types of algorithms that are worth discussing in their own rights (and we are currently planning a corresponding publication).

### 2021-05-12 opencl.tex
Have written reviewer responses

@@
 \begin{itemize}
-	\item Test and trial indices of triangles over which to be integrated.
+	\item Test and trial indices denoting the triangles over which to be integrated.
 	\item Signs of the normal directions for the spaces.
 	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
-	\item Test and trial connectivity, which are lists of node indices that define the corresponding elements of the test and trial grid.
+	\item Test and trial connectivity, which are lists of node indices that define the corresponding triangles of the test and trial grid.
 	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
@@
 OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations.
-In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colours in the test space.
+In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colors in the test space.

### 2021-05-12 performance.tex
Have written reviewer responses

@@
 The speed differences are striking.
-GPU assembly in single precision is a factor of two slower than CPU assembly, and much slower in double precision due to the limited double precision performance of our hardware, though we note that this is not an issue for data center accelerators by Nvidia that are also optimised for double precision.
+GPU assembly in single precision is a factor of two slower than CPU assembly, and much slower in double precision due to the limited double precision performance of our hardware, though we note that this is not an issue for double precision optimised data center accelerators.
 Numba is five times slower for single precision and still around three times slower for double precision than PoCL.
@@
 We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already.
-We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available.
+We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available (this is less a judgement about Numba itself but about the limited optimisations we have done for Numba assembly routines).
 

### 2021-05-28 main.tex
Revisions done.
american spelling and add Numba reference
Added Timo Betcke biography

@@
 
+\bibitem{erichsen}
+S. Erichsen, and A. Sauter,
+``Efficient automatic quadrature in 3-d Galerkin BEM,''
+{\it Computer Methods in Applied Mechanics and Engineering}, vol. 157, no. 3, pp. 215--224, 1998.
+\textsc{doi}: \href{https://doi.org/10.1016/S0045-7825(97)00236-3}{10.1016/S0045-7825(97)00236-3}.
+
 \bibitem{pyopencl}
@@
 
+\bibitem{numba}
+S. K. Lam, A. Pitrou, and S. Seibert,
+``Numba: a LLVM-based Python JIT compiler,''
+{\it Proceedings of the Second Workshop on the LLVM Compiler Infrastructure in HPC}, 2015.
+\textsc{doi}: \href{https://dx.doi.org/10.1145/2833157.2833162}{10.1145/2833157.2833162}.
+
 \bibitem{opencl}
@@
 \bibitem{bempp_maxwell}
-M. W. Scroggs, T. Betcke, E. Burman, Wojciech \'{S}migaj, and E. van 't Wout, ``Software frameworks for integral equations in electromagnetic scattering based on Cader\'{o}n identities''
+M. W. Scroggs, T. Betcke, E. Burman, Wojciech \'{S}migaj, and E. van 't Wout,
+``Software frameworks for integral equations in electromagnetic scattering based on Cader\'{o}n identities''
 {\it Computers \& Mathematics with Applications}, vol. 74, no. 11, pp. 2897--2914, 2017.
@@
 
+\begin{IEEEbiography}{Timo Betcke}{\,}is Professor of Computational Mathematics in the Department of Mathematics at University College London (UCL).
+He originally studied Computational Engineering at Hamburg University of Technology before pursuing a DPhil in Numerical Analysis
+at the University of Oxford, from which he graduated in 2006. Before joining UCL in 2011 he has held research positions at Braunschweig,
+Manchester, and Reading. His work is on the interface of numerical analysis and computational sciences. He was principal investigator
+in two EPSRC software infrastructure grants for the development of Bempp and has been PI and Co-I on a number of other projects related
+to boundary element methods.
+\end{IEEEbiography}
+
 \end{document}

### 2021-05-28 main.tex
add bio

@@
 He originally studied Computational Engineering at Hamburg University of Technology before pursuing a DPhil in Numerical Analysis
-at the University of Oxford, from which he graduated in 2006. Before joining UCL in 2011 he has held research positions at Braunschweig,
+at the University of Oxford, from which he graduated in 2006. Before joining UCL in 2011 he held research positions at Braunschweig,
 Manchester, and Reading. His work is on the interface of numerical analysis and computational sciences. He was principal investigator
@@
 
+\begin{IEEEbiography}{Matthew W. Scroggs}{\,}is a postdoctoral research associate in the Department of Engineering at the
+University of Cambridge, where he works on finite element methods and the open source finite element library FEniCSx.
+Before moving to Cambridge, he obtained a PhD from the Department of Mathematics at University College London, where his
+work focused on boundary element methods and Bempp.
+\end{IEEEbiography}
+
 \end{document}

### 2021-05-28 main.tex
Updated references and merged files

@@
 
-\input{intro}
-\input{bem}
-\input{bempp_overview}
-\input{opencl}
-\input{numba}
-\input{performance}
+\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations.
+The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}.
+Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
+At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl.
+This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}.
+Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp.
+In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.
+
+\subsection{The three language problem}
+The original BEM++ was troubled by what is often called the two language problem.
+It is common for programming language to be either easy for humans to write (e.g., Python) or easy for computers to run and achieve high performance (e.g., C++).
+It is not common, however, for a language to do both of these.
+Due to this, it is common in scientific computing libraries to write a library in a fast low-level language such as C++, while providing a user interface in a higher-level language such as Python.
+
+This was the model followed by BEM++, but the problem would perhaps better be described as a three language problem: as well as the Python and C++ code contained in libraries like this, it is common to also include a significant amount of code in a third interfacing language.
+In the case of BEM++, this third language was first Swig and later Cython, an extension of Python with C data types that can be compiled to include functionality from C++ libraries.
+
+Due to the three languages involved, making changes to the library would often mean having to duplicate changes in three places, with many class structures duplicated in all three languages.
+This made seemingly simple changes into onerous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.
+
+\subsection{Delegating computations with PyOpenCL}
+Prompted by a desire to simplify the library as well as to be able to run on a wide range of CPU and GPU devices, we began a full re-write in 2018, which led to Bempp-cl.
+The aims of this rewrite were to support explicit single-instruction-multiple-data (SIMD) optimization on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python.
+These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.
+
+In addition to the performance benefits, the library redesign has greatly improved the issues related to the three language problem.
+Both OpenCL and Numba are used to compile functions.
+Each function is provided all the data it needs as inputs, so there is no need to duplicate any class structure outside Python.
+The need for a third interfacing language is removed, as the interfacing to the OpenCL kernels is handled by PyOpenCL.
+
+Boundary element methods are particularly suited to this library model, as the main performance critical task is the computation of discrete operators.
+Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any significant decrease in performance.
+
+\bigskip
+
+We begin this article by giving an overview of the boundary element method, and looking at how Bempp can be used to implement such problems.
+Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs and Nvidia GPUs.
+We conclude with some thoughts on the advantages and disadvantages of OpenCL and Numba.
+
+\section{BOUNDARY ELEMENT METHODS WITH BEMPP}
+In this section, we provide a brief introduction to boundary element methods (BEM) and describe the necessary steps for their numerical discretization and solution.
+
+The most simple boundary integral equation is of the form
+\begin{align}
+	\label{eq:bnd_integral}
+	\int_{\Gamma} g(\bx, \by)\phi(\by)\ds[\by] &= f(\bx),&&\bx\in\Gamma.
+\end{align}
+The function $g(\bx, \by)$ is a Green's function, $f$ is a given right-hand side, and $\phi$ is an unknown surface density over the boundary $\Gamma$ of a bounded three dimensional domain $\Omega\subset\mathbb{R}^3$.
+
+As a concrete example, we consider computing the electrostatic capacity of an object $\Omega$.
+In this case, we solve the above equation with $f(\bx)=1$ and $g(\bx,\by)=\frac{1}{4\uppi|\bx - \by|}$.
+Once $\phi$ has been found, the normalized capacity is then obtained using $c = \frac{1}{4\uppi}\int_{\Gamma}\phi(\bx)\ds[\bx]$.
+
+Many practical problems have a significantly more complex structure and can involve block systems of integral equations or even coupling with finite element (FEM) codes.
+Nevertheless, the fundamental structure of what Bempp-cl does is well described using the above simple problem.
+
+\begin{figure}
+	\centering
+	\input{img/sphere}
+	\caption{Discretization of a sphere into flat surface triangles.}
+	\label{fig:triangulation}
+\end{figure}
+
+\begin{figure*}
+	\centering
+	\input{img/triangles}
+	\caption{The four types of intersection of two triangles: the triangles can be (left to right) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle.
+In the first case, standard quadrature is used.
+In the other three cases, singular quadrature rules must be used.}
+	\label{fig:triangles}
+\end{figure*}
+
+The first step is the \textbf{discretization of the surface} $\Gamma$.
+Surfaces are represented in Bempp-cl as a triangulation into flat triangles (see \cref{fig:triangulation}).
+The triangulation is internally represented as an array of node coordinates and an associated connectivity array of node indices that define each triangle.
+In this step, topology data is also computed: in particular, for each triangle, we compute the neighboring triangles and the type of intersection (i.e.
+are they connected by an edge or a vertex, see \cref{fig:triangles}).
+
+Once a triangulation is given we need to define the necessary data structures for the discretization.
+Bempp-cl uses a Galerkin discretization: we introduce a set of basis functions \(\phi_1\) to \(\phi_N\), and define the \textbf{trial space} as the span of these functions.
+We then approximate the solution \(\phi\) of \cref{eq:bnd_integral} by $\phi_h=\sum_{j=1}^N \dvec{x}_j\phi_j$, where $\dvec{x}$ is a vector of coefficients.
+In the most simple case, we can define the function \(\phi_j\) to be equal to 1 on the triangle $\tau_j$ and 0 everywhere else.
+Other spaces are commonly defined to be piecewise polynomials on each triangle.
+
+To discretize \cref{eq:bnd_integral}, we define a \textbf{test (or dual) space} in terms of a basis \(\psi_1\) to \(\psi_N\). The discrete representation of the above problem then takes the form
+$$
+\dmat{A}\dvec{x} = \dvec{b}
+$$
+with
+\begin{align*}
+	\dmat{A}_{ij} &= \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}g(\bx, \by)\phi_j(\by)\ds[\by]\ds[\bx]\\
+	\dvec{b}_i &= \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx].
+\end{align*}
+In the case of piecewise constant trial and test functions, the definition of $\dmat{A}_{ij}$ simplifies to $\dmat{A}_{ij} = \int_{\tau_i}\int_{\tau_j}g(\bx, \by)\ds[\by]\ds[\bx]$.
+
+In Bempp-cl, an operator definition consists of the type of the operator (e.g., Laplace single-layer in the above example),
+the definition of the trial and test spaces, and the definition of the range space. The range space is required for operator
+products and not relevant for the purpose of this paper. The function $f$ is represented as a \textbf{grid function} object, which
+consists of either the dual representation in the forms of the integrals $\dvec{b}_i = \int_{\Gamma}\psi_i(\bx)f(\bx)\ds[\bx]$ or directly through its coefficients $f_j$ in the representation $f=\sum_{j=1}^N f_j\phi_j$.
+
+Once the grid, the spaces, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$.
+For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule such as a symmetric Gauss rule for triangles.
+In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see \cref{fig:triangles}), corresponding singular quadrature rules are used that are based on singularity-removing coordinate transformations \cite{erichsen}.
+
+The values $\dvec{b}_i$ of the right-hand side vector $\dvec{b}$ are similarly computed through a numerical quadrature rule.
+
+In the final step, \textbf{Bempp-cl solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers.
+The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.
+
+In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp-cl:
+\begin{enumerate}
+	\item Import of the surface description as triangulation data.
+	\item Definition of function spaces and relevant operators.
+	\item Discretization into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
+	\item Solution of the matrix problem by either a direct or iterative solver.
+	\item Evaluation of domain potential operators for visualization and post-processing.
+\end{enumerate}
+
+All of these steps are accelerated through the use of either Numba or OpenCL.
+In the following section we provide a high-level overview of the library and how these acceleration techniques are deployed before we dive into the design of the computational kernels.
+
+\section{A HIGH-LEVEL OVERVIEW OF BEMPP-CL}
+
+\begin{figure}
+	\centering
+	\input{img/bempp_overview}
+	\caption{The layout of Bempp-cl with its computational backends.}
+	\label{fig:overview}
+\end{figure}
+
+The main user-visible component of Bempp-cl is the module \pyth{bempp.api}, which defines all user interface functions and other high-level routines.
+In particular, it contains the definitions of the main object types: \mbox{\pyth{Grid},} \mbox{\pyth{Space},} \mbox{\pyth{GridFunction},} and \pyth{Operator}.
+The computational backend routines are contained in the module \pyth{bempp.core}.
+Currently, we support Numba and OpenCL backends.
+An overview of this structure is provided in \cref{fig:overview}.
+
+The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles.
+For larger problems, this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity.
+The price of the improved complexity is significantly more involved data structures and additional approximation errors.
+Bempp-cl provides interfaces to ExaFMM \cite{bempp_exafmm} for large problems.
+Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory.
+Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as efficiently as possible.
+
+Once a user has defined an operator using Bempp-cl, the discretization can be computed by calling the \pyth{weak_form} method.
+Upon calling this method, a regular integrator will be used to assemble all the interactions between non-adjacent elements, and a singular integrator will be used to compute the interactions between adjacent triangles (if the trial and test spaces are defined on different grids, this second integrator is not needed).
+Depending on the user's preferences, these integrators will internally use computational routines defined using either Numba or OpenCL.
+
+For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision.
+The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed.
+If the computational backend is Numba, the call is forwarded to the corresponding Numba discretization routines and executed.
+
+For simple piecewise constant function spaces or other spaces, where the support of each basis function is localized to a single triangle, only one call to the computational routines is necessary.
+If the support of basis functions is larger than a single triangle, different threads may need to sum into the same global degree of freedom.
+
+Outside the operator discretization, Numba is used in the following contexts:
+\begin{itemize}
+	\item Computing the grid topology: this involves iterating through the grid to compute the neighbour relationships between triangles.
+	\item Definition of local-to-global maps for function spaces: again, this requires traversal through the grid and assigning relationships between global and local indices.
+	\item Grid functions: a right-hand side function $f$ can be defined as a Python callable.
+This is just-in-time compiled via Numba and then the product with the corresponding basis functions is integrated in each triangle via numerical quadrature, again via Numba accelerated routines.
+	\item Computing sparse matrix entries, such as for mass matrices that are required to translate between representations of grid functions through basis coefficients or projections, or when we want to evaluate operator products.
+\end{itemize}
+
+As the cost of each of these processes is in general much smaller than the cost of operator discretization, these can be performed using Numba without any need to consider the use of OpenCL for potential further speed up.
+
+\section{ASSEMBLING BOUNDARY INTEGRAL OPERATORS WITH OPENCL}
+
+In this section, we discuss in more detail the assembly of boundary integral operators with OpenCL
+and how we integrated this into our Python workflow. We start with a brief introduction to OpenCL and then
+dive into how we use OpenCL as part of Bempp-cl.
+
+\subsection{What is OpenCL?}
+
+OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers.
+At its core, OpenCL executes compute kernels that can be written in OpenCL C, which is based on C99, or (more recently) in C++, with some restrictions on the allowed operations.
+The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.
+
+OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel.
+Work-items are grouped together into work-groups, which share local memory.
+Barrier synchronization is only allowed within a work-group.
+All work-items are uniquely indexed by a one, two, or three dimensional index space, called \ocl{NDRange}.
+Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device.
+OpenCL allows kernels to be loaded as strings and compiled on-the-fly for a given device, making it well suited for launching from high-productivity languages.
+
+To launch an OpenCL kernel the user must provide relevant data as buffers, which are transferred from the host to the corresponding compute device.
+A kernel string can then be loaded and just-in-time compiled for the device.
+The kernel is then run, and the results can be copied back to the host.
+
+OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types.
+For example, the type \mbox{\ocl{double4}} will allow four double values to be held in a SIMD register.
+This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.
+
+Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner \cite{pyopencl}.
+PyOpenCL automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.
+
+\subsection{OpenCL Assembly in Bempp-cl}
+
+\begin{figure*}
+	\center
+	\input{img/kernel_header}
+	\caption{Definition of the OpenCL compute kernel for scalar integral equations.}
+	\label{fig:kernel_definition}
+\end{figure*}
+
+Bempp-cl has OpenCL kernels for all its boundary operators.
+All operators have the same interface and are launched in the same way.
+In the first step, the relevant data will need to be copied to the compute device.
+This data consists of:
+
+\begin{itemize}
+	\item Test and trial indices denoting the triangles over which to be integrated.
+	\item Signs of the normal directions for the spaces.
+	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
+	\item Test and trial connectivity, which are lists of node indices that define the corresponding triangles of the test and trial grid.
+	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
+	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g., in electromagnetics).
+	\item Quadrature points and quadrature weights.
+	\item A buffer that contains the global assembled matrix.
+	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
+	\item The number of test and trial degrees of freedom.
+	\item A single byte that is set to one if the test and trial grids are disjoint.
+\end{itemize}
+
+An example kernel definition is shown in \cref{fig:kernel_definition}.
+Before the kernel can be launched, it needs to be configured and just-in-time compiled.
+Kernel configuration happens through C-style preprocessor definitions that are passed through the just-in-time compiler.
+These include the names of the test and trial space, the name of the function that evaluates the Green's function, whether we are using single or double precision types, and (for SIMD enhanced kernels) the vector length of the SIMD types.
+For example, in \cref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}. This is substituted with either \ocl{float} or \ocl{double} during just-in-time compilation.
+
+Each work-item computes (using numerical quadrature) all interactions of basis functions on the trial element with basis functions on the test element.
+Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangles are adjacent or identical (see \cref{fig:triangles}).
+To do this, it simply checks if at at least one of the node indices of the test triangle is equal to one of the node indices of the trial triangle.
+If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer: for these triangles, separate singular quadrature rules need to be used.
+The effect is that a few work-items do work that is discarded.
+However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of triangle interactions is $N^2$.
+Hence, only a tiny fraction of work-items are discarded.
+
+\subsection{SIMD optimized kernels}
+When we are running on a CPU and want to take advantage of available SIMD optimizations, we need to make a few modifications to our approach.
+The corresponding kernel works similarly to what is described above, but we compute a batch of interactions between one test triangle and $X$ trial triangles, where $X$ is either 4, 8, or 16 (depending on the number of available SIMD lanes).
+This strategy allows us to optimize almost all floating point operations within a kernel run for SIMD operation.
+If the number of trial elements is not divisible by 4, 8, or 16, then the few remaining trial elements are assembled with the standard non-vectorized kernel.
+
+Each kernel definition is stored in two variants, one with the ending \texttt{\_novec.cl} and another one with the ending \texttt{\_vec.cl}.
+The vectorized variant is configured via preprocessor directives for the desired number of vector lanes.
+Having to develop two OpenCL kernel codes for each operator creates a certain amount of overhead, but once we have implemented the non-vectorized version then, with the help of preprocessor directives and a number of helper functions that do the actual implementation of operations depending on whether the kernel is vectorized or not, it is usually only a matter of an hour or two to convert the non-vectorized kernel into a vectorized version.
+
+Alternatively, some CPU OpenCL runtime environments can (optionally) try and auto-vectorize kernels by batching together work-items on SIMD lanes, similar to what we do manually.
+In our experience, this works well for very simple kernels but often fails for more complex OpenCL kernels.
+This is why we decided to implement this strategy manually.
+
+A completely different SIMD strategy could be taken by batching together quadrature evaluations within a single test/trial pair.
+There are two disadvantages to this approach: first, it only works well if the number of quadrature points is a multiple of the available SIMD lanes.
+Second, other operations such as the geometry calculations for each element then cannot be SIMD optimized as these are only performed once per test/trial pair.
+
+\subsection{Assembling the singular part of integral operators}
+The assembly of the singular part of an integral operator works a bit differently.
+Remember that the singular part consists of triangle parts which are adjacent to each other or identical (the three later cases in \cref{fig:triangles}): there are $\mathcal{O}(N)$ such pairs.
+We are using fully numerical quadrature rules for these integrals that are based on subdividing the four-dimensional integration domain and using transformation techniques to remove the singularities.
+This gives highly accurate evaluation of these integration pairs but requires a large number (typically over $1000$) quadrature points per triangle pair.
+
+For this assembly, we create one work-group for each singular triangle pair.
+Inside this work-group, we have a number of work-items that evaluate the quadrature rules then sum up the results.
+Depending on how two triangles are related to each other, different types of singular quadrature rule are needed.
+We solve this by pre-loading all possible quadrature rules onto the device, and also store for each triangle pair an index pointing to the required quadrature rule so that the kernel function can select the correct rule to evaluate.
+For the singular quadrature rules, we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above.
+At the end, the singular integral values are either summed into the overall result matrix, or (if desired by the user) stored as separate sparse matrix.
+
+\begin{figure}
+	\center
+	\input{img/colouring}
+	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighboring triangles are the same color.
+Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors, so do not share any degrees of freedom.
+[The dots on each cell are included as a visual aid for anyone who prints this article in black and white.]}
+	\label{fig:colouring}
+\end{figure}
+
+\subsection{Avoiding data races in the global assembly}
+Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom.
+To solve this we use standard coloring techniques to split up the computations into chunks that access different data regions.
+To this end, we define two triangles as neighbors if they share at least one global degree of freedom.
+Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space.
+An example coloring is shown in \cref{fig:colouring}.
+
+The compute kernels can then be run color-by-color.
+OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations.
+In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colors in the test space.
+
+\section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}
+The main focus of Numba \cite{numba} within Bempp-cl is to provide accelerated implementations of routines with linear complexity, such as grid iterations, integration of functions over grids, or assembly of certain sparse matrices.
+However, we also provide a fall-back implementation of the OpenCL dense operator assembly in Numba.
+
+With Numba, we use loop parallelism: each loop iteration is the assembly of one test triangle with all trial triangles. We then parallelize over the test triangles through a parallel for-loop.
+Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations.
+However, a much smaller fraction of operations is SIMD optimized due to the lack of targeted SIMD constructs in Numba.
+
+We stress that while Numba provides backends not only for CPU, but also for ROCm and CUDA, we currently only use the CPU component of Numba.
+
+\section{PERFORMANCE BENCHMARKS}
+
+In this section, we provide a number of performance benchmarks.
+The tests were all run on a Dell Precision 7740 Workstation Laptop with 64 GB RAM.
+Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz.
+The CPU supports AVX, AVX2, and AVX-512.
+As a GPU we use an Nvidia Quadro RTX 3000 GPU.
+All benchmark tests were performed in Linux.
+For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment.
+All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba just-in-time compilers did not skew the results.
+Though hardly noticeable by users, for smaller experiments the compilation phase typically takes longer than the actual computation.
+
+\subsection{Dense Operator assembly}
+We start by benchmarking the dense operator assembly. We assemble the matrix $\dmat{A}$ defined by
+$$
+\dmat{A}_{ij} = \int_{\Gamma}\psi_i(\bx)\int_{\Gamma}\frac{\phi_j(\by)}{4\uppi |\bx -\by|}\ds[\by]\ds[\bx],
+$$
+with $\Gamma$ being the unit sphere.
+For the basis functions $\phi_j$ and test functions $\psi_i$, we compare two cases: piecewise constant functions for both (P0 functions); and nodal, piecewise linear, and globally continuous functions for both (P1 functions).
+In the P0 case, each triangle is associated with just one piecewise constant function.
+In the P1 case, each triangle is associated with $3$ linear basis functions, one for each node of the triangle.
+
+We first compare the OpenCL CPU performance for the Intel OpenCL runtime and the PoCL OpenCL runtime driver.
+We run the tests in single precision and double precision.
+The native vector width for both drivers in single precision is $8$, and in double precision is $4$, corresponding to AVX instructions.
+In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision.
+Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this.
+We should hence see up to a factor 2 speed-up between single and double precision assembly.
+
+\begin{figure}
+	\center
+	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp.pdf}
+	\caption{Comparison of the performance of PoCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
+	\label{fig:intel_pocl_laplace_cmp}
+\end{figure}
+
+The timings for a grid with 32,768 elements are shown in \cref{fig:intel_pocl_laplace_cmp}.
+We see the expected speed-up between single precision and double precision evaluation.
+It is interesting to note the difference between the Intel and the PoCL runtime environment.
+For P0 basis functions, the PoCL driver (gray bar) significantly outperforms the Intel driver (orange bar) in both single and double precision.
+For P1 basis functions, however, the Intel driver gives better performance.
+
+In \cref{fig:pocl_single_layer}, we compare specifically the performance of single precision and double precision evaluation for the PoCL driver for various grid sizes in the case of a P0 basis.
+We can see that the speed-up for larger grid sizes is slightly more than just a factor of two.
+The final data point in this graph corresponds to the grid size used for \cref{fig:intel_pocl_laplace_cmp}.
+
+\begin{figure}
+	\center
+	\includegraphics[width=7cm]{img/pocl_single_layer.pdf}
+	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using PoCL and P0 basis functions.}
+	\label{fig:pocl_single_layer}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\begin{tabular}{l|c|c}
+			&   single      &    double\\
+			\hline
+			PoCL   &   0.05s       &    0.08s\\
+			Numba  &   0.25s       &    0.25s\\
+			GPU    &   0.11s       &    3.00s\\
+		\end{tabular}
+	\end{center}
+	\caption{Comparison of PoCL, Numba and GPU assembly for a grid with 2048 elements.}
+	\label{fig:cpu_gpu_numba_compare}
+\end{figure}
+In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly.
+\Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements.
+The speed differences are striking.
+GPU assembly in single precision is a factor of two slower than CPU assembly, and much slower in double precision due to the limited double precision performance of our hardware, though we note that this is not an issue for double precision optimised data center accelerators.
+Numba is five times slower for single precision and still around three times slower for double precision than PoCL.
+The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance.
+Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with tens of thousands of rows and columns on the device.
+
+An alternative method is to compute the matrix-vector product (matvec) $\dmat{A}\dvec{x}$ on the device without first computing the dense matrix $\dmat{A}$.
+This can be done by recomputing all matrix elements during each matvec calculation on-the-fly and not storing them.
+We have done experiments with this, and observed significant speed-ups compared to CPU evaluation as we now only need to transfer single vectors over the bus.
+For larger problems, however, it is not competitive compared to accelerated methods such as FMM, due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM.
+For smaller problems, it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver.
+Hence, we can conclude that there is only limited practical relevance for on-the-fly GPU evaluation of boundary integral operators.
+There are, however, very significant practical advantages of on-the-fly evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.
+
+The Numba performance difference seen in \cref{fig:cpu_gpu_numba_compare} is interesting.
+The main reason for this, we believe, is significantly lower usage of AVX vectorized instructions.
+We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangles for a single test triangle, as we currently do.
+We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already.
+We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available (this is less a judgement about Numba itself but about the limited optimisations we have done for Numba assembly routines).
+
+\subsection{Evaluating domain potentials for post-processing of electromagnetic problems}
+Once an integral equation is solved, one is usually interested in evaluating the solution not only at the boundary but also at points away from it. To do this, we evaluate the integral
+$$
+f(\bx) = \int_{\Gamma}g(\bx, \by)\phi(\by) \ds[\by]
+$$
+for many points $\bx$ away from the boundary $\Gamma$. For example, if we want to visualize a solution, we take the points $\bx$ to be a regular grid of points.
+
+Typically, we want to do only a small number of potential evaluations at the end of a calculation.
+Discretising this operation into a dense matrix and then evaluating the dense matrix-vector product is not practical for larger sizes.
+For very large problems with hundreds of thousands of elements, we use FMM or other accelerated approximate methods.
+For moderately sized problems with a few ten thousand elements up to around a hundred thousand elements (depending on the problem at hand), direct evaluation of this integral for every point $\bx$ is highly efficient.
+We won't go into the details of the corresponding OpenCL kernels here, but we show some results that demonstrate the relative performance on CPUs and GPUs.
+
+For the dense assembly of boundary integral operators, the performance was limited by the bus transfer of the dense matrix.
+For the evaluation of domain potentials, however, we only need to transfer to the device the vector of coefficients of the basis functions for $\phi$, and then transfer back to the host the values at the points $\bx$.
+
+In this section, we consider the evaluation of the electric potential operator, defined by
+\begin{multline}
+	\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
+	-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
+\end{multline}
+where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\uppi|\bx - \by|}$ is the Helmholtz Green's function, and
+$k$ the wavenumber of the problem.
+The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$.
+The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}.
+For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao--Wilton--Glisson) edge-based basis functions are used.
+For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.
+
+In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver.
+In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double precision operations.
+
+\begin{figure}
+	\center
+	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
+	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs Nvidia GPU}
+	\label{fig:efield_domain_potential}
+\end{figure}
+
+\section{SUMMARY}
+With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies.
+Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines.
+Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the ExaFMM library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.
+
+Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models.
+OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU- and GPU-based compute devices with a simple parameter change.
+In this article, we demonstrate Nvidia benchmarks: the same benchmarking code could be used to run on AMD or Intel GPUs.
+
+A disadvantage of our approach is that using OpenCL kernels introduces a second language (C99) to the library.
+Using Numba throughout would give a much more native Python experience, but, while Numba is constantly improving, it is currently difficult to achieve optimal performance for complex operations.
+OpenCL really shines here, as it makes explicit SIMD operations very easy through dedicated constructs.
+Moreoever, OpenCL kernels are completely stack/register based functions, allowing much better compiler optimisations while Numba needs to create every object dynamically, even for very small arrays for objects such as coordinate vectors.
+We need to stress that we have performed very few optimisations specific to Numba, while significant optimisation has gone into the OpenCL codes.
+It is therefore well possible that the performance gap between Numba and OpenCL can be significantly reduced.
+But from other projects our own anecdotal experience is that the more Numba is optimised, the less Pythonic and more C-like Numba functions look.
+So while Numba is a very powerful tool, it requires its own techniques for optimisation, different from standard Python code.
+
+Another important consideration with respect to Python and just-in-time acceleration is the type of algorithms that benefit.
+For the dense assembly of integral operators, we have very simple data structures that can easily be passed to compute kernels.
+More complex data structures with larger mixture of data movement operations and computations (e.g., tree-based algorithms), are much harder to accelerate since the Python layer imposes limits here on the performance.
+
+Overall, with the model of mixed Python/OpenCL/Numba development, we have created a flexible and easy to extend platform for integral equation computations.
+The initial re-implementation efforts by abandoning our old C++ code base are paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance.
+Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.
+
+The current focus of further developments is on letting Bempp-cl take advantage of cluster computing by integrating the \pyth{mpi4py} MPI bindings for Python.
+We have also made big steps forward for large problems by creating a black-box FMM interface that currently interfaces to ExaFMM: this has allowed us to solve problems with 10 million elements on a single workstation.
+We believe that Python-focused development (with some native routines in lower-level languages) is a scalable model and are aiming to exploit this scalability further as we move from single workstation computations to large cluster problems.
+
+\section{Acknowledgements}
+The work of Timo Betcke is supported by Engineering and Physical Sciences Research Council Grant EP/V001531/1.
 
@@
 
+%first
+\bibitem{bempp_orig}
+W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
+``Solving boundary integral problems with BEM++,''
+{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, article 6, 2015.
+\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}
+
+%second
 \bibitem{Bempp-cl}
@@
 
+%third
 \bibitem{erichsen}
@@
 
+%fourth
+\bibitem{bempp_exafmm}
+T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba,
+``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-ExaFMM,''
+2021.
+arXiv (preprint): \href{https://arxiv.org/abs/2103.01048}{2103.01048}.
+
+%fifth
+\bibitem{opencl}
+OpenCL, \url{https://www.khronos.org/opencl/}.
+
+%sixth
 \bibitem{pyopencl}
@@
 
+%seventh
 \bibitem{numba}
@@
 
-\bibitem{opencl}
-OpenCL, \url{https://www.khronos.org/opencl/}.
-
+%eighth
 \bibitem{bempp_maxwell}
@@
 
-\bibitem{bempp_orig}
-W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
-``Solving boundary integral problems with BEM++,''
-{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, article 6, 2015.
-\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}
-
-\bibitem{bempp_exafmm}
-T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba,
-``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-ExaFMM,''
-2021.
-arXiv (preprint): \href{https://arxiv.org/abs/2103.01048}{2103.01048}.
 \end{thebibliography}

### 2021-05-28 numba.tex
Have written reviewer responses
Updated references and merged files

@@
 \section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}
-The main focus of Numba within Bempp-cl is to provide accelerated implementations of routines with linear complexity, such as grid iterations, integration of functions over grids, or assembly of certain sparse matrices.
+The main focus of Numba \cite{numba} within Bempp-cl is to provide accelerated implementations of routines with linear complexity, such as grid iterations, integration of functions over grids, or assembly of certain sparse matrices.
 However, we also provide a fall-back implementation of the OpenCL dense operator assembly in Numba.
 
-With Numba, we use loop parallelism: each loop iteration is the assembly of one test triangle with all trial triangles. We then parallelize over the test triangles.
-All such assemblies are executed through a parallel for-loop.
+With Numba, we use loop parallelism: each loop iteration is the assembly of one test triangle with all trial triangles. We then parallelize over the test triangles through a parallel for-loop.
 Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations.

### 2021-05-28 final_submission/CISESI-2021-04-0035.R1_Betcke.tex
Added final files

--- 2021-05-28 main.tex
@@
 	\centering
-	\input{img/sphere}
+	\input{sphere}
 	\caption{Discretization of a sphere into flat surface triangles.}
@@
 	\centering
-	\input{img/triangles}
+	\input{triangles}
 	\caption{The four types of intersection of two triangles: the triangles can be (left to right) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle.
@@
 	\centering
-	\input{img/bempp_overview}
+	\input{bempp_overview}
 	\caption{The layout of Bempp-cl with its computational backends.}
@@
 	\center
-	\input{img/kernel_header}
+	\input{kernel_header}
 	\caption{Definition of the OpenCL compute kernel for scalar integral equations.}
@@
 	\center
-	\input{img/colouring}
+	\input{colouring}
 	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighboring triangles are the same color.
@@
 	\center
-	\includegraphics[width=6cm]{img/intel_pocl_laplace_comp.pdf}
+	\includegraphics[width=6cm]{intel_pocl_laplace_comp.pdf}
 	\caption{Comparison of the performance of PoCL and the Intel OpenCL runtime for the assembly of the Laplace single-layer boundary operator on a grid with 32,768 elements.}
@@
 	\center
-	\includegraphics[width=7cm]{img/pocl_single_layer.pdf}
+	\includegraphics[width=7cm]{pocl_single_layer.pdf}
 	\caption{Comparison of the single-precision and double-precision performance for various grid sizes using PoCL and P0 basis functions.}
@@
 	\center
-	\includegraphics[width=7cm]{img/efield_domain_potential.pdf}
+	\includegraphics[width=7cm]{efield_domain_potential.pdf}
 	\caption{Evaluation of an electric field potential operator on CPU via PoCL vs Nvidia GPU}
